### 2022-05-11 README.Rmd
first commit

---
title: ohun_paper_analyses
author: AUTOR NAME
output: github_document
date: `r format(Sys.Date(), "%d-%m-%Y")`
editor_options:
  chunk_output_type: console
---

<!-- Short Description  -->
This is a short description of the project

*Updated on `r Sys.time()`*

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}

library(rmarkdown)

knitr::opts_chunk$set(
  collapse = TRUE,
  out.width = "100%"
)
```

## Table of contents
* [General info](#general-info)
  * [Analyses](#Analyses)
    * [To do list](#to-do-list)
      * [Contact](#contact)

## General info

General information about the project (study system, goals, etc)

## Analysis

* Analysis 1

## In progress

*  Something not finished yet

## To-do list

* Do the next analysis

## Status
Project is: _in progress_

## Contact
Created by [Autor name](website_URL)

### 2022-05-11 manuscript/manuscript.Rmd
first commit

---
title: 'Ohun: an R package for optimizing automatic acoustic signal detection'
author:
 - Marcelo Araya-Salas ^1^ * 
 - Grace Smith-Vidaurre ^2^ 
 - Gloriana Chaverri ^3^
 - Fabiola Chirino ^1^
 - Alejandro Rico-Guevara ^4^ 
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: no
bibliography: combined_bibs.bib
citation-style: apa.csl
link-citations: true
editor_options: 
  chunk_output_type: console
---

```{r, fixing citation .bib files, eval = TRUE, echo = FALSE}

# how to cite in Rmarkdown
# https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html

# Here include the path to the libraries for each author. This code will copy/update the library each time the Rmarkdown file is compiled
# Alternatively authors ca
personal_libraries <- c(author_1 = "path/to/lib.file/from/author1/library.bib", author_2 = "path/to/lib.file/from/author2/library.bib")

# update bibtex library
for (i in 1:length(personal_libraries))
if (file.exists(personal_libraries[i]))
 file.copy(from = personal_libraries[i], to = file.path(getwd(), paste0(names(personal_libraries)[i], ".bib")), overwrite = TRUE)

## combine .bib files in compiled.bib ###
# list bib files
bibs <- list.files(pattern = ".bib$")

# remove combined from pool
bibs <- bibs[bibs != "combined_bibs.bib"]

#  read in a list
if (length(bibs) > 0){
  combined_bibs_l <- lapply(bibs, readLines)
  names(combined_bibs_l) <- gsub("\\.bib$", "", bibs)

# combine bibs in a single one
combined_bibs <- unlist(combined_bibs_l)

# save pool libraries as combined_bibs.bib
writeLines(text = combined_bibs, "combined_bibs.bib")

# check if there are duplicated keys
keys <- grep("@article{", combined_bibs, fixed = TRUE, value = TRUE)
keys <- gsub("@article{", "", keys, fixed = TRUE)
keys <- gsub(",", "", keys, fixed = TRUE)

tab_keys <- table(keys)

if (anyDuplicated(keys)){
  print(paste0(sum(tab_keys > 1), " duplicate(s) references found in combined_bibs.bib"))
}
} else writeLines(text = "empty", "combined_bibs.bib")

```

^1^ Affiliation # 1

^2^ Affiliation # 2

\*To whom correspondence should be addressed

### Keywords:

Check [this link](https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html) for details on how to add citations/bibliography in Rmarkdown

---

## Abstract

## Introduction

- 1st paragraph:

    - Animal acoustic signals are widely used in scientific research in a highly diverse areas, ranging from neurobiology to community ecology and evolutionary biology. 
    - The wide usage of animal sounds in research is partly due to the fact that they can be easily registered using non-intrusive methods, in a variety of setting from laboratories to natural areas and equipment for recording these signals has become increasingly inexpensive and accesible.

    In addition, the existence of online repositories and growing number of acoustic data set avialable to the comunnity. As a result, a growing variety of a computational tools for the analysis of acoustic features on those signals is increasingly available, which reflects the growing relevance of bioacustics as a relevant tool in the scientific tool kit. (many of those in the form of freely available software written on programming environments familiar to the scientific community (seewave, warbleR, scikit-mad))

- 2nd paragraph:

    - Bioacoustic approaches can easily generate large amounts of data, which sometimes may proof challenging to analyze in a timely manner.
    - Much effort has been put into developing computational tools that can simplify acoustic analyses. For instance,....
    - However, some key aspects of the analysis remain challenging, despite the great advances witness in the last decade.
    - The annotations that require a precise definition of signal location in frequency and time, most commonly used in behavioral and phenotypic eovlution research, remain among the most time consuming phases of analysis for this type of data. Fortunately, several tools have been develop to assist researchers in this matter. Indeed, new methods for automatic annotation of acoustic signals are contnuously developed. This growing availability of tools, particularly as free software, is expected to further simplify acoustic data processing, making it accesible to wider range of user and scientific questions.

- 3rd paragraph:

    - Here we present the new R package ohun. The package is intended to facilitate the automatic detection of acoustic signals, providing functions to diagnose and optimize detection routines. The package makes use of reference annotations containing the time position of all target signals in a training data set to evaluate the performance of detection routines. This can be done with routine outputs imported from other software as well as detection run with ohun. The package offers an implementation of two automatic detection methods commonly used in bioacustic analysis: energy-based detection and template-based detection. We first explain how acoustic signal detection routines can be evaluate and then showcase the package usage with study cases on zebra-finch songs (*Taenopygia gutata*) and Spix's disc-winged bats (*Thyroptera tricolor*) which represent different recording settings (lab and flight cages) and signal types (sonic mating signals and ultrasonic social calls).

## Diagnosing detection performance

The package makes use of signal detection theory indices to evaluate detection performance. Signal detection theory deals with the process of recovering signals (i.e. target signals) from background noise (not necessarily acoustic noise) and it’s widely used for optimizing this decision making process in the presence of uncertainty. During a detection routine, the detected ‘items’ can be classified into 4 classes: true positives (TPs, target signals correctly identified as signal), false positives (FPs, noise incorrectly identified as ‘signal’), false negatives (FNs, signals incorrectly identified as noise) and true negatives (TNs, background noise correctly identified as noise). However, TNs cannot always be easily defined in the context of acoustic signal detection as noise cannot always be partitioned in discrete units. Hence, the package makes use of TPs, FPs and FNs to calculate three additional indices and can further assist with evaluating the performance of a detection routine.

 - Recall: correct detections relative to total detections (a.k.a. true positive rate or sensitivity; TPs / (TPs + FNs))
- Precision: correct detections relative to total detections (TPs / (TPs + FPs)).
- F1 score: combines recall and precision as the harmonic mean of these two, so it provides a single value for evaluating performance (a.k.a. F-measure or Dice similarity coefficient).

A perfect detection will have no false positives or false negatives, which will result in both recall and precision equal to 1. However, perfect detection cannot always be reached and some compromise between detecting all target signals plus some noise (recall = 1 & precision < 1) and detecting only target signals but not all of them (recall < 1 & precision = 1) might be warranted. The right balance between these two extremes will be given by the relative costs of missing signals and mistaking noise for signals. Hence, these indices provide an useful framework for diagnosing and optimizing the performance of a detection routine.

The package ohun provides a set of tools to evaluate the performance of an acoustic signal detection based on the indices described above. To accomplish this, the result of a detection routine is compared against a reference table containing the time position of all target signals in the sound files.

## Signal detection

Finding the position of signals in a sound file is a challenging task. ohun offers two methods for automatic signal detection: template-based and energy-based detection. These methods are better suited for highly stereotyped or good signal-to-noise ratio (SNR) signals, respectively. If the target signals don’t fit these requirements, more elaborated methods (i.e. machine learning approaches) are warranted:

## Study cases

### Ultrasonic social calls in Spix's disc-winged bats

We recorded 30 individuals of Spix's disc-winged bats at Baru Biological Station, in south-western Costa Rica in January 2018. Bats were captured at their roosting sites (furled leaves of Zingeberaceae plants). Each individual bat was released in a large flight cage (9 x 4 x 3 m) for a 5 min period and their ultrasonic inquiry calls were recorded using a condenser microphone (CM16, Avisoft Bioacoustics, Glienike/
Nordbahn, Germany) through an Avisoft UltraSoundGate 116Hm connected to a laptop computer running Avisoft-Recorder software (sampling rate of 500 kHz, 16-bit amplitude resolution).

Recordings were manually annotated using Raven Pro 1.6 (XXXX). Annotations were created by visual inspection of spectrograms, in which the start and end of signals were determined by the location of the continuous traces of power spectral entropy of target signals. Annotations were made with a time window of 512 samples with a 70% of overlap. Annotations were then imported into R using the R package Rraven (XXXX).

Inquiry calls of Spix's disc-winged bats are stereotyped signals. Most variation is found between individuals although the basic structure of a short, downward broadband frequency modulation (FIG XXX) is always shared. Template-based detection is an useful approach when there is little structural differences in the target signals. Therefore we used the function

### Zebra finch songs

We also used recordings from 18 zebra finch males. These recordings were obtained ....

## Results

## Discusion

- Despite signal detection indices being commonly reported when presenting new automatic detection methods, to our knowledge widely applicable performance evaluating routines have not been made available.
- Can be useful for analysis in which relatively small number of recordings are used and also when only a limited amount of training data is available.

## Improving detection speed

Detection routines can take a long time when working with large amounts of acoustic data (e.g. large sound files and/or many sound files). These are some useful points to keep in mine when trying to make a routine more time-efficient:

- Always test procedures on small data subsets
- `template_detector()` is faster than `energy_detector()`
- Parallelization (see `parallel` argument in most functions) can significantly speed-up routines, but works better on Unix-based operating systems (linux and mac OS)
- Sampling rate matters: detecting signals on low sampling rate files goes faster, so we should avoid having nyquist frequencies (sampling rate / 2) way higher than the highest frequency of the target signals (sound files can be downsampled using warbleR's [`fix_sound_files()`](https://marce10.github.io/warbleR/reference/selection_table.html))
- Large sound files can make the routine crash, use `split_acoustic_data()` to split both reference tables and files into shorter clips.
- Think about using a computer with lots of RAM memory or a computer cluster for working on large amounts of data
- `thinning` argument (which reduces the size of the amplitude envelope) can also speed-up `energy_detector()`

## Additional tips

- Use your knowledge about the signal structure to determine the initial range for the tuning parameters in a detection optimization routine
- If people have a hard time figuring out where a target signal occurs in a recording, detection algorithms will also have a hard time
- Several templates representing the range of variation in signal structure can be used to detect semi-stereotyped signals
- Make sure reference tables contain all target signals and only the target signals. The performance of the detection cannot be better than the reference itself.
- Avoid having overlapping signals or several signals as a single one (like a multi-syllable vocalization) in the reference table when running an energy-based detector
- Low-precision can be improved by training a classification model (e.g. random forest) to tell signals from noise

https://cran.r-project.org/web/packages/bioacoustics/vignettes/tutorial.html

## References

### 2022-05-11 scripts/study_case_taeniopygia.Rmd
first commit

---
title: <center><font size="6"><b>Energy-based automatic detection on zebra finch songs</b></font></center>
subtitle: <center><font size="4"><b>ohun</b></font></center>
author: <center><font size="4"><a href="http://marceloarayasalas.weebly.com/">Marcelo Araya-Salas, PhD</a></font></center>
date: <center>`r format(Sys.Date(), "%d-%m-%Y")`</center>
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: no
fontsize: 12pt 
editor_options: 
  chunk_output_type: console
---

&nbsp;

<div class="alert alert-info">

### Purpose

- Showcase energy-based automatic detection with the R package ohun using zebra-finch songs

</div>

&nbsp;

```{r setup, include = FALSE}

knitr::opts_knit$set(root.dir = normalizePath(".."))

knitr::opts_chunk$set(include = TRUE, eval = TRUE)

data_path <- raw_data_path <- "data/raw/taeniopygia"

preproc_path <- "data/processed/taeniopygia"

figure_path <- "data/processed/taeniopygia/figures"

```

### Load package
```{r libraries, eval = TRUE, echo = FALSE, message=FALSE, warning=FALSE}

x <- c("ohun", "Rraven", "warbleR", "pbapply", "remotes", "DT")

out <- lapply(x, function(y) {

  # check if installed, if not then install 
  if (!y %in% installed.packages()[,"Package"]) 
    install.packages(y)

  # load package
  try(require(y, character.only = T), silent = T)
})

```

```{r, message=FALSE, warning=FALSE}
library(ohun)

library(warbleR)

library(viridis)
```

```{r manual selection table, eval = FALSE, echo = FALSE}

#Read in manual selections made in Raven

mnl_sel_tbl <- imp_raven(path = raw_data_path, name.from.file = TRUE, ext.case = "lower", all.data = TRUE)

glimpse(mnl_sel_tbl)

mnl_sel_tbl <- mnl_sel_tbl %>% 
  dplyr::rename(
    `start` = "Begin Time (s)",
    `end` = "End Time (s)",
    `selec` = "Selection",
    `selection_notes` = "Annotation"
  ) %>% 
  dplyr::mutate(
    bottom.freq = 0.5,
    top.freq = 10
  ) %>% 
  dplyr::select(sound.files, selec, start, end, selec.file, bottom.freq, top.freq, selection_notes)

glimpse(mnl_sel_tbl)

# Calculate signal to noise ratio (SNR)
mnl_sel_tbl_snr <- sig2noise(mnl_sel_tbl, mar = 0.001, type = 1, in.dB = TRUE, before = FALSE, lim.dB = TRUE, bp = c(unique(mnl_sel_tbl$bottom.freq), unique(mnl_sel_tbl$top.freq)), wl = 10, path = raw_data_path)

glimpse(mnl_sel_tbl_snr)

# Write out the combined manual selections out as a selection table (.csv)
write.csv(mnl_sel_tbl_snr, file.path(preproc_path, "manual_selections_Taeniopygia.csv"), row.names = FALSE)

```

```{r manual extended selection table, eval = FALSE, echo = FALSE}
#Write out the combined manual selections out as an extended selection table (RDS).

est <- selection_table(mnl_sel_tbl_snr, whole.recs = FALSE, extended = TRUE, mar = 0.01, parallel = cores, confirm.extended = FALSE, path = raw_data_path)

glimpse(est)
class(est)

# Update sound file names, which now have "_number" after the ".wav" extension
wav_nms <- names(attr(est, "wave.objects"))
wav_nms <- paste(gsub(".wav", "", wav_nms), ".wav", sep = "")
head(wav_nms) # looks good

est2 <- rename_est_waves(est, wav_nms)

# Looks good
head(names(attr(est2, "wave.objects")))
glimpse(est2)

# Remove the "old.sound.file.name" column
est2 <- est2[, -grep("old.sound.file.name", names(est2))]
glimpse(est2)

# Add in original sound file name, since each sound.file value is currently a unique name per row of the table
check_df <- attr(est2, "check.res")
glimpse(check_df)
head(check_df$sound.files)

# Are all sampling rates and bit rates the same?
# Yes: 44.1 kHz sampling rate and bit rate of 16
check_df %>%
  distinct(sample.rate, bits)

# Does the order of selections match between the EST and this attribute data frame? If can, can do the column addition below of original sound file name, start and end coordinates, selection duration
all(est2$sound.files == check_df$sound.files)

est2$duration <- check_df$duration
est2$original_sound_files <- check_df$orig.sound.files
est2$original_start <- check_df$orig.start
est2$original_end <- check_df$orig.end

# Looks good
glimpse(est2)
class(est2)
length(attr(est2, "wave.objects"))

saveRDS(est2, file.path(preproc_path, "manual_selections_Taeniopygia_est.RDS"))

```

```{r automated detection parameters, eval = FALSE, echo = FALSE}

est <- readRDS(file.path(preproc_path, "manual_selections_Taeniopygia_est.RDS"))
# glimpse(est)

mnl_sel_tbl <- read.csv(file.path(preproc_path, "manual_selections_Taeniopygia.csv"))
# glimpse(mnl_sel_tbl)

# Use ohun::feature_reference to get parameters that can help automated detection
# ?ohun::feature_reference

# I get an error when using the EST: "# Error in warbleR::gaps(X = reference, pb = FALSE) : 
  # extended selection tables must be created 'by.song' to be used in song.param()"
# feat_ref <- ohun::feature_reference(est, by.sound.file = FALSE,)

feat_ref <- ohun::feature_reference(mnl_sel_tbl, by.sound.file = FALSE)

# Min and max duration of the manual selections
# Use these for automated detection 
min_dur <- floor(feat_ref[["sel.duration", "min"]]) 
max_dur <- ceiling(feat_ref[["sel.duration", "max"]] )
min_gap <- floor(feat_ref[["gap.duration", "min"]])
bp <- c(feat_ref[["bottom.freq", "min"]] - 0.5, feat_ref[["top.freq", "max"]] + 0.5)

min_dur
max_dur
min_gap
bp

# A single file for testing
wav <- "R39Y212_43533.30691176_3_9_8_31_31.wav"

```

&nbsp;

### Set directory where the sound files and annotations are found

```{r, eval = FALSE, echo = TRUE}

data_path <- "DIRECTORY_WITH_SOUND_FILES_AND_ANNOTATIONS_HERE"

```

&nbsp;

### Read reference annotations
```{r, eval = TRUE, echo = FALSE}

manual_ref <- read.csv(file.path("data/processed/taeniopygia", "manual_selections_Taeniopygia.csv"))

```

```{r, eval = FALSE, echo = TRUE}

manual_ref <- read.csv(file.path(data_path, "manual_selections_Taeniopygia.csv"))

```

&nbsp;

### Create spectrograms to explore vocalization structure

This code creates a multipanel image with multiple spectrograms, one for each of the individuals/recordings in the complete data set
```{r Create catalogs, eval = FALSE}

# select highest signal to noise ratio calls per individual
manual_ref_snr <- signal_2_noise(X = manual_ref, mar = 0.05, path = data_path)

# select 1 example per sound file
high_snr <- manual_ref_snr[ave(-manual_ref_snr$SNR, manual_ref_snr$sound.files, FUN = rank) <= 1, ]

# create catalogs
catalog(X = high_snr, flim = c(0, 11), nrow = 6, ncol = 3, ovlp = 90, height = 15, width = 20, same.time.scale = FALSE, mar = 0.01, wl = 512, gr = FALSE, spec.mar = 0.4, lab.mar = 0.8, rm.axes = TRUE, by.row = TRUE, box = TRUE, pal = viridis, parallel = 10, collevels = seq(-100, 0, 5), path = data_path)

```

```{r catalog, out.width = "100%", echo = FALSE, fig.align= "center"}

knitr::include_graphics(normalizePath(file.path(data_path, "Catalog_p1.jpeg")))

```

&nbsp;

### Split reference annotations for training and testing detection
```{r test and train, eval = TRUE, echo = TRUE}

train_files <- c(
  "BRN7_43435.27985312_12_1_7_46_25.wav",
  "db1_43427.28100313_11_23_7_48_20.wav",
  "Gry35HP_43455.29800260_12_21_8_16_40.wav"
)
test_files <- setdiff(manual_ref$sound.files, train_files)

train_ref <- manual_ref[manual_ref$sound.files %in% train_files, ]
test_ref <- manual_ref[manual_ref$sound.files %in% test_files, ]
```

&nbsp;

### Optimize detection on training data
```{r  optimize automated detection, eval = FALSE, echo = TRUE}

opt_det <- optimize_energy_detector(
  reference = train_ref, 
  files = train_files, 
  threshold = c(1, 5), 
  hop.size = 11.6, 
  smooth = c(5, 10), 
  hold.time = c(0, 5), 
  min.duration = c(5, 15, 25), 
  max.duration = c(275, 300, 325), 
  bp = c(0.5, 10),
  path = data_path
)

```

```{r  save optimized automated detection, eval = FALSE, echo = FALSE}

write.csv(opt_det, file.path(preproc_path, "detection_optimization.csv"), row.names = FALSE)
```

&nbsp;

### Check subset with highest performance (sorted by f1 score)
```{r, eval = FALSE}
# subset with highest performance
sub_opt_det <- opt_det[order(opt_det$f1.score, decreasing = TRUE), ]

sub_opt_det <- sub_opt_det[1:10, c("threshold", "smooth", "hold.time", "min.duration", "max.duration", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f1.score")]

#print 
sub_opt_det

```

```{r, eval = TRUE, echo = FALSE}

opt_det <- read.csv(file.path(preproc_path, "detection_optimization.csv"))

# subset with highest performance
sub_opt_det <- opt_det[order(opt_det$f1.score, decreasing = TRUE), ]

sub_opt_det <- sub_opt_det[1:10, c("threshold", "smooth", "hold.time", "min.duration", "max.duration", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f1.score")]

# print dynamic table
oa_DT <- datatable(sub_opt_det, editable = list(target = "row"), rownames = FALSE, style = "bootstrap",
    filter = "top", options = list(pageLength = 100, autoWidth = TRUE, dom = "ft"),
    autoHideNavigation = TRUE, escape = FALSE)

formatRound(table = oa_DT, columns = sapply(sub_opt_det, is.numeric), 3)
```

&nbsp;

### Select the highest performance parameters based on f1 score
```{r, select best params, eval = TRUE, echo = TRUE}

best_param <- opt_det[which.max(opt_det$f1.score), ]

best_param
```

&nbsp;

### Run detection on test data
```{r, detect on all data, eval = TRUE, echo = TRUE}

opt_det_all <- energy_detector(
  files = test_files, 
  threshold = best_param$threshold, 
  hop.size = 11.6, 
  smooth = best_param$smooth, 
  hold.time = best_param$hold.time, 
  min.duration = best_param$min.duration, 
  max.duration = best_param$max.duration, 
  bp = c(0.5, 10), 
  path = data_path
)
```

&nbsp;

### Evaluate detection performance
```{r, diagnose, eval = TRUE, echo = TRUE}

diagnose_detection(reference = test_ref, detection = opt_det_all, by.sound.file = FALSE)

```

&nbsp;

<div class="alert alert-info">

### Takeaways

- Good performance on detecting zebra finch songs: F1 score was 0.95 for the training data set and 0.96 for the testing data

</div>

&nbsp;

---

<font size="4">Session information</font>

```{r session info, echo=F}

sessionInfo()

```

### 2022-05-11 scripts/study_case_thyroptera.Rmd
first commit

---
title: <center><font size="6"><b>Template-based automatic detection on Spix's disc-winged bat social calls</b></font></center>
subtitle: <center><font size="4"><b>ohun</b></font></center>
author: <center><font size="4"><a href="http://marceloarayasalas.weebly.com/">Marcelo Araya-Salas, PhD</a></font></center>
date: <center>`r format(Sys.Date(), "%d-%m-%Y")`</center>
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: no
fontsize: 12pt 
editor_options: 
  chunk_output_type: console
---

&nbsp;

<div class="alert alert-info">

### Purpose

- Showcase template-based automatic detection with the R package ohun using Spix's disc-winged bat social calls

</div>

&nbsp;

---

### Load packages
```{r packages, message = FALSE, warning = FALSE, echo = FALSE, eval = TRUE, message = FALSE, warning=FALSE}

## add 'developer/' to packages to be installed from github
x <- c("remotes", "maRce10/warbleR", "bioacoustics", "viridis", "ggplot2", "maRce10/ohun", "DT")

aa <- lapply(x, function(y) {

  # get pakage name
  pkg <- strsplit(y, "/")[[1]]
  pkg <- pkg[length(pkg)]

  # check if installed, if not then install 
  if (!pkg %in% installed.packages()[,"Package"])  {

      if (grepl("/", y))  remotes::install_github(y, force = TRUE) else
    install.packages(y) 
    }

  # load package
  try(require(pkg, character.only = T), silent = T)
})

```

```{r chunk opts, eval = TRUE, echo = FALSE}

opts_knit$set(root.dir = "..")

opts_chunk$set( fig.width = 8, fig.height = 3.5)

data_path <- "./data/raw/thyroptera/"
```

```{r, message=FALSE, warning=FALSE}
library(ohun)

library(warbleR)

library(ggplot2)

library(viridis)
```

```{r functions, eval = FALSE, echo = FALSE}

warbleR_options(wav.path = "./data/raw/thyroptera/", wl = 300, parallel = 1, bp = "frange", fast = F, threshold = 15, ovlp = 20)

```

```{r get sample of recordings from flight coordination paper, eval = FALSE, echo = FALSE}
fixed_calls <- read.csv( "./data/processed/manually_fixed_call_detections.csv")
table(fixed_calls)
table(fixed_calls$detection)
length(unique(fixed_calls$sound.files))

# p <- 1
# x <- 1
# while (p > 0.02){
# sub_calls <- fixed_calls[fixed_calls$sound.files %in% sample(unique(fixed_calls$sound.files), 30), ]

set.seed(7)
p <- sum(sub_calls$detection == "manual") / nrow(sub_calls)
# x <- x + 1
# }
sub_calls$channel <- 1
sub_calls <- sub_calls[, c("sound.files", "selec", "channel", "start", "end", "bottom.freq", "top.freq")]

sub_calls$selec <- 1:nrow(sub_calls)

write.csv(sub_calls, "~/Dropbox/Projects/ohun_paper/data/raw/thyroptera/manual_annotations_thyroptera.csv", row.names = FALSE)

wvs <- unique(sub_calls$sound.files)

file.copy(to = file.path("~/Dropbox/Projects/ohun_paper/data/raw/thyroptera/", wvs), from = file.path("~/Dropbox/Recordings/flight_coordination_Thyroptera/converted_sound_files_90_kHz/", wvs))

```

&nbsp;

### Set directory where the sound files and annotations are found

```{r, eval = FALSE, echo = TRUE}

data_path <- "DIRECTORY_WITH_SOUND_FILES_AND_ANNOTATIONS_HERE"

```

&nbsp;

### Read reference annotations
```{r read data, eval = TRUE, echo = FALSE}
# read data
manual_ref <- read.csv("./data/raw/thyroptera/manual_annotations_thyroptera.csv", stringsAsFactors = FALSE)

```

```{r read data 2, eval = FALSE, echo = TRUE}

manual_ref <- read.csv(file.path(data_path, "manual_annotations_thyroptera.csv"), stringsAsFactors = FALSE)
```

&nbsp;

### Create spectrograms to explore vocalization structure

This code creates a multipanel image with multiple spectrograms, one for each of the individuals/recordings in the complete data set
```{r Create catalogs, eval = FALSE}

# select highest signal to noise ratio calls per individual
manual_ref_snr <- signal_2_noise(X = manual_ref, mar = 0.05)

# select 1 example per sound file
high_snr <- manual_ref_snr[ave(-manual_ref_snr$SNR, manual_ref_snr$sound.files, FUN = rank) <= 1, ]

# create catalogs
catalog(X = high_snr, flim = c(10, 45), nrow = 6, ncol = 5, ovlp = 90, height = 15, width = 20, same.time.scale = TRUE, mar = 0.005, wl = 512, gr = FALSE, spec.mar = 0.4, lab.mar = 0.8, rm.axes = TRUE, by.row = TRUE, box = TRUE, pal = viridis, parallel = 10, collevels = seq(-100, 0, 5))

```

```{r catalog, out.width = "100%", echo = FALSE, fig.align= "center"}

knitr::include_graphics(normalizePath(file.path(data_path, "Catalog_p1.jpeg")))

```

&nbsp;

### Split reference annotations for training and testing detection
```{r test and train, eval = TRUE, echo = TRUE}

train_files <- c(
 "ch1MPI2020-01-17_08-44-19_0000028.wav",
 "ch1MPI2020-01-23_10-30-35_0000106.wav",
 "ch1MPI_2020-01-30_15-06-38_0000193.wav",
 "ch1MPI2020-01-16_11-46-10_0000018.wav",
 "ch1MPI2020-01-16_13-40-58_0000019.wav",
 "ch1MPI2020-01-24_11-28-15_0000124.wav"
)
test_files <- setdiff(manual_ref$sound.files, train_files)

train_ref <- manual_ref[manual_ref$sound.files %in% train_files, ]
test_ref <- manual_ref[manual_ref$sound.files %in% test_files, ]
```

&nbsp;

### Get diverse set of templates
```{r cross-correlation detection, eval = FALSE, echo = TRUE}

# find templates
templates <- get_templates(train_ref, path = data_path,  bp = c(10, 50), fast = TRUE, ovlp = , wl = 200, n.templates = 4)

# create ext. selection table
templates_est <- selection_table(templates, extended = TRUE, confirm.extended = FALSE, path = data_path)

templates_est <- rename_est_waves(templates_est, new.sound.files = templates$distance.quantile)

```
&nbsp;

### Create spectrograms of templates

```{r Create catalog templates, eval = FALSE}

# create catalogs
catalog(X = templates_est, flim = c(10, 45), nrow = 2, ncol = 2, ovlp = 90, height = 10, width = 15, same.time.scale = TRUE, mar = 0.005, wl = 512, gr = FALSE, spec.mar = 0.4, lab.mar = 0.8, rm.axes = TRUE, by.row = TRUE, box = TRUE, pal = viridis, parallel = 10, collevels = seq(-100, 0, 5), img.prefix = "templates")

```

```{r catalog templates, out.width = "70%", echo = FALSE, fig.align= "center"}

knitr::include_graphics(normalizePath(file.path(data_path, "templates-Catalog_p1.jpeg")))

```

```{r, eval = FALSE, echo = FALSE}

saveRDS(templates_est, "./data/processed/templates_thyroptera_est.RDS")

```

&nbsp;

### Extract correlation vectors and optimize correlation threshold

```{r,  eval = FALSE, echo = TRUE}

corr_templ <- template_correlator(templates = templates_est,
    path = data_path, files = unique(train_ref$sound.files), parallel = paral, hop.size = 10, ovlp = 70)

corr_templ_train <- template_correlator(
  templates = templates_est,
    path = data_path, 
  files = unique(train_ref$sound.files), 
  hop.size = 10, 
  ovlp = 70
  )

opt_detec_train <- optimize_template_detector(
  reference = train_ref, 
  template.correlations = corr_templ_train,
  threshold = seq(0.05, 0.5, 0.01)
  )

```

```{r, eval = FALSE, echo = FALSE}
templates_est <- readRDS("./data/processed/templates_thyroptera_est.RDS")

paral <- 10

corr_templ_train <- template_correlator(
  templates = templates_est,
    path = data_path, 
  files = unique(train_ref$sound.files), 
  parallel = paral, 
  hop.size = 10, 
  ovlp = 70
  )

opt_detec_train <- optimize_template_detector(
  reference = train_ref, 
  template.correlations = corr_templ_train,
  threshold = seq(0.05, 0.5, 0.01), 
  parallel = paral
  )

opt_detec_train$templates <- gsub("-1", "", opt_detec_train$templates)

saveRDS(opt_detec_train, "./data/processed/optimization_results_3_templates_thyroptera.RDS")

```

&nbsp;

### Create recall-precision plot
```{r, eval = FALSE}

# summarize across sound files
diag_train <- summarize_diagnostic(opt_detec_train, time.diagnostics = FALSE)
agg_sd <- aggregate(cbind(recall, precision) ~ threshold + templates, data = opt_detec_train,
    mean)

# plot
ggplot(agg_sd, aes(x = recall, y = precision, group = templates, color = templates)) +
    geom_line() + geom_point() + scale_color_viridis_d(end = 1) + theme_classic()
```

```{r, eval = TRUE, echo = FALSE}

opt_detec_train <- readRDS("./data/processed/optimization_results_3_templates_thyroptera.RDS")

diag_train <- summarize_diagnostic(opt_detec_train, time.diagnostics = FALSE)

# names(diag_train)

agg_sd <- aggregate(cbind(recall, precision) ~ threshold + templates, data = diag_train,
    mean)

ggplot(agg_sd, aes(x = recall, y = precision, group = templates, color = templates)) +
    geom_line() + geom_point() + scale_color_viridis_d(end = 1) + theme_classic()

```

&nbsp;

### Check subset with highest performance (sorted by f1 score)

```{r, eval = FALSE}

sub_diag_train <-diag_train[order(diag_train$f1.score, decreasing = TRUE), ]

sub_diag_train <- sub_diag_train[1:10, c("templates","threshold", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f1.score")]

sub_diag_train
```

```{r, eval = TRUE, echo = FALSE}

# subset with highest performance
sub_diag_train <-diag_train[order(diag_train$f1.score, decreasing = TRUE), ]

sub_diag_train <- sub_diag_train[1:10, c("templates","threshold", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f1.score")]

# print dynamic table
oa_DT <- datatable(sub_diag_train, editable = list(target = "row"), rownames = FALSE, style = "bootstrap",
    filter = "top", options = list(pageLength = 100, autoWidth = TRUE, dom = "ft"),
    autoHideNavigation = TRUE, escape = FALSE)

formatRound(table = oa_DT, columns = sapply(sub_diag_train, is.numeric), 3)
```

&nbsp;

### Run detection on test files
```{r, detect on all data, eval = FALSE, echo = TRUE}

# get correlation vectors for test files
corr_templ_test <- template_correlator(
  templates = templates_est[templates_est$sound.file == "50%", ],
    path = data_path, files = unique(test_ref$sound.files), 
  parallel = paral, 
  hop.size = 10, 
  ovlp = 70
  )

# extract best template-threshold combination
opt_50 <- optimize_fourier_detec[optimize_fourier_detec$templates == "50%", ]
opt_thresh <- opt_50$threshold[which.max(opt_50$f1.score)]

# detect on test files
detec_test <- template_detector(
  template.correlations = corr_templ_test,
    threshold = opt_thresh
  )
```

&nbsp;

### Evaluate detection performance on test files
```{r, diagnose detect on all data, eval = FALSE, echo = TRUE}

diagnose_detection(reference = test_ref, detection = detec_test)

```

```{r, eval = FALSE, echo = FALSE}
saveRDS(detec_test, "./data/processed/detection_thyroptera_all.RDS")
```

```{r, eval = TRUE, echo = FALSE}

detec_test <- readRDS("./data/processed/detection_thyroptera_all.RDS")

diagnose_detection(reference = test_ref, detection = detec_test)
```

&nbsp;

<div class="alert alert-info">

### Takeaways

- Good performance on detecting Spix's disc-winged social calls: F1 score was 0.99 for the training data set and 0.94 for the testing data

</div>

&nbsp;

---

<font size="4">Session information</font>

```{r session info, echo=F}

sessionInfo()

```

### 2022-05-11 scripts/study_cases_bat_social_calls.Rmd
first commit

---
title: <center><font size="6"><b>Ohun</b></font></center>
subtitle: <center><font size="4"><b>Study cases</b></font></center>
author: <center><font size="4"><a href="http://marceloarayasalas.weebly.com/">Marcelo Araya-Salas</a> & <a href="http:///">Gloriana Chaverri</a></font></center>
date: <center>`r format(Sys.Date(), "%d-%m-%Y")`</center>
output:
  html_document:
    theme: default
    highlight: pygments
    code_folding: hide
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: yes
      smooth_scroll: yes
fontsize: 12pt 
editor_options: 
  chunk_output_type: console
---

## Load packages
```{r packages, message=FALSE, warning = FALSE, echo = TRUE, eval = TRUE, include = TRUE}

## add 'developer/' to packages to be installed from github
x <- c("pbapply", "viridis", "knitr", "remotes", "maRce10/warbleR", "maRce10/ohun", "DT", "ranger")

aa <- lapply(x, function(y) {

  # get pakage name
  pkg <- strsplit(y, "/")[[1]]
  pkg <- pkg[length(pkg)]

  # check if installed, if not then install 
  if (!pkg %in% installed.packages()[,"Package"])  {

      if (grepl("/", y))  remotes::install_github(y, force = TRUE) else
    install.packages(y) 
    }

  # load package
  try(require(pkg, character.only = T), silent = T)
})

```

```{r set knitting options,  eval = TRUE, echo = FALSE}

opts_knit$set(dpi = 100, fig.width = 12, warning = FALSE, root.dir = "..", fig.align = 'center', message = FALSE)

```

## Set functions and global parameters
```{r functions, eval = TRUE, echo = TRUE}

cols <- viridis(10, alpha = 0.6)

warbleR_options(wav.path = "./data/raw")

```

# Bat calls data

## File descriptions

manual_detections_Raven_bats4testing.csv: Manual detection of all visible bat vocalizations (Raven)

manual_detections_bats4testing.csv: Manual detection of all high quality bat vocalizations (Avisoft)

automated_detections_bats4testing.csv: Automated detection with Eric's function

# Read data
```{r}

reference_all <- read.csv("./data/raw/manual_detections_Raven_bats4testing.csv")

reference_best <- read.csv("./data/raw/manual_detections_bats4testing.csv")

ericks_detection <- read.csv("./data/raw/automated_detections_bats4testing.csv")

```

# Check references

Exhaustive detection (all signals) 
```{r}
# all
check_sels(reference_all, pb = FALSE)
feature_reference(reference_all, path = "./data/raw")
```

High quality subset
```{r}

check_sels(reference_best, pb = FALSE)
feature_reference(reference_best, path = "./data/raw")
```

Detection from Erick's algorithm
```{r}

check_sels(ericks_detection, pb = FALSE)
feature_reference(ericks_detection, path = "./data/raw")

```

# Compare references (high quality vs exhaustive)

```{r}

# all vs best
diagnose_detection(reference = reference_all, detection = reference_best)

```
&nbsp;

<div class="alert alert-info">

- *High quality reference has ~70% of all signals*

</div>

&nbsp;

# Diagnose Erick's detection

Compare to exhaustive detection:

```{r}

diagnose_detection(reference = reference_all, detection = ericks_detection)

```
&nbsp;

<div class="alert alert-info">

- *Erick's algorithm detected ~50% of all signals*
- *44% of the detected signals were noise (no bat calls)*

</div>

&nbsp;

Compare to high quality subset:

```{r}

diagnose_detection(reference = reference_best, detection = ericks_detection)

```
&nbsp;

<div class="alert alert-info">

- *Erick's algorithm detected ~70% of high quality signals*
- *43% of the detected signals were noise (no bat calls)*

</div>

&nbsp;

# Run ohun energy detector

```{r, eval = FALSE}

ohun_opt_detec <- optimize_energy_detector(reference = reference_best, threshold = c(0.5, 1:4), smooth = 1:4, hold.time = 1:8, min.duration = c(0.05, 0.1, 0.2, 0.3, 0.4, 1, 2), max.duration = c(30, 40, 50), parallel = 10, path = "./data/raw", bp = c(5, 100))

saveRDS(ohun_opt_detec, "./data/processed/high_quality_reference_optimization.RDS")

ohun_opt_detec_all <- optimize_energy_detector(reference = reference_all, threshold = c(0.5, 1:4), smooth = 1:4, hold.time = 1:8, min.duration = c(0.05, 0.1, 0.2, 0.3, 0.4, 1, 2), max.duration = c(30, 40, 50), parallel = 10, path = "./data/raw", bp = c(5, 100))

saveRDS(ohun_opt_detec_all, "./data/processed/all_calls_reference_optimization.RDS")

ohun_opt_detec_ericks <- optimize_energy_detector(reference = ericks_detection, threshold = c(0.5, 1:4), smooth = 1:4, hold.time = 1:8, min.duration = c(0.05, 0.1, 0.2, 0.3, 0.4, 1, 2), max.duration = c(30, 40, 50), parallel = 10, path = "./data/raw", bp = c(5, 100))

saveRDS(ohun_opt_detec_ericks, "./data/processed/ericks_reference_optimization.RDS")

```

# ohun detection performance

Compare to exhaustive reference (only the highest sensitivity iteration is shown)
```{r}

ohun_opt_detec_all <- readRDS("./data/processed/all_calls_reference_optimization.RDS")

ohun_opt_detec_all <- ohun_opt_detec_all[which.max(ohun_opt_detec_all$sensitivity), c(1:11, 17:18)]

ohun_opt_detec_all
# oa_DT <- datatable(ohun_opt_detec_all, editable = list(
#   target = 'row'
# ), rownames = FALSE, style = "bootstrap",  filter = 'top', options = list(
#   pageLength = 100, autoWidth = TRUE, dom = 'ft'
# ), autoHideNavigation = TRUE, escape = FALSE)
# 
# formatRound(table = oa_DT, columns = sapply(ohun_opt_detec_all, is.numeric), 3)

```
&nbsp;

<div class="alert alert-info">

- *Optimized ohun detection got ~70% of all signals*
- *24% of the detected signals were noise (no bat calls)*

</div>

&nbsp;

Compare to high quality reference (only the highest sensitivity iteration is shown)
```{r}

ohun_opt_detec <- readRDS("./data/processed/high_quality_reference_optimization.RDS")

ohun_opt_detec <- ohun_opt_detec[which.max(ohun_opt_detec$sensitivity), c(1:11, 17:18)]

ohun_opt_detec
# oa_DT <- datatable(ohun_opt_detec, editable = list(
#   target = 'row'
# ), rownames = FALSE, style = "bootstrap",  filter = 'top', options = list(
#   pageLength = 100, autoWidth = TRUE, dom = 'ft'
# ), autoHideNavigation = TRUE, escape = FALSE)
# 
# formatRound(table = oa_DT, columns = sapply(ohun_opt_detec, is.numeric), 3)

```
&nbsp;

<div class="alert alert-info">

- *Optimized ohun detection got ~74% of high quality signals*
- *32% of the detected signals were noise (no bat calls)*

</div>

&nbsp;

Compare to Erick's detection (only the highest sensitivity iteration is shown)
```{r}

ohun_opt_detec_ericks <- readRDS("./data/processed/ericks_reference_optimization.RDS")

ohun_opt_detec_ericks <- ohun_opt_detec_ericks[which.max(ohun_opt_detec_ericks$sensitivity), c(1:11, 17:18)]

ohun_opt_detec_ericks
# oa_DT <- datatable(ohun_opt_detec_ericks, editable = list(
#   target = 'row'
# ), rownames = FALSE, style = "bootstrap",  filter = 'top', options = list(
#   pageLength = 100, autoWidth = TRUE, dom = 'ft'
# ), autoHideNavigation = TRUE, escape = FALSE)
# 
# formatRound(table = oa_DT, columns = sapply(ohun_opt_detec_ericks, is.numeric), 3)

```
&nbsp;

<div class="alert alert-info">

- *Optimized ohun detection got ~71% of signals detected by Erick's algorithm*
- *31% of the detected signals were noise (no bat calls)*

</div>

&nbsp;

# Filter using Random Forest

Before random forest filtering
```{r, eval = TRUE}
# best parameters from all signal optimization
ohun_detec <- energy_detector(threshold = 1, smooth = 1, hold.time = 1, min.duration = 0.1, max.duration = 50, parallel = 10, path = "./data/raw", bp = c(5, 100), pb = FALSE)

diagnose_detection(reference_all, ohun_detec)
```

After random forest filtering
```{r, eval = TRUE}
ohun_detec <- label_detection(reference = reference_all, detection = ohun_detec, pb = FALSE)

# measure spectrographic parameters
spectral_parameters <- spectro_analysis(ohun_detec, bp = c(5, 100), fast = TRUE, ovlp = 70, parallel = 10, pb = FALSE)

spectral_parameters$SNR <- sig2noise(ohun_detec, mar = 0.01, pb = FALSE)$SNR

spectral_parameters$class <- ohun_detec$detection.class

spectral_parameters$class[spectral_parameters$class != "false.positive"] <- "true.positive"

# make it a factor for ranger to work 
spectral_parameters$class <- as.factor(spectral_parameters$class)

# run RF model on spectral parameters
  rfm <-
    ranger(
      class ~ .,
      data = spectral_parameters[, !names(spectral_parameters) %in% c("sound.files", "selec")],
      num.trees = 10000,
      importance = "impurity",
      seed = 10
    )

  # table(lab_detec$detection.class)
ohun_detec$pred.class <- rfm$predictions

rf_ohun_detec <- ohun_detec[ohun_detec$pred.class == "true.positive", ]

diagnose_detection(reference = reference_all, detection = rf_ohun_detec, pb = FALSE)

```

&nbsp;

<div class="alert alert-info">

- *Random forest filtered ohun detection got ~68% of all signals*
- *~4% of the detected signals were noise (no bat calls)*

</div>

&nbsp;

---

<font size="4">R session information</font>

```{r session info, echo=F}

sessionInfo()

```

### 2022-05-12 manuscript/manuscript.Rmd
edit manuscript

@@
 } else writeLines(text = "empty", "combined_bibs.bib")
+
+library(knitr)
+
+```
+
+```{r, echo = FALSE}
+
+# knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
+
+opts_knit$set(root.dir = "..")
+
+# opts_chunk$set(root.dir = "..")
 
@@
 Bats were captured at their roosting sites (furled leaves of Zingeberaceae plants).
-Each individual bat was released in a large flight cage (9 x 4 x 3 m) for a 5 min period and their ultrasonic inquiry calls were recorded using a condenser microphone (CM16, Avisoft Bioacoustics, Glienike/
-Nordbahn, Germany) through an Avisoft UltraSoundGate 116Hm connected to a laptop computer running Avisoft-Recorder software (sampling rate of 500 kHz, 16-bit amplitude resolution).
+Each individual bat was released in a large flight cage (9 x 4 x 3 m) for a 5 min period and their ultrasonic inquiry calls were recorded using a condenser microphone (CM16, Avisoft Bioacoustics, Glienike/Nordbahn, Germany) through an Avisoft UltraSoundGate 116Hm connected to a laptop computer running Avisoft-Recorder software, with a sampling rate of 500 kHz and a 16-bit amplitude resolution.
 
@@
 Inquiry calls of Spix's disc-winged bats are stereotyped signals.
-Most variation is found between individuals although the basic structure of a short, downward broadband frequency modulation (FIG XXX) is always shared.
+Most variation is found between individuals although the basic structure of a short, downward broadband frequency modulation (Fig.
+bat-spectro) is always shared.
 Template-based detection is an useful approach when there is little structural differences in the target signals.
-Therefore we used the function
+Therefore we used the function.
+
+```{r catalog, out.width = "100%", echo = FALSE, fig.align= "center"}
+
+# knitr::include_graphics("./data/raw/thyroptera/Catalog_p1.jpeg")
+
+knitr::include_graphics(normalizePath(file.path("./data/raw/thyroptera", "Catalog_p1.jpeg")))
+
+list.files("./data/raw/thyroptera")
+print(getwd())
+```
+
+Fig. bat-spectro.
 
@@
 
-We also used recordings from 18 zebra finch males. These recordings were obtained ....
+We used recordings from 18 zebra finch males. These recordings were obtained ....
 

### 2022-07-13 manuscript/manuscript (neuro-ubuntu's conflicted copy 2022-07-13).Rmd
first commit

---
title: 'ohun: an R package for optimizing automatic acoustic signal detection'
author:
- Marcelo Araya-Salas ^1,^ ^2,^ ^3^ *
- Grace Smith-Vidaurre ^4^
- Gloriana Chaverri ^3^
- Fabiola Chirino ^1^
- Alejandro Rico-Guevara ^5^
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: no
  word_document:
    toc: yes
    toc_depth: '3'
  pdf_document:
    toc: yes
    toc_depth: '3'
always_allow_html: yes
bibliography: combined_bibs.bib
citation-style: apa.csl
link-citations: yes
editor_options:
  chunk_output_type: console
---

```{r load packages, echo = FALSE, message = FALSE, warning=FALSE}

# github packages must include user name ("user/package")
# knitr is require for creating html/pdf/word reports
# kableExtra is used to print pretty formatted tables 
# formatR is used for soft-wrapping code
# klippy is used for adding a copy button to each code block

library(knitr)
library(ohun)
library(ggplot2)
library(viridis)
library(kableExtra)
library(rmdwc)

pkgs <- c("kableExtra", "knitr", "rprojroot", "sciRmdTheme", "rmdwc", "whereami", "maRce10/ohun", "ggplot2", "viridis", "rmdwc")

# install/ load packages
out <- lapply(pkgs, function(y) {

  # get pakage name
  pkg <- strsplit(y, "/")[[1]]
  pkg <- pkg[length(pkg)]

  # check if installed, if not then install 
  if (!pkg %in% installed.packages()[,"Package"])  {

      if (grepl("/", y))  remotes::install_github(y, force = TRUE) else
    install.packages(y) 
    }

  # load package
  a <- try(require(pkg, character.only = T), silent = T)

  if (!a) remove.packages(pkg)
  })

```

```{r theme setting, echo=FALSE}

# simplest default settings
sciRmdTheme::set.theme()

```

```{r custom data frame printing, echo = FALSE, message=FALSE}

# this is a customized printing style data frames 
# screws up tibble function
tibble <- function(x, ...) { 
  x <- kbl(x, digits=4, align= 'c', row.names = FALSE) 
   x <- kable_styling(x, position ="center", full_width = FALSE,  bootstrap_options = c("striped", "hover", "condensed", "responsive")) 
   asis_output(x)
}

registerS3method("knit_print", "data.frame", tibble)

```

```{r set working directory, echo=FALSE}

# set working directory as project directory or one directory above,
rootdir <- try(rprojroot::find_rstudio_root_file(), silent = TRUE)
if (is(rootdir, "try-error")) rootdir <-  ".."
opts_knit$set(root.dir = rootdir)

```

```{r, fixing citation .bib files, eval = TRUE, echo = FALSE}

# how to cite in Rmarkdown
# https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html

# Here include the path to the libraries for each author. This code will copy/update the library each time the Rmarkdown file is compiled
# Alternatively authors ca
personal_libraries <- c(author_1 = "/home/m/Documents/mendeley_libraries/ohun_paper.bib")

# update bibtex library
for (i in 1:length(personal_libraries))
if (file.exists(personal_libraries[i]))
 file.copy(from = personal_libraries[i], to = file.path(dirname(whereami::thisfile()), paste0(names(personal_libraries)[i], ".bib")), overwrite = TRUE)

## combine .bib files in compiled.bib ###
# list bib files
bibs <- list.files(pattern = ".bib$", path = dirname(whereami::thisfile()), full.names = TRUE)

# remove combined from pool
bibs <- bibs[basename(bibs) != "combined_bibs.bib"]

#  read in a list
if (length(bibs) > 0){
  combined_bibs_l <- lapply(bibs, readLines)
  names(combined_bibs_l) <- gsub("\\.bib$", "", basename(bibs))

# combine bibs in a single one
combined_bibs <- unlist(combined_bibs_l)

# save pool libraries as combined_bibs.bib
writeLines(text = combined_bibs, file.path(dirname(whereami::thisfile()), "combined_bibs.bib"))

# check if there are duplicated keys
keys <- grep("@article{", combined_bibs, fixed = TRUE, value = TRUE)
keys <- gsub("@article{", "", keys, fixed = TRUE)
keys <- gsub(",", "", keys, fixed = TRUE)

tab_keys <- table(keys)

if (anyDuplicated(keys)){
  print(paste0(sum(tab_keys > 1), " duplicate(s) references found in combined_bibs.bib"))
}
} else writeLines(text = "empty", file.path(dirname(whereami::thisfile()), "combined_bibs.bib"))

```

&nbsp;

^1^ Centro de Investigación en Neurociencias, Universidad de Costa Rica, San José, Costa Rica

^2^ Escuela de Biología, Universidad de Costa Rica, San José, Costa Rica

^3^ Sede del Sur, Universidad de Costa Rica, Golfito, Costa Rica

^4^ GRACE'S AFFILIATION

^5^  Department of Biology, University of Washington, Seattle, USA

\* *To whom correspondence should be addressed*

### Keywords:

---

&nbsp;

```{r word count, eval = TRUE, echo = FALSE, results='asis'}

# word count without code chunks
word_count <- rmdcount(whereami::thisfile())

# word count with code chunks
word_count_chunk <- rmdcount(whereami::thisfile(), exclude='')

# print results
cat(paste0("Word count: ", word_count[3], " (including code chunks: ", word_count_chunk[3], ")"))

```

Check [this link](https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html) for details on how to add citations/bibliography in Rmarkdown

---

## Abstract

Animal acoustic signals are widely used in a diversity of research areas. This partly due to the relative ease for registering them for a wide range of taxonomic groups and research settings. However, bioacoustic research can easily generate large amounts of data which might proof challenging to analyze in a timely manner. Many tools for the automatic detection of signals are currently available. Nonetheless, choosing the right approach for a specific task might turn difficult as there are no widely available tools that can provide a common framework for evaluating detection performance. Here we present the new R package ohun. The package is intended to facilitate the automatic detection of acoustic signals, by providing functions to diagnose and optimize detection routines. The package makes use of reference annotations containing the time position of target signals in a training data set to evaluate the performance of detection routines, using common signal detection theory indices. This can be done with routine outputs imported from other software and from detection run within the package itself. In includes an implementation of two automatic detection methods commonly used in bioacustic analysis: energy-based detection and template-based detection. We provide examples on how to evalute detections from external software and showcase the package usage with study cases on zebra-finch (*Taenopygia gutata*) songs and Spix's disc-winged bat (*Thyroptera tricolor*) ultrasonic social calls.

## Introduction

Animal acoustic signals are traits widely used to investigate a variety of questions in highly diverse areas, ranging from neurobiology to community ecology and evolutionary biology. The wide usage of animal sounds in research is partly due to the fact that they can be easily registered using non-intrusive methods, can be obtained in a variety of settings from laboratories to natural areas and the equipment required for registering an analyzing these signals is increasingly inexpensive. In addition, the existence of online repositories and growing number has allowed facilitated the research of animal acoustic signals at larger taxonomic and geographic scales. As a result, a growing variety of a computational tools for the analysis of acoustic features on those signals is increasingly available, reflecting the growing relevance of bioacustics in the scientific tool kit.

<!-- (many of those in the form of freely available software written on programming environments familiar to the scientific community (seewave, warbleR, scikit-mad)) -->

Bioacoustic research can easily generate large amounts of data, which sometimes may proof challenging to analyze in a timely manner. Annotations that require precise location of signals in frequency and time, most commonly used in behavioral and phenotypic evolution research, remain among the most time consuming phases of analysis for this type of data. Fortunately, several tools have been develop to assist researchers in this matter. Indeed, new methods for automatic annotation of acoustic signals are continuously developed (reviewd by @Stowell2022). This growing availability of tools, particularly as free software, is expected to further simplify acoustic data processing, making it accessible to a wider range of users and scientific questions. However, this diversity of tools posits a challange as it can also be hard to navigate. given the little use of widespread metrics that can evaluate the behavior of tools under different scenarios.

Here we present the new R package ohun. The package is intended to facilitate the automatic detection of acoustic signals, providing functions to diagnose and optimize detection routines. The package makes use of reference annotations containing the time position of target signals in a training data set to evaluate the performance of detection routines. This can be done with routine outputs imported from other software as well as detection run within the package itself. The package also offers an implementation of two automatic detection methods commonly used in bioacustic analysis: energy-based detection and template-based detection. We first explain how acoustic signal detection routines can be evaluate and then showcase the package usage with study cases on zebra-finch songs (*Taenopygia gutata*) and Spix's disc-winged bats (*Thyroptera tricolor*) which represent different recording settings (lab and flight cages) and signal types (sonic mating signals and ultrasonic social calls).

## Diagnosing detection performance

The package makes use of signal detection theory indices to evaluate detection performance. Signal detection theory deals with the process of recovering signals (i.e. target signals) from background noise (not necessarily acoustic noise) and it’s widely used for optimizing this decision making process in the presence of uncertainty. During a detection routine, the detected ‘items’ can be classified into 4 classes: true positives (TPs, target signals correctly identified as signal), false positives (FPs, noise incorrectly identified as ‘signal’), false negatives (FNs, signals incorrectly identified as noise) and true negatives (TNs, background noise correctly identified as noise). However, TNs cannot always be easily defined in the context of acoustic signal detection as noise cannot always be partitioned in discrete units. Hence, the package makes use of TPs, FPs and FNs to calculate three additional indices that can further assist with evaluating the performance of a detection routine.

 - Recall: correct detections relative to total detections (a.k.a. true positive rate or sensitivity; TPs / (TPs + FNs))
- Precision: correct detections relative to total detections (TPs / (TPs + FPs)).
- F1 score: combines recall and precision as the harmonic mean of these two, providing a single value for evaluating performance (a.k.a. F-measure or Dice similarity coefficient).

A perfect detection will have no false positives or false negatives, which will result in both recall and precision equal to 1. However, perfect detection cannot always be reached and some compromise between detecting all target signals plus some noise (recall = 1 & precision < 1) and detecting only target signals but not all of them (recall < 1 & precision = 1) might be warranted. The right balance between these two extremes will be given by the relative costs of missing signals and mistaking noise for signals, given the specific goals of the study. These indices provide an useful framework for diagnosing and optimizing the performance of a detection routine.

The package offers tools to evaluate the performance of an acoustic signal detection based on the indices described above. To accomplish this, annotations derived from a detection routine are compared against a reference table containing the time position of all target signals in the sound files. For instance, the following code evaluates a routine run in Raven Pro 1.6 (XXXX) using the "band limited energy detector" option (minimum frequency: 0.8 kHz; maximum frequency: 22 kHz; minimum duration: 0.54989 s; maximum duration: 0.54989s; minimum separation: 0.02268 s) on a subset of the Zebra finch recordings described below:

```{r, eval = TRUE, echo = FALSE}

library(Rraven)

# read data
manual_ref_thy <- read.csv(file.path(rootdir, "//data/raw/thyroptera/", "manual_annotations_thyroptera.csv"))

manual_ref_tae <- read.csv(file.path(rootdir, "/data/processed/taeniopygia", "manual_selections_Taeniopygia.csv"))

raven_detec <- imp_raven(path = file.path(rootdir, "/data/raw/"), files = "Raven.Band.Limited.Energy.Detector.selections.txt", warbler.format = TRUE)

# raven_ref <- manual_ref[manual_ref$sound.files %in% raven_detec$sound.files, ]

```

```{r, eval = TRUE, echo = TRUE}

diagnose_detection(reference = manual_ref_tae, detection = raven_detec)

```

The output shows the indices described above. The function also allows to detail those indices by sound file. Here we show the first 6 files:

```{r, eval = TRUE, echo = TRUE}

diag_raven <- diagnose_detection(reference = manual_ref_tae, detection = raven_detec, by.sound.file = TRUE)

head(diag_raven)
```

SOmetime

Diagnostics from routines using different tuning parameters can be useful for identifying the parameter values that optimize a detection. The process of evaluating different routines for detection optimization is incorporated into the two signal detection approaches provided natively by ohun, which we depict in the following section. Note that the detection with Raven Pro does not necessarily reflect the best performance of this software and has been included only as an example on evaluating detection from external sources rather than a direct comparison of performance to ohun.

In addition, several functions can facilitate getting acoustic data sets in the right format prior to automatic detection. Having some sense of the format and size of the sound files to be analyzed is an important step for avoiding downstream errors and adjusting our time performance expectations. The function `feature_acoustic_data` prints a summary of the duration, size and format of all the recordings in a folder:

```{r, eval = FALSE, echo = TRUE}

feature_acoustic_data(path = "PATH_TO_FILES")

```

```{r, eval = TRUE, echo = FALSE}

feature_acoustic_data(path = file.path(rootdir, "data/raw/taeniopygia/"))

```

## Signal detection with ohun

The package offers two methods for automatic signal detection: template-based and energy-based detection. These methods are better suited for highly stereotyped or good signal-to-noise ratio (SNR) signals, respectively. If the target signals do not fit these requirements, more elaborated methods (i.e. machine learning approaches) are warranted.

## Study cases

### Template detection on ultrasonic social calls of  Spix's disc-winged bats

We recorded `r length(unique(manual_ref_thy$sound.files))` individuals of Spix's disc-winged bats at Baru Biological Station, in south-western Costa Rica in January 2020. Bats were captured at their roosting sites (furled leaves of Zingeberaceae plants). Each individual bat was released in a large flight cage (9 x 4 x 3 m) for a 5 min period and their ultrasonic inquiry calls were recorded using a condenser microphone (CM16, Avisoft Bioacoustics, Glienike/Nordbahn, Germany) through an Avisoft UltraSoundGate 116Hm plugged into a laptop computer running Avisoft-Recorder software. Recordings were made at a sampling rate of 500 kHz and an amplitude resolution of 16 bits.

Recordings were manually annotated using Raven Pro 1.6 (XXXX). Annotations were created by visual inspection of spectrograms, in which the start and end of signals were determined by the location of the continuous traces of power spectral entropy of the target signals. A total of `r nrow(manual_ref_thy)` calls were annotated (~`r round(nrow(manual_ref_thy) / 30, 0)` calls per recording). Annotations were made with a time window of 200 samples and 70% of overlap. Annotations were then imported into R using the package Rraven (XXXX).

Inquiry calls of Spix's disc-winged bats are structurally stereotyped. Most variation is found between individuals although the basic form of a short, downward broadband frequency modulation is always shared (Fig. BAT-SPECTRO, Araya-Salas et al 2021 ontogeny).

```{r catalog, out.width = "100%", echo = FALSE, fig.align= "left",fig.cap="Fig. BAT-SPECTRO. Example spectrograms of Spix's disc-winged social calls for each of the 30 recordings used in the analysis. The highest signal-to-noise ratio call by sound file are shown. The time scale range is 0.071 s and the frequency range 10-44 kHz", error=TRUE}

# knitr::include_graphics(normalizePath(file.path("./data/raw/thyroptera", "Catalog_p1.jpeg")))
knitr::include_graphics("~/Dropbox/Projects/ohun_paper/data/raw/thyroptera/Catalog_p1.jpeg")

```

Template-based detection is a useful approach when there is little structural differences in the target signals. We used this approach in ohun to detect inquiry calls. To do this, we tested the performance of three acoustic templates on a training subset of five sound files. The function `get_templates` finds several signals representative of the variation in signal structure. This function measures several spectrographic parameters which are then summarized using Principal Component Analysis. The first two components are used to project the acoustic space. On this space the function defines sub-spaces as equal-size slices of a circle centered at the centroid of the acoustic space. Templates are then selected as those closer to the centroid within each of the sub-spaces, including the centroid for the entire acoustic space. The user needs to define the number of sub-spaces in which the acoustic space will be split.

```{r, eval = FALSE}

# read manual annotations
manual_ref_thy <- read.csv("manual_annotations_thyroptera.csv")

# get random subset of 5 sound files for training
set.seed(1)
train_files <- sample(unique(manual_ref_thy$sound.files), size = 5)
train_ref <- manual_ref_thy[manual_ref_thy$sound.files %in% train_files, ]

# the rest for testin g
test_files <- setdiff(manual_ref_thy$sound.files, train_files)
test_ref <- manual_ref_thy[manual_ref_thy$sound.files %in% test_files, ]

# find templates
templates <- get_templates(train_ref, path = data_path,  bp = c(10, 50), ovlp = 70, wl = 200, n.sub.spaces = 3)

```

```{r, echo = FALSE, fig.height= 6, fig.cap= "FIG. ACOUSTIC-SPACE. Acoustic space defined as the the first two component of a Principal Component Analysis on spectrographic parameters. Templates are selected as those closer to the centroid for each of the sub-spaces. Gray dashed lines delimit the region of sub-spaces. Yellow circles around points highlight the position of the signals selected as templates."}

# read manual annotations
manual_ref_thy <- read.csv(file.path(rootdir, "/data/raw/thyroptera/", "manual_annotations_thyroptera.csv"))

# get random subset of 5 sound files for training
set.seed(1)
train_files <- sample(unique(manual_ref_thy$sound.files), size = 5)
train_ref <- manual_ref_thy[manual_ref_thy$sound.files %in% train_files, ]

# the rest for testing
test_files <- setdiff(manual_ref_thy$sound.files, train_files)
test_ref <- manual_ref_thy[manual_ref_thy$sound.files %in% test_files, ]

source("~/Dropbox/R_package_testing/ohun/R/get_templates.R")

templates <- get_templates(train_ref, path = file.path(rootdir, "/data/raw/thyroptera/"),  bp = c(10, 50), ovlp = 70, wl = 200, n.sub.spaces = 3)

```

```{r, eval = TRUE, echo = FALSE}

data_path <- file.path(rootdir, "/data/raw/thyroptera/")

manual_ref_thy <- read.csv(file.path(data_path, "manual_annotations_thyroptera.csv"))

# get random subset of 5 sound files for training
set.seed(1)
train_files <- sample(unique(manual_ref_thy$sound.files), size = 5)
train_ref <- manual_ref_thy[manual_ref_thy$sound.files %in% train_files, ]

# the rest for testing
test_files <- setdiff(manual_ref_thy$sound.files, train_files)
test_ref <- manual_ref_thy[manual_ref_thy$sound.files %in% test_files, ]
```

The output of the `get_templates` includes an acoustic space plot (FIG. ACOUSTIC-SPACE) in which the position of the signals selected as templates is highlighted. In the following code we used those templates for detecting calls. The code iterates a template-based detection on the training data set across a range of correlation thresholds, in order to find the combination of threshold and template with the best performance.

```{r, eval = FALSE}

# get correlation vectors
corr_templ_train <- template_correlator(
  templates = templates,
    path = data_path, 
  files = unique(train_ref$sound.files), 
  hop.size = 10, 
  ovlp = 70
  )

# evaluate detection for different correlation thresholds
opt_detec_train <- optimize_template_detector(
  reference = train_ref, 
  template.correlations = corr_templ_train,
  threshold = seq(0.05, 0.5, 0.01)
  )

```

Note that the correlation vectors are estimated first (i.e. vectors of correlation values across sound files, `template_correlator()`) and then the correlation thresholds are optimized on them (`optimize_template_detector()`). The output of `optimize_template_detector()` contains the detection performance indices for each combination of templates and thresholds. Table TEMPLATE PERFORMANCE shows the two highest performance runs for each template.

```{r, eval = TRUE, echo = FALSE}

opt_detec_train <- readRDS(file.path(rootdir, "/data//processed/optimization_results_3_templates_thyroptera.RDS"))

# opt_detec_train <- opt_detec_train[order(opt_detec_train$templates, opt_detec_train$f1.score, decreasing = TRUE), ]

# sub_opt_detec_train <- opt_detec_train[ave(-opt_detec_train$f1.score, opt_detec_train$templates, FUN = rank) <= 4, ]

sub_odt <- lapply(unique(opt_detec_train$templates), function(x) {

    X <- opt_detec_train[opt_detec_train$templates == x, ]
    X <- X[order(X$f1.score, decreasing = TRUE), ]
    return(X[1:2, ])
    })

sub_opt_detec_train <- do.call(rbind, sub_odt)

kbl <- kable(sub_opt_detec_train[, c("threshold", "templates", "true.positives", "false.positives", "false.negatives", "recall", "precision", "f1.score")], row.names = FALSE, caption = "TABLE TEMPLATE PERFORMANCE. Performance diagnostic of template-based detections using four templates across several threshold values. Only the two highest performance iterations for each template are shown.")

kable_styling(kbl, bootstrap_options = c("striped", "hover",  full_width = FALSE))

```

We can explore the performance of each template in more detail by looking at the change in F1 score across thresholds (FIG. THRESHOLD vs F1.SCORE).

```{r, eval = TRUE, echo = FALSE, fig.height= 3, fig.align='center', fig.cap= "FIG. THRESHOLD vs F1.SCORE. Shows the changes in F1 score across the range of threshold values"}

agg_f1 <- aggregate(f1.score ~ threshold + templates, data = opt_detec_train,
mean)

ggplot(agg_f1, aes(x = threshold, y = f1.score, group = templates, color = templates)) +
    geom_line() + 
    geom_point() + 
    scale_color_viridis_d(end = 1, labels = c("Centroid", "1st", "2nd", "3rd"), alpha = 0.7) + 
    labs(x= "Cross-correlation threshold", y = "F1 score", color = "Templates")

```

In this example the "centroid" template, produced the best performance (TABLE TEMPLATE PERFORMANCE; FIG. THRESHOLD vs F1.SCORE). Hence, we will use this template for detecting calls on the rest of the data. The following code extracts this template from the reference annotation table and use it to find inquiry calls on the testing data set:

```{r, detect on all data 67, eval = FALSE, echo = TRUE}

# get correlation vectors for test files
corr_templ_test <- template_correlator(
  templates = templates[templates$sound.file == "centroid", ],
    path = data_path, files = unique(test_ref$sound.files), 
  hop.size = 10, 
  ovlp = 70
  )

# detect on test files
detec_test <- template_detector(
  template.correlations = corr_templ_test,
    threshold = 0.45
  )

diagnose_detection(reference = test_ref, detection = detec_test)

```

```{r, eval = TRUE, echo = FALSE}

detec_test <- readRDS(file.path(rootdir, "/data/processed/detection_thyroptera_all_67%_template.RDS"))

diagnose_detection(reference = test_ref, detection = detec_test)
```

The last line of codes evalutes the detection on the test data set, which shows a good performance for both recall and precision.

### Energy-based detection on zebra finch songs

We used recordings from 18 zebra finch males. These recordings were obtained ....

Zebra finch vocalizations are composed by multiple elements (i.e. distinct patterns of continuous traces of power spectral entropy in the spectrogram separated by time gaps) that can vary importantly in key features as duration and frequency range (ZEBRAFINCH-SPECTRO). However, as the recorded signals show a good signal-to-noise ratio, they can potentially be detected using an energy-based approach.  
```{r catalog zebra f, out.width = "100%", echo = FALSE, fig.align= "left",fig.cap="Fig. ZEBRAFINCH-SPECTRO. Example spectrograms of male zebra finch songs for each of the 18 sound files used in the analysis. The highest signal-to-noise ratio call by sound file are shown. The time scale range is 0.359 s and the frequency range 0-11 kHz. Signals have been highlighted for visualization purposes only.", error=TRUE}

knitr::include_graphics(normalizePath(file.path(rootdir, "data/raw/taeniopygia/same.time-Catalog_p1.jpeg")))

```

The following code loads the reference annotations and split them in two data set for training (3 sound files) and testing (15 sound files):
```{r, eval = TRUE, echo = FALSE}

manual_ref_tae <- read.csv(file.path(rootdir, "/data/processed/taeniopygia", "manual_selections_Taeniopygia.csv"))

set.seed(450)    
train_files <- sample(unique(manual_ref_tae$sound.files), 3)

test_files <- setdiff(manual_ref_tae$sound.files, train_files)

train_ref <- manual_ref_tae[manual_ref_tae$sound.files %in% train_files, ]
test_ref <- manual_ref_tae[manual_ref_tae$sound.files %in% test_files, ]

```

```{r, eval = FALSE, echo = TRUE}

manual_ref_tae <- read.csv("manual_selections_Taeniopygia.csv")

set.seed(450)    
train_files <- sample(unique(manual_ref_tae$sound.files), 3)    
test_files <- setdiff(manual_ref_tae$sound.files, train_files)

train_ref <- manual_ref_tae[manual_ref_tae$sound.files %in% train_files, ]
test_ref <- manual_ref_tae[manual_ref_tae$sound.files %in% test_files, ]

```

Now we can optimize the detection parameters using the function `optimize_energy_detector`. This function runs a detection for all possible combinations of tuning parameters. The code below tries three minimum duration and maximum duration values and two hold time values (which merges signals within the specified time interval):

```{r}
feature_reference(test_ref)
```

```{r  optimize automated detection, eval = FALSE, echo = TRUE}

opt_det_train <- optimize_energy_detector(
  reference = train_ref, 
  files = train_files, 
  threshold = c(1, 5), 
  hop.size = 11.6, 
  smooth = c(5, 10), 
  hold.time = c(0, 5), 
  min.duration = c(5, 15, 25), 
  max.duration = c(275, 300, 325), 
  bp = c(0.5, 10)
)

```

The output (`opt_det_train`) shows the performance indices for each of those combinations, here we show the 10 combinations with the highest F1 score:

```{r, eval = FALSE, echo = TRUE}

# subset with highest performance
opt_det_train <- opt_det_train[order(opt_det_train$f1.score, decreasing = TRUE), ]

head(opt_det_train, 10)
```

```{r, eval = TRUE, echo = FALSE}

opt_det_train <- read.csv(file.path(rootdir, "/data/processed/taeniopygia", "detection_optimization.csv"))

# subset with highest performance
sub_opt_det_train <- opt_det_train[order(opt_det_train$f1.score, decreasing = TRUE), ]

sub_opt_det_train <- sub_opt_det_train[1:10, c("threshold", "smooth", "hold.time", "min.duration", "max.duration", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f1.score")]

sub_opt_det_train
```

We now can use the tuning parameter values that produce the best performance to detect signals on the test data set:

```{r, select best params, eval = FALSE, echo = TRUE}

best_param <- opt_det_train[which.max(opt_det_train$f1.score), ]

det_test <- energy_detector(
  files = test_files, 
  threshold = best_param$threshold, 
  hop.size = 11.6, 
  smooth = best_param$smooth, 
  hold.time = best_param$hold.time, 
  min.duration = best_param$min.duration, 
  max.duration = best_param$max.duration, 
  bp = c(0.5, 10)
)
```

As our reference annotations include all sound files we can evaluate the performance of the detection on the test set as well:

```{r , eval = TRUE, echo = FALSE}

det_test <- read.csv(file.path(rootdir, "/data/processed/taeniopygia", "detection_test_data.csv"))

```

```{r, diagnose, eval = TRUE, echo = TRUE}

diagnose_detection(reference = test_ref, detection = det_test, by.sound.file = FALSE)

```
&nbsp;

The performance on the test data set was also acceptable with a F1 score of 0.95. Note that in the example we use a small subset of sound files for training. More training data might be needed for optimizing a detection routine on larger data sets or recordings with more variable signals or background noise levels.

### Additional tools

The package offers additional tools to simplify acoustic signal detection. The function 'feature_reference' allows to extract information about the duration and frequency range of acoustic signals than can then be used to define tuning parameter values. This can be particularly useful for energy-based detection in which time and frequency attributes of the target signals are required. Detected signals can be labeled as false or true positives with the function 'label_detection'. This allows users to explore the structure of false positives and figure out ways to exclude them. The function 'filter_detections' can be used to remove ambiguous signals (i.e. those labeled as split or merged detection) keeping only those that maximize a specific criterium (i.e. the highest template correlacion). Finally, note that  several templates representing the range of variation in signal structure can be used to detect semi-stereotyped signals when running template-based detection ('template_detection' function).

## Discusion

Here we have showed how to evaluate the performance of acoustic signal detection routines using the package ohun. The package can evaluate detection outputs imported from other software, as well as its own detection routines. The latter can be iterated over combinations of tuning parameters in order to find those values that optimize a detection. Despite signal detection indices being commonly reported when presenting new automatic detection methods, to our knowledge widely applicable performance evaluating routines have not been made available in a free, open source platform. Providing a common framework for acoustic signal detection evaluation can simplify comparing performances of different tools and the selection of those better suited to our system and research question. Note that the tools offered by ohun for diagnosing a detection should not be necessarily limited to acoustic data. Any type of automatic detection in which the time of occurrence of discrete events needs to be found can be evaluated and optimized by comparing it to a reference annotation. For instance detection of specific behaviors in video analysis of animal motor activity (Sturman et al 2020; Hsu & Yttri 2021; DeepEthogram).

The package provides two native detection methods: template-based  and energy-based detection. Compared to new deep learning approaches for finding the occurrence of acoustic signals, the two native methods are relatively simple tools. However, these methods have been widely used by the bioacoustic community and, under the appropriate conditions can reach adequate performance, as we have shown in our two study cases. Deep learning methods tend to require larger computational power and more complex training routines. This might bring unnecessary difficulties when dealing when less challenging detection tasks. Therefore the availability of a wide range of approaches is desired in order to simplify finding the most appropriate tool for the intricacies of our study system and research goals. Note that the tools offered in ohun can be used in a pipeline in which detected signals are further classified using more elaborated discrimination algortihms. Acoustic parameters can be used to quantify the structure of the signals and tell apart target from non-target signals.

Detection routines can take a long time when working with large amounts of acoustic data (e.g. large recordings and/or many files). These are some tips that can help make a routine more time-efficient. 1. Always test procedures on small data subsets. Make sure you are getting decent results on a small subset of recordings before trying to scale up the analysis. 2. Template-based detection is almost always faster than energy-based detection. 4. Run routines in parallel. Parallelization (i.e. the ability to distribute tasks over several cores in your computer) can significantly speed-up routines. All functions for automatic detection and performance evaluation allow users to run analysis in parallel (see `parallel` argument in those functions). Hence a computer with several cores can be helpful for improving efficiency. 4. Sampling rate matters. Detecting signals on low sampling rate files is faster, so we must avoid having Nyquist frequencies much higher than the highest frequency of the target signals. 5. Try using a computer with lots of RAM memory or a computer cluster for working on large amounts of data. These rules are not restricted to ohun and can also be helpful to speed-up routines in other software packages.

There are some additional things to be considered when aiming to automatically detect acoustic signals. When running automatic signals detection try to use your knowledge about the signal structure to determine the initial range. This can be extremely helpful for narrowing down possible parameter values, particularly for energy-based detection. As a general rule, if people have a hard time figuring out where a target signal occurs, detection algorithms will also have a hard time. In cases in which occurrences are ambiguous low performances are expected. Make sure reference tables contain all target signals and only the target signals, otherwise performance optimization can be misleading as the performance of a detection cannot be better than the reference itself. Lastly, avoid having overlapping signals or several signals as a single one (like a multi-syllable vocalization) in the reference table when running an energy-based detector, as they are likely to be identified as separeted units.

https://cran.r-project.org/web/packages/bioacoustics/vignettes/tutorial.html

## References

Sturman, O., von Ziegler, L., Schläppi, C. et al. Deep learning-based behavioral analysis reaches human accuracy and is capable of outperforming commercial solutions. Neuropsychopharmacol. 45, 1942–1952 (2020). https://doi.org/10.1038

Hsu, A.I., Yttri, E.A. B-SOiD, an open-source unsupervised algorithm for identification and fast prediction of behaviors. Nat Commun 12, 5188 (2021). https://doi.org/10.1038/s41467-021-25420-x

### 2022-07-13 manuscript/manuscript.Rmd
first commit

--- 2022-05-11 manuscript/manuscript.Rmd
@@
 ---
-title: 'Ohun: an R package for optimizing automatic acoustic signal detection'
+title: 'ohun: an R package for optimizing automatic acoustic signal detection'
 author:
- - Marcelo Araya-Salas ^1^ * 
- - Grace Smith-Vidaurre ^2^ 
- - Gloriana Chaverri ^3^
- - Fabiola Chirino ^1^
- - Alejandro Rico-Guevara ^4^ 
-date: "`r Sys.Date()`"
 output:
@@
       smooth_scroll: no
+  word_document:
+    toc: yes
+    toc_depth: '3'
+  pdf_document:
+    toc: yes
+    toc_depth: '3'
+always_allow_html: yes
 bibliography: combined_bibs.bib
 citation-style: apa.csl
-link-citations: true
+link-citations: yes
 editor_options:
@@
 ---
+
+```{r load packages, echo = FALSE, message = FALSE, warning=FALSE}
+
+# github packages must include user name ("user/package")
+# knitr is require for creating html/pdf/word reports
+# kableExtra is used to print pretty formatted tables 
+# formatR is used for soft-wrapping code
+# klippy is used for adding a copy button to each code block
+
+pkgs <- c("kableExtra", "knitr", "rprojroot", "sciRmdTheme", "rmdwc", "whereami", "maRce10/ohun", "ggplot2", "viridis", "rmdwc")
+
+# install/ load packages
+out <- lapply(pkgs, function(y) {
+
+  # get pakage name
+  pkg <- strsplit(y, "/")[[1]]
+  pkg <- pkg[length(pkg)]
+
+  # check if installed, if not then install 
+  if (!pkg %in% installed.packages()[,"Package"])  {
+
+      if (grepl("/", y))  remotes::install_github(y, force = TRUE) else
+    install.packages(y) 
+    }
+
+  # load package
+  a <- try(require(pkg, character.only = T), silent = T)
+
+  if (!a) remove.packages(pkg)
+  })
+
+```
+
+```{r theme setting, echo=FALSE}
+# simplest default settings
+sciRmdTheme::set.theme()
+```
+
+```{r custom data frame printing, echo = FALSE, message=FALSE}
+
+# this is a customized printing style data frames 
+# screws up tibble function
+tibble <- function(x, ...) { 
+  x <- kbl(x, digits=4, align= 'c', row.names = FALSE) 
+   x <- kable_styling(x, position ="center", full_width = FALSE,  bootstrap_options = c("striped", "hover", "condensed", "responsive")) 
+   asis_output(x)
+}
+
+registerS3method("knit_print", "data.frame", tibble)
+
+```
+
+```{r set working directory, echo=FALSE}
+
+# set working directory as project directory or one directory above,
+rootdir <- try(rprojroot::find_rstudio_root_file(), silent = TRUE)
+if (is(rootdir, "try-error")) rootdir <-  ".."
+opts_knit$set(root.dir = rootdir)
+
+```
 
@@
 # Alternatively authors ca
-personal_libraries <- c(author_1 = "path/to/lib.file/from/author1/library.bib", author_2 = "path/to/lib.file/from/author2/library.bib")
+personal_libraries <- c(author_1 = "/home/m/Documents/mendeley_libraries/ohun_paper.bib")
 
@@
 if (file.exists(personal_libraries[i]))
- file.copy(from = personal_libraries[i], to = file.path(getwd(), paste0(names(personal_libraries)[i], ".bib")), overwrite = TRUE)
+ file.copy(from = personal_libraries[i], to = file.path(dirname(whereami::thisfile()), paste0(names(personal_libraries)[i], ".bib")), overwrite = TRUE)
 
@@
 # list bib files
-bibs <- list.files(pattern = ".bib$")
+bibs <- list.files(pattern = ".bib$", path = dirname(whereami::thisfile()), full.names = TRUE)
 
 # remove combined from pool
-bibs <- bibs[bibs != "combined_bibs.bib"]
+bibs <- bibs[basename(bibs) != "combined_bibs.bib"]
 
@@
   combined_bibs_l <- lapply(bibs, readLines)
-  names(combined_bibs_l) <- gsub("\\.bib$", "", bibs)
+  names(combined_bibs_l) <- gsub("\\.bib$", "", basename(bibs))
 
@@
 # save pool libraries as combined_bibs.bib
-writeLines(text = combined_bibs, "combined_bibs.bib")
+writeLines(text = combined_bibs, file.path(dirname(whereami::thisfile()), "combined_bibs.bib"))
 
@@
 }
-} else writeLines(text = "empty", "combined_bibs.bib")
+} else writeLines(text = "empty", file.path(dirname(whereami::thisfile()), "combined_bibs.bib"))
 
@@
 
-^1^ Affiliation # 1
+```{r read data, echo=FALSE}
 
-^2^ Affiliation # 2
+path_zebra_finch <- file.path(rootdir, "data/raw/taeniopygia/")
 
-\*To whom correspondence should be addressed
+# read data
+manual_ref_thy <- read.csv(file.path(rootdir, "//data/raw/thyroptera/", "manual_annotations_thyroptera.csv"))
+
+manual_ref_tae <- read.csv(file.path(rootdir, "/data/processed/taeniopygia", "manual_selections_Taeniopygia.csv"))
+
+```
+
+Marcelo Araya-Salas ^1,^ ^2,^ ^3^ *,
+Grace Smith-Vidaurre ^4^,
+Gloriana Chaverri ^3^,
+Fabiola Chirino ^1^ &
+Alejandro Rico-Guevara ^5^
+
+&nbsp;
+
+^1^ Centro de Investigación en Neurociencias, Universidad de Costa Rica, San José, Costa Rica
+
+^2^ Escuela de Biología, Universidad de Costa Rica, San José, Costa Rica
+
+^3^ Sede del Sur, Universidad de Costa Rica, Golfito, Costa Rica
+
+^4^ GRACE'S AFFILIATION
+
+^5^  Department of Biology, University of Washington, Seattle, USA
+
+\* *To whom correspondence should be addressed*
 
 ### Keywords:
+
+---
+
+&nbsp;
+
+```{r word count, eval = TRUE, echo = FALSE, results='asis'}
+
+# word count without code chunks
+word_count <- rmdcount(whereami::thisfile())
+
+# word count with code chunks
+word_count_chunk <- rmdcount(whereami::thisfile(), exclude='')
+
+# print results
+cat(paste0("Word count: ", word_count[3], " (including code chunks: ", word_count_chunk[3], ")"))
+
+```
 
@@
 
+Animal acoustic signals are widely used in a diversity of research areas.
+This is partly due to the relative ease in which they can be registered for wide range of taxonomic groups and research settings.
+However, bioacoustic research can easily generate large amounts of data which might proof challenging to analyze in a timely manner.
+Many tools for the automatic detection of signals are currently available.
+However, choosing the right approach for a specific task might turn difficult as there are no available tools that can provide a common framework for evaluating detection performance.
+Here we present the new R package ohun.
+The package is intended to facilitate the automatic detection of acoustic signals, by providing functions to diagnose and optimize detection routines as well as compare the performance of different detection approaches.
+The package makes use of reference annotations containing the time position of target signals in a training data set to evaluate the performance of detection routines, using common signal detection theory indices.
+This can be done with routine outputs imported from other software and from detection run within the package itself.
+The package also provides a set functions to organize acoustic data sets in a format amenable for detection analyses.
+It also includes an implementation of two automatic detection methods commonly used in bioacustic analysis: energy-based detection and template-based detection.
+We included examples on how to evalute detections from external software and showcase the package usage with study cases on zebra-finch (*Taenopygia gutata*) songs and Spix's disc-winged bat (*Thyroptera tricolor*) ultrasonic social calls.
+Finally, we provide some general suggestions to improve detection performance.
+
 ## Introduction
 
-- 1st paragraph:
+Animal acoustic signals are widely used to investigate a variety of questions in highly diverse areas, ranging from neurobiology to community ecology and evolutionary biology.
+The wide usage of animal sounds in research is partly due to the fact that they can be easily registered using non-intrusive methods, can be obtained in a variety of settings from laboratories to natural areas, and the equipment required for registering an analyzing these signals has become increasingly inexpensive.
+In addition, the existence of online repositories and growing number has facilitated the study of these traits at larger taxonomic and geographic scales.
+However, the adoption of bioacustic approaches in research can also imply large amounts of data (i.e.
+lots of recordings), which can be challenge to analyzed manually by human observers.
+As a result, a growing number of computational tools for the analysis of acoustic features on those signals is increasingly available (reviewd by @Stowell2022), reflecting the need of automated approaches for efficiently conducting these analyses.
 
-    - Animal acoustic signals are widely used in scientific research in a highly diverse areas, ranging from neurobiology to community ecology and evolutionary biology. 
-    - The wide usage of animal sounds in research is partly due to the fact that they can be easily registered using non-intrusive methods, in a variety of setting from laboratories to natural areas and equipment for recording these signals has become increasingly inexpensive and accesible.
+<!-- Many of these tools are supplied in the form of freely available software written on programming environments familiar to the scientific community (seewave, warbleR, scikit-mad), further simplifying their implementation for research.
+-->
+<!-- Bioacoustic research can easily generate large amounts of data, which sometimes may proof challenging to analyze in a timely manner.
+Annotations that require precise location of signals in frequency and time, most commonly used in behavioral and phenotypic evolution research, remain among the most time consuming phases of analysis for this type of data.
+Fortunately, several tools have been develop to assist researchers in this matter.
+Indeed, new methods for automatic annotation of acoustic signals are continuously developed .
+-->
 
-    In addition, the existence of online repositories and growing number of acoustic data set avialable to the comunnity.
-As a result, a growing variety of a computational tools for the analysis of acoustic features on those signals is increasingly available, which reflects the growing relevance of bioacustics as a relevant tool in the scientific tool kit.
-(many of those in the form of freely available software written on programming environments familiar to the scientific community (seewave, warbleR, scikit-mad))
+The growing availability of tools for automatic detection of acoustic events, particularly as free software, is expected to further simplify acoustic data processing, making it accessible to a wider range of users and scientific questions.
+However, this diversity of tools also posits a challenge as it can be hard to navigate.
+In this regard, the use of standard approaches for evaluating the performance of automatic detection tools might prove helpful to inform the decision about which method better fit our study system.
+The performance of automatic acoustic signal detection routines have been evaluated using standard indices from signal detection theory (CITATIONS).
+In its basic form, performance is assessed by comparing the output of a detection routine against a "gold standard" reference in which all the target signals have been annotated.
+This comparison allows to quantify the number of correctly detected signals (true positives), wrongly detected signals (false positives) and missed signals (false negatives) as well as additional metrics derived from these indices (e.g.
+recall, precision).
+However, the fact that acoustic signals are not being evaluated as discrete classification units (as opposed to for instance identifying species from pictures) demand additional information to fully diagnose detection performance.
+This is particularly relevant when the precise time position of signals is needed as is common for research in which measures of the acoustic structure of signals is the main goal.
+For instance, the same signal can be detected as several separated signals, the inferred time position can be far off from the the target signal position, or several signals can be detected as a single one.
+Therefore, metrics that account for this other performance dimensions can be a valuable tool to properly diagnose automatic acoustic signal detection.
 
-- 2nd paragraph:
+Here we present the new R package ohun.
+The package is intended to facilitate the automatic detection of acoustic signals, providing functions to diagnose specific aspects of acoustic detection routines in order to simplify their optimization.
+The package makes use of reference annotations containing the time position of target signals in a training data set to evaluate the performance of detection routines.
+This can be done with routine outputs imported from other software as well as detection run within the package itself.
+The package also provides a set functions to explore acoustic data sets and organize them in a format amenable for detection analyses.
+In addition, it offers implementations of two automatic detection methods commonly used in bioacustic analysis: energy-based detection and template-based detection.
+We explain how to explore and format acoustic data sets, how acoustic signal detection routines can be evaluate and then showcase the package usage with study cases on zebra-finch songs (*Taenopygia gutata*) and Spix's disc-winged bats (*Thyroptera tricolor*) which represent different recording settings (lab and flight cages) and signal types (sonic mating signals and ultrasonic social calls).
 
-    - Bioacoustic approaches can easily generate large amounts of data, which sometimes may proof challenging to analyze in a timely manner.
-    - Much effort has been put into developing computational tools that can simplify acoustic analyses. For instance,....
-    - However, some key aspects of the analysis remain challenging, despite the great advances witness in the last decade.
-    - The annotations that require a precise definition of signal location in frequency and time, most commonly used in behavioral and phenotypic eovlution research, remain among the most time consuming phases of analysis for this type of data.
-Fortunately, several tools have been develop to assist researchers in this matter.
-Indeed, new methods for automatic annotation of acoustic signals are contnuously developed.
-This growing availability of tools, particularly as free software, is expected to further simplify acoustic data processing, making it accesible to wider range of user and scientific questions.
+## Formatting acoustic data sets
 
-- 3rd paragraph:
+Having some sense of the format and size of the sound files to be analyzed is an important step for avoiding downstream errors and informing our time performance expectations.
+Several functions can facilitate getting acoustic data sets in the right format prior to automatic detection.
+The function `feature_acoustic_data` prints a summary of the duration, size and format of all the recordings in a folder.
+Here we explore the acoustic data set of Zebra finch's songs (suppl.
+mat):
 
-    - Here we present the new R package ohun.
-The package is intended to facilitate the automatic detection of acoustic signals, providing functions to diagnose and optimize detection routines.
-The package makes use of reference annotations containing the time position of all target signals in a training data set to evaluate the performance of detection routines.
-This can be done with routine outputs imported from other software as well as detection run with ohun.
-The package offers an implementation of two automatic detection methods commonly used in bioacustic analysis: energy-based detection and template-based detection.
-We first explain how acoustic signal detection routines can be evaluate and then showcase the package usage with study cases on zebra-finch songs (*Taenopygia gutata*) and Spix's disc-winged bats (*Thyroptera tricolor*) which represent different recording settings (lab and flight cages) and signal types (sonic mating signals and ultrasonic social calls).
+```{r, eval = FALSE, echo = TRUE}
+
+# working directory
+path_zebra_finch <- "path_to_zebra_finch_files"
+
+feature_acoustic_data(path = path_zebra_finch)
+
+```
+
+```{r, eval = TRUE, echo = FALSE}
+
+feature_acoustic_data(path = path_zebra_finch)
+
+```
+
+In this case all recordings have format (.wav files, 44.1 kHz sampling rate, 16 bit resolution and a single channel).
+We can also check the duration and size of files.
+This information is important as some tuning parameters of detection routines can behave differently depending on file format (e.g.
+time window size is affected by sampling rate) or simply because some software might can only work on specific sound file formats.
+In addition, long sound files might be hard to analyzed in regular computers and might have to be split in shorter clips.
+In the latter case the function `split_acoustic_data` can be used to produce those clips:
+
+```{r, eval = FALSE, echo = FALSE}
+source("~/Dropbox/R_package_testing/ohun/R/split_acoustic_data.R")
+split_info <- split_acoustic_data(path = path_zebra_finch, sgmt.dur = 5)
+
+write.csv(split_info, "./data/processed/split_data_for_manuscript.csv", row.names = FALSE)
+
+```
+
+```{r, eval = FALSE, echo = TRUE}
+
+split_info <- split_acoustic_data(path = path_zebra_finch, sgmt.dur = 5)
+
+head(split_info)
+
+```
+
+```{r, eval = TRUE, echo=FALSE}
+
+split_info <- read.csv(file.path(rootdir, "data/processed/split_data_for_manuscript.csv"))
+
+head(split_info)
+```
+
+The output shows to which time segments in the original sound files the clips belong to.
+If an annotation table is supplied (argument 'X') the function will adjust the annotations so they refer to the position of the signals in the clips.
+This can be helpful when reference tables have been annotated on the original sound files.
+
+Annotations can also be explored using the function `feature_reference`.
+It returns about the mean and range of signal and gap duration (time intervals between selections), bottom and top frequency and of the number of annotations by sound file.
+If the path to the sound files is supplied the duty cycle (fraction of a sound file corresponding to signals) and peak amplitude (highest amplitude in a detection) are also returned:
+
+```{r, eval = FALSE, echo = TRUE}
+
+# read reference annotations
+manual_ref_tae <- read.csv(file.path(path_zebra_finch, "manual_selections_Taeniopygia.csv"))
+
+# explore annotations
+feature_reference(reference = manual_ref_tae, path = path_zebra_finch)
+
+```
+
+```{r, eval = TRUE, echo = FALSE}
+
+manual_ref_tae <- read.csv(file.path(rootdir, "/data/processed/taeniopygia", "manual_selections_Taeniopygia.csv"))
+
+feature_reference(reference = manual_ref_tae, path = path_zebra_finch)
+
+```
 
@@
 However, TNs cannot always be easily defined in the context of acoustic signal detection as noise cannot always be partitioned in discrete units.
-Hence, the package makes use of TPs, FPs and FNs to calculate three additional indices and can further assist with evaluating the performance of a detection routine.
+Hence, the package makes use of TPs, FPs and FNs to calculate three additional indices that can further assist with evaluating the performance of a detection routine.
 
@@
 - Precision: correct detections relative to total detections (TPs / (TPs + FPs)).
-- F1 score: combines recall and precision as the harmonic mean of these two, so it provides a single value for evaluating performance (a.k.a. F-measure or Dice similarity coefficient).
+- F1 score: combines recall and precision as the harmonic mean of these two, providing a single value for evaluating performance (a.k.a. F-measure or Dice similarity coefficient).
 
@@
 However, perfect detection cannot always be reached and some compromise between detecting all target signals plus some noise (recall = 1 & precision < 1) and detecting only target signals but not all of them (recall < 1 & precision = 1) might be warranted.
-The right balance between these two extremes will be given by the relative costs of missing signals and mistaking noise for signals.
-Hence, these indices provide an useful framework for diagnosing and optimizing the performance of a detection routine.
+The right balance between these two extremes will be given by the relative costs of missing signals and mistaking noise for signals, given the specific goals of the study.
+These indices provide an useful framework for diagnosing and optimizing the performance of a detection routine.
 
-The package ohun provides a set of tools to evaluate the performance of an acoustic signal detection based on the indices described above.
-To accomplish this, the result of a detection routine is compared against a reference table containing the time position of all target signals in the sound files.
+The package offers tools to evaluate the performance of an acoustic signal detection based on the indices described above.
+To accomplish this, annotations derived from a detection routine are compared against a reference table containing the time position of all target signals in the sound files.
+For instance, the following code evaluates a routine run in Raven Pro 1.6 (XXXX) using the "band limited energy detector" option (minimum frequency: 0.8 kHz; maximum frequency: 22 kHz; minimum duration: 0.54989 s; maximum duration: 0.54989s; minimum separation: 0.02268 s) on a subset of the Zebra finch recordings described below:
 
-## Signal detection
+```{r, eval = TRUE, echo = FALSE}
 
-Finding the position of signals in a sound file is a challenging task.
-ohun offers two methods for automatic signal detection: template-based and energy-based detection.
+library(Rraven)
+
+raven_detec <- imp_raven(path = file.path(rootdir, "/data/raw/"), files = "Raven.Band.Limited.Energy.Detector.selections.txt", warbler.format = TRUE)
+
+# raven_ref <- manual_ref[manual_ref$sound.files %in% raven_detec$sound.files, ]
+
+```
+
+```{r, eval = TRUE, echo = TRUE}
+
+diagnose_detection(reference = manual_ref_tae, detection = raven_detec)
+
+```
+
+The output shows the indices described above. The function also allows to detail those indices by sound file. Here we show the first 6 files:
+
+```{r, eval = TRUE, echo = TRUE}
+
+diag_raven <- diagnose_detection(reference = manual_ref_tae, detection = raven_detec, by.sound.file = TRUE)
+
+head(diag_raven)
+```
+
+Diagnostics from routines using different tuning parameters can be used to identify the parameter values that optimize a detection.
+The process of evaluating different routines for detection optimization is incorporated into the two signal detection approaches provided natively by ohun, which we depict in the following section.
+Note that the detection with Raven Pro does not necessarily reflect the best performance of this software and has been included only as an example on evaluating detection from external sources rather than a direct comparison of performance to ohun.
+
+## Signal detection with ohun
+
+The package offers two methods for automatic signal detection: template-based and energy-based detection.
 These methods are better suited for highly stereotyped or good signal-to-noise ratio (SNR) signals, respectively.
-If the target signals don’t fit these requirements, more elaborated methods (i.e.
-machine learning approaches) are warranted:
+If the target signals do not fit these requirements, more elaborated methods (i.e.
+machine/deep learning approaches) are warranted.
 
@@
 
-### Ultrasonic social calls in Spix's disc-winged bats
+### Template detection on ultrasonic social calls of  Spix's disc-winged bats
 
-We recorded 30 individuals of Spix's disc-winged bats at Baru Biological Station, in south-western Costa Rica in January 2018.
+We recorded `r length(unique(manual_ref_thy$sound.files))` individuals of Spix's disc-winged bats at Baru Biological Station, in south-western Costa Rica in January 2020.
 Bats were captured at their roosting sites (furled leaves of Zingeberaceae plants).
-Each individual bat was released in a large flight cage (9 x 4 x 3 m) for a 5 min period and their ultrasonic inquiry calls were recorded using a condenser microphone (CM16, Avisoft Bioacoustics, Glienike/
-Nordbahn, Germany) through an Avisoft UltraSoundGate 116Hm connected to a laptop computer running Avisoft-Recorder software (sampling rate of 500 kHz, 16-bit amplitude resolution).
+Each individual bat was released in a large flight cage (9 x 4 x 3 m) for a 5 min period and their ultrasonic inquiry calls were recorded using a condenser microphone (CM16, Avisoft Bioacoustics, Glienike/Nordbahn, Germany) through an Avisoft UltraSoundGate 116Hm plugged into a laptop computer running Avisoft-Recorder software.
+Recordings were made at a sampling rate of 500 kHz and an amplitude resolution of 16 bits.
 
 Recordings were manually annotated using Raven Pro 1.6 (XXXX).
-Annotations were created by visual inspection of spectrograms, in which the start and end of signals were determined by the location of the continuous traces of power spectral entropy of target signals.
-Annotations were made with a time window of 512 samples with a 70% of overlap.
-Annotations were then imported into R using the R package Rraven (XXXX).
+Annotations were created by visual inspection of spectrograms, in which the start and end of signals were determined by the location of the continuous traces of power spectral entropy of the target signals.
+A total of `r nrow(manual_ref_thy)` calls were annotated (~`r round(nrow(manual_ref_thy) / 30, 0)` calls per recording).
+Annotations were made with a time window of 200 samples and 70% of overlap.
+Annotations were then imported into R using the package Rraven (XXXX).
 
-Inquiry calls of Spix's disc-winged bats are stereotyped signals.
-Most variation is found between individuals although the basic structure of a short, downward broadband frequency modulation (FIG XXX) is always shared.
-Template-based detection is an useful approach when there is little structural differences in the target signals.
-Therefore we used the function
+Inquiry calls of Spix's disc-winged bats are structurally stereotyped.
+Most variation is found between individuals although the basic form of a short, downward broadband frequency modulation is always shared (Fig.
+BAT-SPECTRO, Araya-Salas et al 2021 ontogeny).
 
-### Zebra finch songs
+```{r catalog, out.width = "100%", echo = FALSE, fig.align= "left",fig.cap="Fig.
+BAT-SPECTRO.
+Example spectrograms of Spix's disc-winged social calls for each of the 30 recordings used in the analysis.
+The highest signal-to-noise ratio call by sound file are shown.
+The time scale range is 71 ms and the frequency range 10-44 kHz", error=TRUE}
 
-We also used recordings from 18 zebra finch males. These recordings were obtained ....
+# knitr::include_graphics("~/Dropbox/Projects/ohun_paper/data/raw/thyroptera/Catalog_p1.jpeg")
+knitr::include_graphics((file.path(rootdir, "data/raw/thyroptera/Catalog_p1.jpeg")))
+```
 
-## Results
+Template-based detection is a useful approach when there is little structural differences in the target signals.
+We used this approach in ohun to detect inquiry calls.
+To do this, we tested the performance of three acoustic templates on a training subset of five sound files.
+The function `get_templates` finds several signals representative of the variation in signal structure.
+This function measures several spectrographic parameters which are then summarized using Principal Component Analysis.
+The first two components are used to project the acoustic space.
+On this space the function defines sub-spaces as equal-size slices of a circle centered at the centroid of the acoustic space.
+Templates are then selected as those closer to the centroid within each of the sub-spaces, including the centroid for the entire acoustic space.
+The user needs to define the number of sub-spaces in which the acoustic space will be split.
+
+```{r, eval = FALSE}
+
+# read manual annotations
+manual_ref_thy <- read.csv("manual_annotations_thyroptera.csv")
+
+# get random subset of 5 sound files for training
+set.seed(1)
+train_files <- sample(unique(manual_ref_thy$sound.files), size = 5)
+train_ref <- manual_ref_thy[manual_ref_thy$sound.files %in% train_files, ]
+
+# the rest for testin g
+test_files <- setdiff(manual_ref_thy$sound.files, train_files)
+test_ref <- manual_ref_thy[manual_ref_thy$sound.files %in% test_files, ]
+
+# find templates
+templates <- get_templates(train_ref, path = data_path,  bp = c(10, 50), ovlp = 70, wl = 200, n.sub.spaces = 3)
+
+```
+
+```{r, echo = FALSE, fig.height= 6, fig.cap= "FIG.
+ACOUSTIC-SPACE.
+Acoustic space defined as the the first two component of a Principal Component Analysis on spectrographic parameters.
+Templates are selected as those closer to the centroid for each of the sub-spaces.
+Gray dashed lines delimit the region of sub-spaces.
+Yellow circles around points highlight the position of the signals selected as templates."}
+
+# read manual annotations
+manual_ref_thy <- read.csv(file.path(rootdir, "/data/raw/thyroptera/", "manual_annotations_thyroptera.csv"))
+
+# get random subset of 5 sound files for training
+set.seed(1)
+train_files <- sample(unique(manual_ref_thy$sound.files), size = 5)
+train_ref <- manual_ref_thy[manual_ref_thy$sound.files %in% train_files, ]
+
+# the rest for testing
+test_files <- setdiff(manual_ref_thy$sound.files, train_files)
+test_ref <- manual_ref_thy[manual_ref_thy$sound.files %in% test_files, ]
+
+source("~/Dropbox/R_package_testing/ohun/R/get_templates.R")
+
+templates <- get_templates(train_ref, path = file.path(rootdir, "/data/raw/thyroptera/"),  bp = c(10, 50), ovlp = 70, wl = 200, n.sub.spaces = 3)
+
+```
+
+```{r, eval = TRUE, echo = FALSE}
+
+data_path <- file.path(rootdir, "/data/raw/thyroptera/")
+
+manual_ref_thy <- read.csv(file.path(data_path, "manual_annotations_thyroptera.csv"))
+
+# get random subset of 5 sound files for training
+set.seed(1)
+train_files <- sample(unique(manual_ref_thy$sound.files), size = 5)
+train_ref <- manual_ref_thy[manual_ref_thy$sound.files %in% train_files, ]
+
+# the rest for testing
+test_files <- setdiff(manual_ref_thy$sound.files, train_files)
+test_ref <- manual_ref_thy[manual_ref_thy$sound.files %in% test_files, ]
+```
+
+The output of the `get_templates` includes an acoustic space plot (FIG.
+ACOUSTIC-SPACE) in which the position of the signals selected as templates is highlighted.
+In the following code we used those templates for detecting calls.
+The code iterates a template-based detection on the training data set across a range of correlation thresholds, in order to find the combination of threshold and template with the best performance.
+
+```{r, eval = FALSE}
+
+# get correlation vectors
+corr_templ_train <- template_correlator(
+  templates = templates,
+    path = data_path, 
+  files = unique(train_ref$sound.files), 
+  hop.size = 10, 
+  ovlp = 70
+  )
+
+# evaluate detection for different correlation thresholds
+opt_detec_train <- optimize_template_detector(
+  reference = train_ref, 
+  template.correlations = corr_templ_train,
+  threshold = seq(0.05, 0.5, 0.01)
+  )
+
+```
+
+Note that the correlation vectors are estimated first (i.e.
+vectors of correlation values across sound files, `template_correlator()`) and then the correlation thresholds are optimized on them (`optimize_template_detector()`).
+The output of `optimize_template_detector()` contains the detection performance indices for each combination of templates and thresholds.
+Table TEMPLATE PERFORMANCE shows the two highest performance runs for each template.
+
+```{r, eval = TRUE, echo = FALSE}
+
+opt_detec_train <- readRDS(file.path(rootdir, "/data//processed/optimization_results_3_templates_thyroptera.RDS"))
+
+# opt_detec_train <- opt_detec_train[order(opt_detec_train$templates, opt_detec_train$f1.score, decreasing = TRUE), ]
+
+# sub_opt_detec_train <- opt_detec_train[ave(-opt_detec_train$f1.score, opt_detec_train$templates, FUN = rank) <= 4, ]
+
+sub_odt <- lapply(unique(opt_detec_train$templates), function(x) {
+
+    X <- opt_detec_train[opt_detec_train$templates == x, ]
+    X <- X[order(X$f1.score, decreasing = TRUE), ]
+    return(X[1:2, ])
+    })
+
+sub_opt_detec_train <- do.call(rbind, sub_odt)
+
+kbl <- kable(sub_opt_detec_train[, c("threshold", "templates", "true.positives", "false.positives", "false.negatives", "recall", "precision", "f1.score")], row.names = FALSE, caption = "TABLE TEMPLATE PERFORMANCE.
+Performance diagnostic of template-based detections using four templates across several threshold values.
+Only the two highest performance iterations for each template are shown.")
+
+kable_styling(kbl, bootstrap_options = c("striped", "hover",  full_width = FALSE))
+
+```
+
+We can explore the performance of each template in more detail by looking at the change in F1 score across thresholds (FIG. THRESHOLD vs F1.SCORE).
+
+```{r, eval = TRUE, echo = FALSE, fig.height= 3, fig.align='center', fig.cap= "FIG. THRESHOLD vs F1.SCORE. Shows the changes in F1 score across the range of threshold values"}
+
+agg_f1 <- aggregate(f1.score ~ threshold + templates, data = opt_detec_train,
+mean)
+
+ggplot(agg_f1, aes(x = threshold, y = f1.score, group = templates, color = templates)) +
+    geom_line() + 
+    geom_point() + 
+    scale_color_viridis_d(end = 1, labels = c("Centroid", "1st", "2nd", "3rd"), alpha = 0.7) + 
+    labs(x= "Cross-correlation threshold", y = "F1 score", color = "Templates")
+
+```
+
+In this example the "centroid" template, produced the best performance (TABLE TEMPLATE PERFORMANCE; FIG.
+THRESHOLD vs F1.SCORE).
+Hence, we will use this template for detecting calls on the rest of the data.
+The following code extracts this template from the reference annotation table and use it to find inquiry calls on the testing data set:
+
+```{r, detect on all data 67, eval = FALSE, echo = TRUE}
+
+# get correlation vectors for test files
+corr_templ_test <- template_correlator(
+  templates = templates[templates$sound.file == "centroid", ],
+    path = data_path, files = unique(test_ref$sound.files), 
+  hop.size = 10, 
+  ovlp = 70
+  )
+
+# detect on test files
+detec_test <- template_detector(
+  template.correlations = corr_templ_test,
+    threshold = 0.45
+  )
+
+diagnose_detection(reference = test_ref, detection = detec_test)
+
+```
+
+```{r, eval = TRUE, echo = FALSE}
+
+detec_test <- readRDS(file.path(rootdir, "/data/processed/detection_thyroptera_all_67%_template.RDS"))
+
+diagnose_detection(reference = test_ref, detection = detec_test)
+```
+
+The last line of codes evalutes the detection on the test data set, which shows a good performance for both recall and precision.
+
+### Energy-based detection on zebra finch songs
+
+We used recordings from 18 zebra finch males. These recordings were obtained ....
+
+Zebra finch vocalizations are composed by multiple elements (i.e.
+distinct patterns of continuous traces of power spectral entropy in the spectrogram separated by time gaps) that can vary importantly in key features as duration and frequency range (ZEBRAFINCH-SPECTRO).
+However, as the recorded signals show a good signal-to-noise ratio, they can potentially be detected using an energy-based approach.
+```{r catalog zebra f, out.width = "100%", echo = FALSE, fig.align= "left",fig.cap="Fig.
+ZEBRAFINCH-SPECTRO.
+Example spectrograms of male zebra finch songs for each of the 18 sound files used in the analysis.
+The highest signal-to-noise ratio call by sound file are shown.
+The time scale range is 359 ms and the frequency range 0-11 kHz.
+Signals have been highlighted for visualization purposes only.", error=TRUE}
+
+knitr::include_graphics((file.path(rootdir, "data/raw/taeniopygia/same.time-Catalog_p1.jpeg")))
+
+```
+
+The following code loads the reference annotations and split them in two data set for training (3 sound files) and testing (15 sound files):
+```{r, eval = TRUE, echo = FALSE}
+
+manual_ref_tae <- read.csv(file.path(rootdir, "/data/processed/taeniopygia", "manual_selections_Taeniopygia.csv"))
+
+set.seed(450)    
+train_files <- sample(unique(manual_ref_tae$sound.files), 3)
+
+test_files <- setdiff(manual_ref_tae$sound.files, train_files)
+
+train_ref <- manual_ref_tae[manual_ref_tae$sound.files %in% train_files, ]
+test_ref <- manual_ref_tae[manual_ref_tae$sound.files %in% test_files, ]
+
+```
+
+```{r, eval = FALSE, echo = TRUE}
+
+manual_ref_tae <- read.csv("manual_selections_Taeniopygia.csv")
+
+set.seed(450)    
+train_files <- sample(unique(manual_ref_tae$sound.files), 3)    
+test_files <- setdiff(manual_ref_tae$sound.files, train_files)
+
+train_ref <- manual_ref_tae[manual_ref_tae$sound.files %in% train_files, ]
+test_ref <- manual_ref_tae[manual_ref_tae$sound.files %in% test_files, ]
+
+```
+
+Now we can optimize the detection parameters using the function `optimize_energy_detector`.
+This function runs a detection for all possible combinations of tuning parameters.
+The code below tries three minimum duration and maximum duration values and two hold time values (which merges signals within the specified time interval):
+
+```{r}
+feature_reference(test_ref)
+```
+
+```{r  optimize automated detection, eval = FALSE, echo = TRUE}
+
+opt_det_train <- optimize_energy_detector(
+  reference = train_ref, 
+  files = train_files, 
+  threshold = c(1, 5), 
+  hop.size = 11.6, 
+  smooth = c(5, 10), 
+  hold.time = c(0, 5), 
+  min.duration = c(5, 15, 25), 
+  max.duration = c(275, 300, 325), 
+  bp = c(0.5, 10)
+)
+
+```
+
+The output (`opt_det_train`) shows the performance indices for each of those combinations, here we show the 10 combinations with the highest F1 score:
+
+```{r, eval = FALSE, echo = TRUE}
+
+# subset with highest performance
+opt_det_train <- opt_det_train[order(opt_det_train$f1.score, decreasing = TRUE), ]
+
+head(opt_det_train, 10)
+```
+
+```{r, eval = TRUE, echo = FALSE}
+
+opt_det_train <- read.csv(file.path(rootdir, "/data/processed/taeniopygia", "detection_optimization.csv"))
+
+# subset with highest performance
+sub_opt_det_train <- opt_det_train[order(opt_det_train$f1.score, decreasing = TRUE), ]
+
+sub_opt_det_train <- sub_opt_det_train[1:10, c("threshold", "smooth", "hold.time", "min.duration", "max.duration", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f1.score")]
+
+sub_opt_det_train
+```
+
+We now can use the tuning parameter values that produce the best performance to detect signals on the test data set:
+
+```{r, select best params, eval = FALSE, echo = TRUE}
+
+best_param <- opt_det_train[which.max(opt_det_train$f1.score), ]
+
+det_test <- energy_detector(
+  files = test_files, 
+  threshold = best_param$threshold, 
+  hop.size = 11.6, 
+  smooth = best_param$smooth, 
+  hold.time = best_param$hold.time, 
+  min.duration = best_param$min.duration, 
+  max.duration = best_param$max.duration, 
+  bp = c(0.5, 10)
+)
+```
+
+As our reference annotations include all sound files we can evaluate the performance of the detection on the test set as well:
+
+```{r , eval = TRUE, echo = FALSE}
+
+det_test <- read.csv(file.path(rootdir, "/data/processed/taeniopygia", "detection_test_data.csv"))
+
+```
+
+```{r, diagnose, eval = TRUE, echo = TRUE}
+
+diagnose_detection(reference = test_ref, detection = det_test, by.sound.file = FALSE)
+
+```
+&nbsp;
+
+The performance on the test data set was also acceptable with a F1 score of 0.95.
+Note that in the example we use a small subset of sound files for training.
+More training data might be needed for optimizing a detection routine on larger data sets or recordings with more variable signals or background noise levels.
+
+### Additional tools
+
+The package offers additional tools to simplify acoustic signal detection.
+The function 'feature_reference' allows to extract information about the duration and frequency range of acoustic signals than can then be used to define tuning parameter values.
+This can be particularly useful for energy-based detection in which time and frequency attributes of the target signals are required.
+Detected signals can be labeled as false or true positives with the function 'label_detection'.
+This allows users to explore the structure of false positives and figure out ways to exclude them.
+The function 'filter_detections' can be used to remove ambiguous signals (i.e.
+those labeled as split or merged detection) keeping only those that maximize a specific criterium (i.e.
+the highest template correlacion).
+Finally, note that  several templates representing the range of variation in signal structure can be used to detect semi-stereotyped signals when running template-based detection ('template_detection' function).
 
@@
 
-- Despite signal detection indices being commonly reported when presenting new automatic detection methods, to our knowledge widely applicable performance evaluating routines have not been made available.
-- Can be useful for analysis in which relatively small number of recordings are used and also when only a limited amount of training data is available.
+Here we have showed how to evaluate the performance of acoustic signal detection routines using the package ohun.
+The package can evaluate detection outputs imported from other software, as well as its own detection routines.
+The latter can be iterated over combinations of tuning parameters in order to find those values that optimize a detection.
+Despite signal detection indices being commonly reported when presenting new automatic detection methods, to our knowledge widely applicable performance evaluating routines have not been made available in a free, open source platform.
+Providing a common framework for acoustic signal detection evaluation can simplify comparing performances of different tools and the selection of those better suited to our system and research question.
+Note that the tools offered by ohun for diagnosing a detection should not be necessarily limited to acoustic data.
+Any type of automatic detection in which the time of occurrence of discrete events needs to be found can be evaluated and optimized by comparing it to a reference annotation.
+For instance detection of specific behaviors in video analysis of animal motor activity (Sturman et al 2020; Hsu & Yttri 2021; DeepEthogram).
 
-## Improving detection speed
+The package provides two native detection methods: template-based  and energy-based detection.
+Compared to new deep learning approaches for finding the occurrence of acoustic signals, the two native methods are relatively simple tools.
+However, these methods have been widely used by the bioacoustic community and, under the appropriate conditions can reach adequate performance, as we have shown in our two study cases.
+Deep learning methods tend to require larger computational power and more complex training routines.
+This might bring unnecessary difficulties when dealing when less challenging detection tasks.
+Therefore the availability of a wide range of approaches is desired in order to simplify finding the most appropriate tool for the intricacies of our study system and research goals.
+Note that the tools offered in ohun can be used in a pipeline in which detected signals are further classified using more elaborated discrimination algortihms.
+Acoustic parameters can be used to quantify the structure of the signals and tell apart target from non-target signals.
 
 Detection routines can take a long time when working with large amounts of acoustic data (e.g.
-large sound files and/or many sound files).
-These are some useful points to keep in mine when trying to make a routine more time-efficient:
+large recordings and/or many files).
+These are some tips that can help make a routine more time-efficient.
+1.
+Always test procedures on small data subsets.
+Make sure you are getting decent results on a small subset of recordings before trying to scale up the analysis.
+2.
+Template-based detection is almost always faster than energy-based detection.
+4.
+Run routines in parallel.
+Parallelization (i.e.
+the ability to distribute tasks over several cores in your computer) can significantly speed-up routines.
+All functions for automatic detection and performance evaluation allow users to run analysis in parallel (see `parallel` argument in those functions).
+Hence a computer with several cores can be helpful for improving efficiency.
+4.
+Sampling rate matters.
+Detecting signals on low sampling rate files is faster, so we must avoid having Nyquist frequencies much higher than the highest frequency of the target signals.
+5.
+Try using a computer with lots of RAM memory or a computer cluster for working on large amounts of data.
+These rules are not restricted to ohun and can also be helpful to speed-up routines in other software packages.
 
-- Always test procedures on small data subsets
-- `template_detector()` is faster than `energy_detector()`
-- Parallelization (see `parallel` argument in most functions) can significantly speed-up routines, but works better on Unix-based operating systems (linux and mac OS)
-- Sampling rate matters: detecting signals on low sampling rate files goes faster, so we should avoid having nyquist frequencies (sampling rate / 2) way higher than the highest frequency of the target signals (sound files can be downsampled using warbleR's [`fix_sound_files()`](https://marce10.github.io/warbleR/reference/selection_table.html))
-- Large sound files can make the routine crash, use `split_acoustic_data()` to split both reference tables and files into shorter clips.
-- Think about using a computer with lots of RAM memory or a computer cluster for working on large amounts of data
-- `thinning` argument (which reduces the size of the amplitude envelope) can also speed-up `energy_detector()`
-
-## Additional tips
-
-- Use your knowledge about the signal structure to determine the initial range for the tuning parameters in a detection optimization routine
-- If people have a hard time figuring out where a target signal occurs in a recording, detection algorithms will also have a hard time
-- Several templates representing the range of variation in signal structure can be used to detect semi-stereotyped signals
-- Make sure reference tables contain all target signals and only the target signals. The performance of the detection cannot be better than the reference itself.
-- Avoid having overlapping signals or several signals as a single one (like a multi-syllable vocalization) in the reference table when running an energy-based detector
-- Low-precision can be improved by training a classification model (e.g. random forest) to tell signals from noise
+There are some additional things to be considered when aiming to automatically detect acoustic signals.
+When running automatic signals detection try to use your knowledge about the signal structure to determine the initial range.
+This can be extremely helpful for narrowing down possible parameter values, particularly for energy-based detection.
+As a general rule, if people have a hard time figuring out where a target signal occurs, detection algorithms will also have a hard time.
+In cases in which occurrences are ambiguous low performances are expected.
+Make sure reference tables contain all target signals and only the target signals, otherwise performance optimization can be misleading as the performance of a detection cannot be better than the reference itself.
+Lastly, avoid having overlapping signals or several signals as a single one (like a multi-syllable vocalization) in the reference table when running an energy-based detector, as they are likely to be identified as separeted units.
 
@@
 ## References
+
+Sturman, O., von Ziegler, L., Schläppi, C.
+et al.
+Deep learning-based behavioral analysis reaches human accuracy and is capable of outperforming commercial solutions.
+Neuropsychopharmacol.
+45, 1942–1952 (2020).
+https://doi.org/10.1038
+
+Hsu, A.I., Yttri, E.A. B-SOiD, an open-source unsupervised algorithm for identification and fast prediction of behaviors. Nat Commun 12, 5188 (2021). https://doi.org/10.1038/s41467-021-25420-x

### 2022-07-13 scripts/study_case_taeniopygia.Rmd
first commit

@@
 
-```{r manual selection table, eval = FALSE, echo = FALSE}
-
-#Read in manual selections made in Raven
-
-mnl_sel_tbl <- imp_raven(path = raw_data_path, name.from.file = TRUE, ext.case = "lower", all.data = TRUE)
-
-glimpse(mnl_sel_tbl)
-
-mnl_sel_tbl <- mnl_sel_tbl %>% 
-  dplyr::rename(
-    `start` = "Begin Time (s)",
-    `end` = "End Time (s)",
-    `selec` = "Selection",
-    `selection_notes` = "Annotation"
-  ) %>% 
-  dplyr::mutate(
-    bottom.freq = 0.5,
-    top.freq = 10
-  ) %>% 
-  dplyr::select(sound.files, selec, start, end, selec.file, bottom.freq, top.freq, selection_notes)
-
-glimpse(mnl_sel_tbl)
-
-# Calculate signal to noise ratio (SNR)
-mnl_sel_tbl_snr <- sig2noise(mnl_sel_tbl, mar = 0.001, type = 1, in.dB = TRUE, before = FALSE, lim.dB = TRUE, bp = c(unique(mnl_sel_tbl$bottom.freq), unique(mnl_sel_tbl$top.freq)), wl = 10, path = raw_data_path)
-
-glimpse(mnl_sel_tbl_snr)
-
-# Write out the combined manual selections out as a selection table (.csv)
-write.csv(mnl_sel_tbl_snr, file.path(preproc_path, "manual_selections_Taeniopygia.csv"), row.names = FALSE)
-
-```
-
-```{r manual extended selection table, eval = FALSE, echo = FALSE}
-#Write out the combined manual selections out as an extended selection table (RDS).
-
-est <- selection_table(mnl_sel_tbl_snr, whole.recs = FALSE, extended = TRUE, mar = 0.01, parallel = cores, confirm.extended = FALSE, path = raw_data_path)
-
-glimpse(est)
-class(est)
-
-# Update sound file names, which now have "_number" after the ".wav" extension
-wav_nms <- names(attr(est, "wave.objects"))
-wav_nms <- paste(gsub(".wav", "", wav_nms), ".wav", sep = "")
-head(wav_nms) # looks good
-
-est2 <- rename_est_waves(est, wav_nms)
-
-# Looks good
-head(names(attr(est2, "wave.objects")))
-glimpse(est2)
-
-# Remove the "old.sound.file.name" column
-est2 <- est2[, -grep("old.sound.file.name", names(est2))]
-glimpse(est2)
-
-# Add in original sound file name, since each sound.file value is currently a unique name per row of the table
-check_df <- attr(est2, "check.res")
-glimpse(check_df)
-head(check_df$sound.files)
-
-# Are all sampling rates and bit rates the same?
-# Yes: 44.1 kHz sampling rate and bit rate of 16
-check_df %>%
-  distinct(sample.rate, bits)
-
-# Does the order of selections match between the EST and this attribute data frame? If can, can do the column addition below of original sound file name, start and end coordinates, selection duration
-all(est2$sound.files == check_df$sound.files)
-
-est2$duration <- check_df$duration
-est2$original_sound_files <- check_df$orig.sound.files
-est2$original_start <- check_df$orig.start
-est2$original_end <- check_df$orig.end
-
-# Looks good
-glimpse(est2)
-class(est2)
-length(attr(est2, "wave.objects"))
-
-saveRDS(est2, file.path(preproc_path, "manual_selections_Taeniopygia_est.RDS"))
-
-```
-
-```{r automated detection parameters, eval = FALSE, echo = FALSE}
-
-est <- readRDS(file.path(preproc_path, "manual_selections_Taeniopygia_est.RDS"))
-# glimpse(est)
-
-mnl_sel_tbl <- read.csv(file.path(preproc_path, "manual_selections_Taeniopygia.csv"))
-# glimpse(mnl_sel_tbl)
-
-# Use ohun::feature_reference to get parameters that can help automated detection
-# ?ohun::feature_reference
-
-# I get an error when using the EST: "# Error in warbleR::gaps(X = reference, pb = FALSE) : 
-  # extended selection tables must be created 'by.song' to be used in song.param()"
-# feat_ref <- ohun::feature_reference(est, by.sound.file = FALSE,)
-
-feat_ref <- ohun::feature_reference(mnl_sel_tbl, by.sound.file = FALSE)
-
-# Min and max duration of the manual selections
-# Use these for automated detection 
-min_dur <- floor(feat_ref[["sel.duration", "min"]]) 
-max_dur <- ceiling(feat_ref[["sel.duration", "max"]] )
-min_gap <- floor(feat_ref[["gap.duration", "min"]])
-bp <- c(feat_ref[["bottom.freq", "min"]] - 0.5, feat_ref[["top.freq", "max"]] + 0.5)
-
-min_dur
-max_dur
-min_gap
-bp
-
-# A single file for testing
-wav <- "R39Y212_43533.30691176_3_9_8_31_31.wav"
-
-```
-
 &nbsp;
@@
 
+high_snr$bottom.freq <- -1
+high_snr$top.freq <- 12
+
 # create catalogs
-catalog(X = high_snr, flim = c(0, 11), nrow = 6, ncol = 3, ovlp = 90, height = 15, width = 20, same.time.scale = FALSE, mar = 0.01, wl = 512, gr = FALSE, spec.mar = 0.4, lab.mar = 0.8, rm.axes = TRUE, by.row = TRUE, box = TRUE, pal = viridis, parallel = 10, collevels = seq(-100, 0, 5), path = data_path)
+catalog(X = high_snr, flim = c(0, 11), nrow = 6, ncol = 3, ovlp = 90, height = 15, width = 20, same.time.scale = TRUE, mar = 0.01, wl = 512, gr = FALSE, spec.mar = 0, lab.mar = 0.001, rm.axes = TRUE, by.row = TRUE, box = TRUE, pal = viridis, parallel = 10, collevels = seq(-140, 0, 5), path = data_path, img.prefix = "same.time", highlight = TRUE, alpha = 0.2)
 
@@
 ### Split reference annotations for training and testing detection
-```{r test and train, eval = TRUE, echo = TRUE}
+```{r trains set, eval = FALSE, echo = FALSE}
 
+# train set must be this one
 train_files <- c(
@@
 )
+
+```
+
+```{r test and train, eval = TRUE, echo = TRUE}
+
+set.seed(450)    
+train_files <- sample(unique(manual_ref$sound.files), 3)
+
 test_files <- setdiff(manual_ref$sound.files, train_files)
@@
 ```{r, eval = FALSE}
+
 # subset with highest performance
@@
 ### Run detection on test data
-```{r, detect on all data, eval = TRUE, echo = TRUE}
+```{r, detect on all data, eval = FALSE, echo = TRUE}
 
-opt_det_all <- energy_detector(
+det_test <- energy_detector(
   files = test_files, 
@@
 
+```{r , eval = FALSE, echo = FALSE}
+
+write.csv(det_test, file.path(preproc_path, "detection_test_data.csv"), row.names = FALSE)
+
+```
+
 &nbsp;
@@
 ### Evaluate detection performance
+
+```{r , eval = TRUE, echo = FALSE}
+
+det_test <- read.csv(file.path(preproc_path, "detection_test_data.csv"))
+
+```
+
 ```{r, diagnose, eval = TRUE, echo = TRUE}
 
-diagnose_detection(reference = test_ref, detection = opt_det_all, by.sound.file = FALSE)
+diagnose_detection(reference = test_ref, detection = det_test, by.sound.file = FALSE)
 

### 2022-07-13 scripts/study_case_thyroptera.Rmd
first commit

@@
 data_path <- "./data/raw/thyroptera/"
+
+source("~/Dropbox/R_package_testing/ohun/R/get_templates.R")
 ```
@@
 ```{r get sample of recordings from flight coordination paper, eval = FALSE, echo = FALSE}
+
 fixed_calls <- read.csv( "./data/processed/manually_fixed_call_detections.csv")
@@
 
-manual_ref <- read.csv(file.path(data_path, "manual_annotations_thyroptera.csv"), stringsAsFactors = FALSE)
+manual_ref <- read.csv(file.path(data_path, "manual_annotations_thyroptera.csv"))
+
 ```
@@
 # create catalogs
-catalog(X = high_snr, flim = c(10, 45), nrow = 6, ncol = 5, ovlp = 90, height = 15, width = 20, same.time.scale = TRUE, mar = 0.005, wl = 512, gr = FALSE, spec.mar = 0.4, lab.mar = 0.8, rm.axes = TRUE, by.row = TRUE, box = TRUE, pal = viridis, parallel = 10, collevels = seq(-100, 0, 5))
+catalog(X = high_snr, flim = c(10, 45), nrow = 6, ncol = 5, ovlp = 90, height = 10, width = 20, same.time.scale = TRUE, mar = 0.005, wl = 512, gr = FALSE, spec.mar = 0, lab.mar = 0.001, rm.axes = TRUE, by.row = TRUE, box = FALSE, pal = viridis, collevels = seq(-100, 0, 5), labels = NULL)
+
+```
+
+```{r fig for paper, eval = FALSE, echo = FALSE}
+
+catalog(X = high_snr[1:15, ], flim = c(10, 45), nrow = 3, ncol = 5, ovlp = 90, height = 7.5, width = 20, same.time.scale = TRUE, mar = 0.005, wl = 512, gr = FALSE, spec.mar = 0.4, lab.mar = 0.2, rm.axes = TRUE, by.row = TRUE, box = TRUE, pal = viridis, parallel = 1, collevels = seq(-100, 0, 5), img.prefix = "thyroptera_15_calls", labels = NULL)
 
@@
 
-train_files <- c(
- "ch1MPI2020-01-17_08-44-19_0000028.wav",
- "ch1MPI2020-01-23_10-30-35_0000106.wav",
- "ch1MPI_2020-01-30_15-06-38_0000193.wav",
- "ch1MPI2020-01-16_11-46-10_0000018.wav",
- "ch1MPI2020-01-16_13-40-58_0000019.wav",
- "ch1MPI2020-01-24_11-28-15_0000124.wav"
-)
+set.seed(1)
+
+train_files <- sample(unique(manual_ref$sound.files), size = 5)
+
 test_files <- setdiff(manual_ref$sound.files, train_files)
@@
 ### Get diverse set of templates
-```{r cross-correlation detection, eval = FALSE, echo = TRUE}
+```{r get templates, eval = TRUE, echo = TRUE, fig.height=7}
+# find templates
+templates <- get_templates(train_ref, path = data_path,  bp = c(10, 50), fast = TRUE, ovlp = , wl = 200, n.sub.spaces = 3)
+```
 
-# find templates
-templates <- get_templates(train_ref, path = data_path,  bp = c(10, 50), fast = TRUE, ovlp = , wl = 200, n.templates = 4)
-
+```{r get templates 2, eval = FALSE, echo = TRUE}
 # create ext. selection table
@@
 
-templates_est <- rename_est_waves(templates_est, new.sound.files = templates$distance.quantile)
+templates_est <- rename_est_waves(templates_est, new.sound.files = templates$template)
 
@@
 
-corr_templ <- template_correlator(templates = templates_est,
-    path = data_path, files = unique(train_ref$sound.files), parallel = paral, hop.size = 10, ovlp = 70)
-
 corr_templ_train <- template_correlator(
@@
   template.correlations = corr_templ_train,
-  threshold = seq(0.05, 0.5, 0.01)
+  threshold = seq(0.1, 0.9, 0.05)
   )
@@
 ```{r, eval = FALSE, echo = FALSE}
+
 templates_est <- readRDS("./data/processed/templates_thyroptera_est.RDS")
@@
   template.correlations = corr_templ_train,
-  threshold = seq(0.05, 0.5, 0.01), 
+  threshold = seq(0.1, 0.9, 0.05), 
   parallel = paral
@@
 
-opt_detec_train$templates <- gsub("-1", "", opt_detec_train$templates)
+opt_detec_train$templates <- gsub("-1$", "", opt_detec_train$templates)
 
@@
 
-### Create recall-precision plot
-```{r, eval = FALSE}
-
-# summarize across sound files
-diag_train <- summarize_diagnostic(opt_detec_train, time.diagnostics = FALSE)
-agg_sd <- aggregate(cbind(recall, precision) ~ threshold + templates, data = opt_detec_train,
-    mean)
-
-# plot
-ggplot(agg_sd, aes(x = recall, y = precision, group = templates, color = templates)) +
-    geom_line() + geom_point() + scale_color_viridis_d(end = 1) + theme_classic()
-```
-
+### Create plot with threshold vs f1 score
 ```{r, eval = TRUE, echo = FALSE}
@@
 
-diag_train <- summarize_diagnostic(opt_detec_train, time.diagnostics = FALSE)
-
-# names(diag_train)
-
-agg_sd <- aggregate(cbind(recall, precision) ~ threshold + templates, data = diag_train,
+agg_f1 <- aggregate(f1.score ~ threshold + templates, data = opt_detec_train,
 mean)
 
-ggplot(agg_sd, aes(x = recall, y = precision, group = templates, color = templates)) +
+ggplot(agg_f1, aes(x = threshold, y = f1.score, group = templates, color = templates)) +
 geom_line() + geom_point() + scale_color_viridis_d(end = 1) + theme_classic()
@@
 
-sub_diag_train <-diag_train[order(diag_train$f1.score, decreasing = TRUE), ]
+opt_detec_train <- opt_detec_train[order(opt_detec_train$f1.score, decreasing = TRUE), ]
 
-sub_diag_train <- sub_diag_train[1:10, c("templates","threshold", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f1.score")]
+sub_opt_detec_train <- opt_detec_train[1:10, c("templates","threshold", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f1.score")]
 
-sub_diag_train
+sub_opt_detec_train
 ```
@@
 # subset with highest performance
-sub_diag_train <-diag_train[order(diag_train$f1.score, decreasing = TRUE), ]
+opt_detec_train <- opt_detec_train[order(opt_detec_train$f1.score, decreasing = TRUE), ]
 
-sub_diag_train <- sub_diag_train[1:10, c("templates","threshold", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f1.score")]
+sub_opt_detec_train <- opt_detec_train[1:10, c("templates","threshold", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f1.score")]
 
 # print dynamic table
-oa_DT <- datatable(sub_diag_train, editable = list(target = "row"), rownames = FALSE, style = "bootstrap",
+oa_DT <- datatable(sub_opt_detec_train, editable = list(target = "row"), rownames = FALSE, style = "bootstrap",
     filter = "top", options = list(pageLength = 100, autoWidth = TRUE, dom = "ft"),
@@
 
-formatRound(table = oa_DT, columns = sapply(sub_diag_train, is.numeric), 3)
+formatRound(table = oa_DT, columns = sapply(sub_opt_detec_train, is.numeric), 3)
+
 ```
@@
 ### Run detection on test files
-```{r, detect on all data, eval = FALSE, echo = TRUE}
+
+We used a threshold of 0.45 as te centroid template seems to behave well across thresholds.
+```{r, detect on all data 67, eval = FALSE, echo = TRUE}
 
@@
 corr_templ_test <- template_correlator(
-  templates = templates_est[templates_est$sound.file == "50%", ],
+  templates = templates_est[templates_est$sound.file == "centroid", ],
     path = data_path, files = unique(test_ref$sound.files), 
-  parallel = paral, 
   hop.size = 10, 
@@
 
-# extract best template-threshold combination
-opt_50 <- optimize_fourier_detec[optimize_fourier_detec$templates == "50%", ]
-opt_thresh <- opt_50$threshold[which.max(opt_50$f1.score)]
-
 # detect on test files
@@
   template.correlations = corr_templ_test,
-    threshold = opt_thresh
+    threshold = 0.45
   )
+
 ```
@@
 ### Evaluate detection performance on test files
-```{r, diagnose detect on all data, eval = FALSE, echo = TRUE}
+```{r, diagnose detect on all data 67, eval = FALSE, echo = TRUE}
 
@@
 ```{r, eval = FALSE, echo = FALSE}
-saveRDS(detec_test, "./data/processed/detection_thyroptera_all.RDS")
+
+saveRDS(detec_test, "./data/processed/detection_thyroptera_all_67%_template.RDS")
+
 ```
@@
 
-detec_test <- readRDS("./data/processed/detection_thyroptera_all.RDS")
+detec_test <- readRDS("./data/processed/detection_thyroptera_all_67%_template.RDS")
 
-diagnose_detection(reference = test_ref, detection = detec_test)
+dd <-diagnose_detection(reference = test_ref, detection = detec_test)
+
+dd
+
+dd$f1.score
+
 ```
@@
 
-- Good performance on detecting Spix's disc-winged social calls: F1 score was 0.99 for the training data set and 0.94 for the testing data
+- Good performance on detecting Spix's disc-winged social calls: F1 score was `r round(dd$f1.score, 3)` for the training data set and `r round(max(opt_detec_train$f1.score), 3)` for the testing data
+
+- Some merged positives were found but notice that there are some overlapping signals in the reference annotation as well (can be checked with `warbleR::overlapping_sels(manual_ref)`)
 

### 2022-07-13 manuscript/manuscript.Rmd
manually change docx

@@
 output:
+  word_document:
+    toc: yes
+    toc_depth: '3'
   html_document:
@@
       smooth_scroll: no
-  word_document:
-    toc: yes
-    toc_depth: '3'
   pdf_document:
@@
 lots of recordings), which can be challenge to analyzed manually by human observers.
-As a result, a growing number of computational tools for the analysis of acoustic features on those signals is increasingly available (reviewd by @Stowell2022), reflecting the need of automated approaches for efficiently conducting these analyses.
+As a result, a growing number of computational tools for the analysis of acoustic features on those signals is increasingly available (reviewed by @Stowell2022), reflecting the need of automated approaches for efficiently conducting these analyses.
 

### 2022-07-13 README.Rmd
Update README.Rmd

@@
 title: ohun_paper_analyses
-author: AUTOR NAME
+author: Marcelo Araya Salas
 output: github_document
@@
 <!-- Short Description  -->
-This is a short description of the project
+Data analysis and manuscript writting for method paper on the R package ohun
 
@@
 * [General info](#general-info)
-  * [Analyses](#Analyses)
     * [To do list](#to-do-list)
@@
 
-General information about the project (study system, goals, etc)
-
-## Analysis
-
-* Analysis 1
+Data analysis and manuscript writting for method paper on the R package ohun
 
@@
 
-*  Something not finished yet
+*  Writting
 
@@
 
-* Do the next analysis
+* Add citations
 
@@
 ## Contact
-Created by [Autor name](website_URL)
+Created by [Marcelo Araya-Salas](marceloarayasalas.weebly.com/)

### 2022-07-13 README.Rmd
update readme

@@
 title: ohun_paper_analyses
-author: Marcelo Araya Salas
+author: "Marcelo Araya Salas"
 output: github_document
-date: `r format(Sys.Date(), "%d-%m-%Y")`
 editor_options:

### 2022-07-13 manuscript/manuscript.Rmd
update manuscript

--- 2022-07-13 manuscript/manuscript.Rmd
@@
 
-```{r}
-feature_reference(test_ref)
-```
-
 ```{r  optimize automated detection, eval = FALSE, echo = TRUE}

### 2022-07-27 manuscript/manuscript.Rmd
update ohun paper

@@
 ---
-title: 'ohun: an R package for optimizing automatic acoustic signal detection'
+title: 'ohun: an R package for diagnosing and optimizing automatic acoustic signal detection'
 author:
@@
 always_allow_html: yes
-bibliography: combined_bibs.bib
-citation-style: apa.csl
+bibliography: references.bib
+citation-style: methods-in-ecology-and-evolution.csl
 link-citations: yes
@@
 
-pkgs <- c("kableExtra", "knitr", "rprojroot", "sciRmdTheme", "rmdwc", "whereami", "maRce10/ohun", "ggplot2", "viridis", "rmdwc")
+pkgs <- c("kableExtra", "knitr", "rprojroot", "sciRmdTheme", "rmdwc", "whereami", "maRce10/ohun", "ggplot2", "viridis", "rmdwc", "RefManageR", "bibtex")
 
@@
 if (is(rootdir, "try-error")) rootdir <-  ".."
+
 opts_knit$set(root.dir = rootdir)
@@
 
-```{r, fixing citation .bib files, eval = TRUE, echo = FALSE}
+```{r, fixing citation .bib files, eval = TRUE, echo = FALSE, message=FALSE, warning=FALSE}
 
-# how to cite in Rmarkdown
+# how to cite in Rmarkdown:
 # https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html
+
+# find citation style files (csl) here:
+# https://github.com/stackpr/csl-styles
 
@@
 # Alternatively authors ca
-personal_libraries <- c(author_1 = "/home/m/Documents/mendeley_libraries/ohun_paper.bib")
+personal_libraries <- c(author_1 = "/home/m/Documents/mendeley_libraries/ohun_paper.bib", author_2 = "/home/m/Documentos/mendeley_bibtex_files/ohun_paper.bib")
 
-# update bibtex library
-for (i in 1:length(personal_libraries))
-if (file.exists(personal_libraries[i]))
- file.copy(from = personal_libraries[i], to = file.path(dirname(whereami::thisfile()), paste0(names(personal_libraries)[i], ".bib")), overwrite = TRUE)
+# location of current Rmarkdown
+rmd_dir <- dirname(whereami::thisfile()) 
+if (is.null(rmd_dir)) rmd_dir <- file.path(rootdir, "manuscript")
 
-## combine .bib files in compiled.bib ###
-# list bib files
-bibs <- list.files(pattern = ".bib$", path = dirname(whereami::thisfile()), full.names = TRUE)
+# read personal library
+personal_libraries <- personal_libraries[file.exists(personal_libraries)]
+personal_bib <- suppressMessages(ReadBib(personal_libraries))
 
-# remove combined from pool
-bibs <- bibs[basename(bibs) != "combined_bibs.bib"]
+# read paper's references.bib if it exists
+if (file.exists(file.path(rootdir, "references.bib"))){
+    combined_bib <- suppressMessages(ReadBib(file.path(rmd_dir, "references.bib")))
 
-#  read in a list
-if (length(bibs) > 0){
-  combined_bibs_l <- lapply(bibs, readLines)
-  names(combined_bibs_l) <- gsub("\\.bib$", "", basename(bibs))
+    merge_bib <- suppressMessages(merge(personal_bib, combined_bib, fields.to.check = "title", ignore.case = TRUE))
+} else
+    merge_bib <- personal_bib
 
-# combine bibs in a single one
-combined_bibs <- unlist(combined_bibs_l)
-
-# save pool libraries as combined_bibs.bib
-writeLines(text = combined_bibs, file.path(dirname(whereami::thisfile()), "combined_bibs.bib"))
-
-# check if there are duplicated keys
-keys <- grep("@article{", combined_bibs, fixed = TRUE, value = TRUE)
-keys <- gsub("@article{", "", keys, fixed = TRUE)
-keys <- gsub(",", "", keys, fixed = TRUE)
-
-tab_keys <- table(keys)
-
-if (anyDuplicated(keys)){
-  print(paste0(sum(tab_keys > 1), " duplicate(s) references found in combined_bibs.bib"))
-}
-} else writeLines(text = "empty", file.path(dirname(whereami::thisfile()), "combined_bibs.bib"))
+# save combined references
+WriteBib(bib = merge_bib, file = file.path(rmd_dir, "references.bib"), biblatex = TRUE)
 
@@
 Grace Smith-Vidaurre ^4^,
-Gloriana Chaverri ^3^,
+Gloriana Chaverri ^3^, ^5^ 
 Fabiola Chirino ^1^ &
-Alejandro Rico-Guevara ^5^
+Alejandro Rico-Guevara ^6^
 
@@
 
-^5^  Department of Biology, University of Washington, Seattle, USA
+^5^  Smithsonian Tropical Research Institute, Panama City, Panamá
+
+^6^  Department of Biology, University of Washington, Seattle, USA
 
@@
 
-### Keywords:
+**Keywords**: bioacoustics, automatic detection, animal vocalizations
 
@@
 This is partly due to the relative ease in which they can be registered for wide range of taxonomic groups and research settings.
-However, bioacoustic research can easily generate large amounts of data which might proof challenging to analyze in a timely manner.
+However, bioacoustic research can easily generate large amounts of data which might prove challenging to analyze in a timely manner.
 Many tools for the automatic detection of signals are currently available.
@@
 This can be done with routine outputs imported from other software and from detection run within the package itself.
-The package also provides a set functions to organize acoustic data sets in a format amenable for detection analyses.
-It also includes an implementation of two automatic detection methods commonly used in bioacustic analysis: energy-based detection and template-based detection.
-We included examples on how to evalute detections from external software and showcase the package usage with study cases on zebra-finch (*Taenopygia gutata*) songs and Spix's disc-winged bat (*Thyroptera tricolor*) ultrasonic social calls.
+The package also provides a set of functions to organize acoustic data sets in a format amenable for detection analyses.
+It also includes an implementation of two automatic detection methods commonly used in bioacoustic analysis: energy-based detection and template-based detection.
+We included examples on how to evaluate detections from external software and showcase the package usage with study cases on zebra-finch (*Taenopygia gutata*) songs and Spix's disc-winged bat (*Thyroptera tricolor*) ultrasonic social calls.
 Finally, we provide some general suggestions to improve detection performance.
@@
 The wide usage of animal sounds in research is partly due to the fact that they can be easily registered using non-intrusive methods, can be obtained in a variety of settings from laboratories to natural areas, and the equipment required for registering an analyzing these signals has become increasingly inexpensive.
-In addition, the existence of online repositories and growing number has facilitated the study of these traits at larger taxonomic and geographic scales.
+In addition, the existence of online repositories has facilitated the study of these traits at larger taxonomic and geographic scales.
 However, the adoption of bioacustic approaches in research can also imply large amounts of data (i.e.
-lots of recordings), which can be challenge to analyzed manually by human observers.
+lots of recordings), which can be challenging to analyzed manually.
 As a result, a growing number of computational tools for the analysis of acoustic features on those signals is increasingly available (reviewed by @Stowell2022), reflecting the need of automated approaches for efficiently conducting these analyses.
@@
 In this regard, the use of standard approaches for evaluating the performance of automatic detection tools might prove helpful to inform the decision about which method better fit our study system.
-The performance of automatic acoustic signal detection routines have been evaluated using standard indices from signal detection theory (CITATIONS).
+The performance of automatic acoustic signal detection routines has been evaluated using standard indices from signal detection theory (CITATIONS).
 In its basic form, performance is assessed by comparing the output of a detection routine against a "gold standard" reference in which all the target signals have been annotated.
@@
 For instance, the same signal can be detected as several separated signals, the inferred time position can be far off from the the target signal position, or several signals can be detected as a single one.
-Therefore, metrics that account for this other performance dimensions can be a valuable tool to properly diagnose automatic acoustic signal detection.
+Therefore, metrics that account for these other performance dimensions can be a valuable tool to properly diagnose automatic acoustic signal detection.
 
@@
 This can be done with routine outputs imported from other software as well as detection run within the package itself.
-The package also provides a set functions to explore acoustic data sets and organize them in a format amenable for detection analyses.
-In addition, it offers implementations of two automatic detection methods commonly used in bioacustic analysis: energy-based detection and template-based detection.
-We explain how to explore and format acoustic data sets, how acoustic signal detection routines can be evaluate and then showcase the package usage with study cases on zebra-finch songs (*Taenopygia gutata*) and Spix's disc-winged bats (*Thyroptera tricolor*) which represent different recording settings (lab and flight cages) and signal types (sonic mating signals and ultrasonic social calls).
+The package also provides a set of functions to explore acoustic data sets and organize them in a format amenable for detection analyses.
+In addition, it offers implementations of two automatic detection methods commonly used in bioacoustic analysis: energy-based detection and template-based detection.
+We explain how to explore and format acoustic data sets, how acoustic signal detection routines can be evaluated and then showcase the package usage with study cases on zebra-finch songs (*Taenopygia gutata*) and Spix's disc-winged bats (*Thyroptera tricolor*) which represent different recording settings (lab and flight cages) and signal types (sonic mating signals and ultrasonic social calls).
 
@@
 time window size is affected by sampling rate) or simply because some software might can only work on specific sound file formats.
-In addition, long sound files might be hard to analyzed in regular computers and might have to be split in shorter clips.
+In addition, long sound files might be hard to analyze in regular computers and might have to be split in shorter clips.
 In the latter case the function `split_acoustic_data` can be used to produce those clips:
@@
 The right balance between these two extremes will be given by the relative costs of missing signals and mistaking noise for signals, given the specific goals of the study.
-These indices provide an useful framework for diagnosing and optimizing the performance of a detection routine.
+These indices provide a useful framework for diagnosing and optimizing the performance of a detection routine.
 
@@
 To accomplish this, annotations derived from a detection routine are compared against a reference table containing the time position of all target signals in the sound files.
-For instance, the following code evaluates a routine run in Raven Pro 1.6 (XXXX) using the "band limited energy detector" option (minimum frequency: 0.8 kHz; maximum frequency: 22 kHz; minimum duration: 0.54989 s; maximum duration: 0.54989s; minimum separation: 0.02268 s) on a subset of the Zebra finch recordings described below:
+For instance, the following code evaluates a routine run in Raven Pro 1.6 (Cornell Lab of Ornithology) using the "band limited energy detector" option (minimum frequency: 0.8 kHz; maximum frequency: 22 kHz; minimum duration: 0.03968 s; maximum duration: 0.54989s; minimum separation: 0.02268 s) on a subset of the Zebra finch recordings described below:
 
-```{r, eval = TRUE, echo = FALSE}
+```{r read and save raven automatic detection selection files, eval = FALSE, echo = FALSE}
 
@@
 
-raven_detec <- imp_raven(path = file.path(rootdir, "/data/raw/"), files = "Raven.Band.Limited.Energy.Detector.selections.txt", warbler.format = TRUE)
+raven_detec <- imp_raven(path = file.path(rootdir, "/data/raw/"), files = c("22.kHz_sep.0.02322.selections.txt.txt", "15.kHz_sep.0.02322.selections.txt", "10.kHz_sep.0.02322.selections.txt", "10.kHz_sep.0.005.selections.txt"), warbler.format = TRUE)
 
-# raven_ref <- manual_ref[manual_ref$sound.files %in% raven_detec$sound.files, ]
+raven_detec$tuning_parameters <- gsub(".selections.txt", "", raven_detec$selec.file)
+raven_detec$tuning_parameters <- gsub("kHz_", "kHz; ", raven_detec$tuning_parameters)
+raven_detec$tuning_parameters <- gsub(".kHz", " kHz", raven_detec$tuning_parameters)
+raven_detec$tuning_parameters <- gsub("sep.", "sep: ", raven_detec$tuning_parameters)
+raven_detec$tuning_parameters <- paste0("band: 1-", raven_detec$tuning_parameters, " s")
+
+write.csv(raven_detec, "./data/processed/combined_raven_detection.csv", row.names = FALSE)
+
+```
+
+```{r, eval = FALSE, echo = TRUE}
+
+raven_detec <- read.csv("combined_raven_detection.csv")
+
+diagnose_detection(reference = manual_ref_tae, detection = raven_detec, by = "tuning_parameters")
 
@@
 
-diagnose_detection(reference = manual_ref_tae, detection = raven_detec)
+raven_detec <- read.csv("./data/processed/combined_raven_detection.csv")
+
+diagnose_detection(reference = manual_ref_tae, detection = raven_detec, by = "tuning_parameters")
 
@@
 
-The output shows the indices described above. The function also allows to detail those indices by sound file. Here we show the first 6 files:
+The output shows the indices described above. The function also allows to detail those indices by sound file. Here we show the first 10 files:
 
@@
 
-diag_raven <- diagnose_detection(reference = manual_ref_tae, detection = raven_detec, by.sound.file = TRUE)
+diag_raven <- diagnose_detection(reference = manual_ref_tae, detection = raven_detec, by = "tuning_parameters", by.sound.file = TRUE)
 
-head(diag_raven)
+head(diag_raven, 10)
 ```
@@
 Diagnostics from routines using different tuning parameters can be used to identify the parameter values that optimize a detection.
-The process of evaluating different routines for detection optimization is incorporated into the two signal detection approaches provided natively by ohun, which we depict in the following section.
+These process of evaluating different routines for detection optimization is incorporated into the two signal detection approaches provided natively by ohun, which we depict in the following section.
 Note that the detection with Raven Pro does not necessarily reflect the best performance of this software and has been included only as an example on evaluating detection from external sources rather than a direct comparison of performance to ohun.
@@
 These methods are better suited for highly stereotyped or good signal-to-noise ratio (SNR) signals, respectively.
-If the target signals do not fit these requirements, more elaborated methods (i.e.
+If the target signals do not fit these requirements, more elaborate methods (i.e.
 machine/deep learning approaches) are warranted.
@@
 
-We recorded `r length(unique(manual_ref_thy$sound.files))` individuals of Spix's disc-winged bats at Baru Biological Station, in south-western Costa Rica in January 2020.
+We recorded `r length(unique(manual_ref_thy$sound.files))` individuals of Spix's disc-winged bats (*Thyroptera tricolor*) at Baru Biological Station, in south-western Costa Rica in January 2020.
 Bats were captured at their roosting sites (furled leaves of Zingeberaceae plants).
@@
 
-Recordings were manually annotated using Raven Pro 1.6 (XXXX).
+Recordings were manually annotated using Raven Pro 1.6 (Cornell Lab of Ornithology).
 Annotations were created by visual inspection of spectrograms, in which the start and end of signals were determined by the location of the continuous traces of power spectral entropy of the target signals.
 A total of `r nrow(manual_ref_thy)` calls were annotated (~`r round(nrow(manual_ref_thy) / 30, 0)` calls per recording).
-Annotations were made with a time window of 200 samples and 70% of overlap.
-Annotations were then imported into R using the package Rraven (XXXX).
+Annotations were made with a time window of 200 samples and 70% of overlap, and were then imported into R using the package Rraven (@Araya-Salas2017e).
 
-Inquiry calls of Spix's disc-winged bats are structurally stereotyped.
-Most variation is found between individuals although the basic form of a short, downward broadband frequency modulation is always shared (Fig.
+Inquiry calls of Spix's disc-winged bats are structurally stereotyped (@Chaverri2010b).
+Most variation is found between individuals, although the basic form of a short, downward broadband frequency modulation is always shared (Fig.
 BAT-SPECTRO, Araya-Salas et al 2021 ontogeny).
@@
 
-# knitr::include_graphics("~/Dropbox/Projects/ohun_paper/data/raw/thyroptera/Catalog_p1.jpeg")
-knitr::include_graphics((file.path(rootdir, "data/raw/thyroptera/Catalog_p1.jpeg")))
+knitr::include_graphics(file.path(rootdir, "data/raw/thyroptera/Catalog_p1.jpeg"))
+
 ```
 
-Template-based detection is a useful approach when there is little structural differences in the target signals.
+Template-based detection is a useful approach when there are few structural differences in the target signals.
 We used this approach in ohun to detect inquiry calls.
@@
 ACOUSTIC-SPACE.
-Acoustic space defined as the the first two component of a Principal Component Analysis on spectrographic parameters.
+Acoustic space defined as the the first two components of a Principal Component Analysis on spectrographic parameters.
 Templates are selected as those closer to the centroid for each of the sub-spaces.
@@
 
-The last line of codes evalutes the detection on the test data set, which shows a good performance for both recall and precision.
+The last line of codes evaluates the detection on the test data set, which shows a good performance for both recall and precision.
 
-### Energy-based detection on zebra finch songs
+### Energy-based detection on zebra finch vocalizations
 
-We used recordings from 18 zebra finch males. These recordings were obtained ....
-
+We used recordings from 18 zebra finch males recorded at obtained from the Rockefeller University Field Research Center Song Library (http://ofer.sci.ccny.cuny.edu/songs, @Tchernichovski2021).
+Recordings contain undirected vocalizations (songs or calls) of single males recorded in sound attenuation chambers using Sound Analysis Pro.
 Zebra finch vocalizations are composed by multiple elements (i.e.
@@
 
-knitr::include_graphics((file.path(rootdir, "data/raw/taeniopygia/same.time-Catalog_p1.jpeg")))
+knitr::include_graphics(file.path(rootdir, "data/raw/taeniopygia/same.time-Catalog_p1.jpeg"))
 
@@
 
+Reference annotations were made manually on the oscillogram with the spectrogram and audio as a guide using Raven Lite 2.0.1 (Cornell Lab of Ornithology).
 The following code loads the reference annotations and split them in two data set for training (3 sound files) and testing (15 sound files):
@@
 
-The output (`opt_det_train`) shows the performance indices for each of those combinations, here we show the 10 combinations with the highest F1 score:
+The output (`opt_det_train`) shows the performance indices for each of those combinations. Hhere we show the 10 combinations with the highest F1 score:
 
@@
 The performance on the test data set was also acceptable with a F1 score of 0.95.
-Note that in the example we use a small subset of sound files for training.
+Note that in the example we used a small subset of sound files for training.
 More training data might be needed for optimizing a detection routine on larger data sets or recordings with more variable signals or background noise levels.
@@
 The package offers additional tools to simplify acoustic signal detection.
-The function 'feature_reference' allows to extract information about the duration and frequency range of acoustic signals than can then be used to define tuning parameter values.
-This can be particularly useful for energy-based detection in which time and frequency attributes of the target signals are required.
 Detected signals can be labeled as false or true positives with the function 'label_detection'.
@@
 
-Here we have showed how to evaluate the performance of acoustic signal detection routines using the package ohun.
+Here we have shown how to evaluate the performance of acoustic signal detection routines using the package ohun.
 The package can evaluate detection outputs imported from other software, as well as its own detection routines.
@@
 Note that the tools offered by ohun for diagnosing a detection should not be necessarily limited to acoustic data.
-Any type of automatic detection in which the time of occurrence of discrete events needs to be found can be evaluated and optimized by comparing it to a reference annotation.
-For instance detection of specific behaviors in video analysis of animal motor activity (Sturman et al 2020; Hsu & Yttri 2021; DeepEthogram).
+Any type of automatic detection in which the time of occurrence of discrete events needs to be found can be evaluated and optimized by comparing it to a reference annotation, for instance, detection of specific behaviors in video analysis of animal motor activity (Sturman et al 2020; Hsu & Yttri 2021; DeepEthogram).
+The package can also be used to assess the repeatability of manually made annotations by comparing annotations from different observers on the same data sets.
 
@@
 However, these methods have been widely used by the bioacoustic community and, under the appropriate conditions can reach adequate performance, as we have shown in our two study cases.
-Deep learning methods tend to require larger computational power and more complex training routines.
-This might bring unnecessary difficulties when dealing when less challenging detection tasks.
+Deep learning methods tend to require larger computational power, larger training data sets and, in some cases, more complex training routines.
+This might bring unnecessary difficulties when dealing with when less challenging detection tasks.
 Therefore the availability of a wide range of approaches is desired in order to simplify finding the most appropriate tool for the intricacies of our study system and research goals.
-Note that the tools offered in ohun can be used in a pipeline in which detected signals are further classified using more elaborated discrimination algortihms.
-Acoustic parameters can be used to quantify the structure of the signals and tell apart target from non-target signals.
+Note that the tools offered in ohun can be used in a pipeline in which detected signals are further classified using more elaborated discrimination algorithms.
+Acoustic parameters can be used to quantify the structure of the signals and distinguish target from non-target signals.
 
@@
 Template-based detection is almost always faster than energy-based detection.
-4.
+3.
 Run routines in parallel.
@@
 Make sure reference tables contain all target signals and only the target signals, otherwise performance optimization can be misleading as the performance of a detection cannot be better than the reference itself.
-Lastly, avoid having overlapping signals or several signals as a single one (like a multi-syllable vocalization) in the reference table when running an energy-based detector, as they are likely to be identified as separeted units.
+Lastly, avoid having overlapping signals or several signals as a single one (like a multi-syllable vocalization) in the reference table when running an energy-based detector, as they are likely to be identified as separated units.
 
-https://cran.r-project.org/web/packages/bioacoustics/vignettes/tutorial.html
+### Acknowledgements
 
+Nazareth Rojas, Silvia Chaves-Ramírez, Mariela Sánchez-Chavarría, Miriam Gioiosa, Cristian Castillo-Salazar, and José Pablo Barrantes for their help collecting acoustic data for Spix’s disc-winged bats and Jorge Elizondo for his help on manually annotating recordings.
+We also thank the Centro Biológico Hacienda Barú for their continuous support of our research.
+This study was partly funded by a CONARE-Max Planck grant from the Consejo Nacional de Rectores.
+
+### Ethical note
+
+All sampling protocols followed guidelines approved by the American Society of Mammalogists for capture, handling and care of mammals (@Sikes2016) and the ASAB/ABS Guidelines for the use of animals in research.
+This study was conducted in accordance with the ethical standards for animal welfare of the Costa Rican Ministry of Environment and Energy, Sistema Nacional de Áreas de Conservación, permit no.
+SINAC-ACOPAC-RES-INV-008-2017 (Decree No.
+32553-MINAE).
+Protocols were also approved by the University of Costa Rica’s Institutional Animal Care and Use Committee (CICUA-42-2018).
+
+### Supplementary materials
+
+Supplementary material associated with this article is available in XXXXXXXXXXXXXXXXX
 ## References

### 2022-08-07 manuscript/manuscript.Rmd
fabi changes

@@
 Gloriana Chaverri ^3^, ^5^
+Juan Carlos Brenes ^1^ &
 Fabiola Chirino ^1^ &

### 2022-08-11 manuscript/manuscript.Rmd
added comments from authors

--- 2022-07-27 manuscript/manuscript.Rmd
@@
 ---
-title: 'ohun: an R package for diagnosing and optimizing automatic acoustic signal detection'
+title: 'ohun: an R package for diagnosing and optimizing automatic sound event detection'
 author:
@@
 tibble <- function(x, ...) { 
-  x <- kbl(x, digits=4, align= 'c', row.names = FALSE) 
+  x <- kbl(x, digits=3, align= 'c', row.names = FALSE) 
    x <- kable_styling(x, position ="center", full_width = FALSE,  bootstrap_options = c("striped", "hover", "condensed", "responsive")) 
@@
 # Alternatively authors ca
-personal_libraries <- c(author_1 = "/home/m/Documents/mendeley_libraries/ohun_paper.bib", author_2 = "/home/m/Documentos/mendeley_bibtex_files/ohun_paper.bib")
+personal_libraries <- c(author_1 = "/home/m/Documents/mendeley_libraries/ohun_paper.bib", author_2 = "/home/neuro/Documents/export.bib")
 
 # location of current Rmarkdown
-rmd_dir <- dirname(whereami::thisfile()) 
+if (!is.null(whereami::thisfile()))
+    rmd_dir <- dirname(whereami::thisfile()) else
+        rmd_dir <- NULL
 if (is.null(rmd_dir)) rmd_dir <- file.path(rootdir, "manuscript")
@@
 personal_libraries <- personal_libraries[file.exists(personal_libraries)]
+
+if (length(personal_libraries) > 0){
 personal_bib <- suppressMessages(ReadBib(personal_libraries))
@@
 WriteBib(bib = merge_bib, file = file.path(rmd_dir, "references.bib"), biblatex = TRUE)
-
+}
 ```
@@
 Marcelo Araya-Salas ^1,^ ^2,^ ^3^ *,
-Grace Smith-Vidaurre ^4^,
-Gloriana Chaverri ^3^, ^5^ 
+Grace Smith-Vidaurre ^4,^ ^5,^ ^6^, 
+Gloriana Chaverri ^3^, ^7^
+Juan Carlos Brenes ^1^ &
 Fabiola Chirino ^1^ &
-Alejandro Rico-Guevara ^6^
+Alejandro Rico-Guevara ^8,^ ^9^
 
@@
 
-^4^ GRACE'S AFFILIATION
+^4^ Laboratory of Neurogenetics of Language, Rockefeller University, New York, NY, USA
 
-^5^  Smithsonian Tropical Research Institute, Panama City, Panamá
+^5^ Rockefeller University Field Center, Millbrook, NY, USA
 
-^6^  Department of Biology, University of Washington, Seattle, USA
+^6^ Department of Biological Sciences, University of Cincinnati, Cincinnati, OH, USA
+
+^7^  Smithsonian Tropical Research Institute, Panama City, Panamá
+
+^8^  Department of Biology, University of Washington, Seattle, USA
+
+^9^ Burke Museum of Natural History and Culture, University of Washington, Seattle, USA
 
@@
 Animal acoustic signals are widely used in a diversity of research areas.
-This is partly due to the relative ease in which they can be registered for wide range of taxonomic groups and research settings.
-However, bioacoustic research can easily generate large amounts of data which might prove challenging to analyze in a timely manner.
-Many tools for the automatic detection of signals are currently available.
-However, choosing the right approach for a specific task might turn difficult as there are no available tools that can provide a common framework for evaluating detection performance.
-Here we present the new R package ohun.
-The package is intended to facilitate the automatic detection of acoustic signals, by providing functions to diagnose and optimize detection routines as well as compare the performance of different detection approaches.
-The package makes use of reference annotations containing the time position of target signals in a training data set to evaluate the performance of detection routines, using common signal detection theory indices.
+This is partly due to the relative ease with which sounds can be registered across a wide range of taxonomic groups and research settings.
+However, bioacoustics research can easily generate large amounts of data which might prove challenging to analyze in a timely manner.
+Many tools for the automated detection of sounds are currently available.
+Choosing the right approach for a specific task can be difficult as there are no available tools that can provide a common framework for evaluating detection performance.
+Here we present *ohun* an R package intended to facilitate the automated detection of sound events.
+*ohun* provides functions to diagnose and optimize detection routines as well as to compare the performance of different detection approaches.
+The package makes use of reference annotations containing the time position of target sounds in a training data set to evaluate the performance of detection routines, using common signal detection theory indices.
 This can be done with routine outputs imported from other software and from detection run within the package itself.
@@
 It also includes an implementation of two automatic detection methods commonly used in bioacoustic analysis: energy-based detection and template-based detection.
-We included examples on how to evaluate detections from external software and showcase the package usage with study cases on zebra-finch (*Taenopygia gutata*) songs and Spix's disc-winged bat (*Thyroptera tricolor*) ultrasonic social calls.
+We show how ohun can be used to automatically detect vocal signals with case studies of adult male zebra finch (*Taenopygia gutata*) songs and Spix's disc-winged bat (*Thyroptera tricolor*) ultrasonic social calls.
+We also include examples of how to evaluate the detection performance of ohun and external software.
 Finally, we provide some general suggestions to improve detection performance.
@@
 
-Animal acoustic signals are widely used to investigate a variety of questions in highly diverse areas, ranging from neurobiology to community ecology and evolutionary biology.
-The wide usage of animal sounds in research is partly due to the fact that they can be easily registered using non-intrusive methods, can be obtained in a variety of settings from laboratories to natural areas, and the equipment required for registering an analyzing these signals has become increasingly inexpensive.
-In addition, the existence of online repositories has facilitated the study of these traits at larger taxonomic and geographic scales.
-However, the adoption of bioacustic approaches in research can also imply large amounts of data (i.e.
-lots of recordings), which can be challenging to analyzed manually.
-As a result, a growing number of computational tools for the analysis of acoustic features on those signals is increasingly available (reviewed by @Stowell2022), reflecting the need of automated approaches for efficiently conducting these analyses.
+Animal acoustic signals are widely used to address a variety of questions in highly diverse areas, ranging from neurobiology [@Burgdorf2011; @schoneich2020] to taxonomy [@Kohler2017; @gew2019], community ecology [@zsebok2021; @tiwari2022] and evolutionary biology [@Odom2021; @Medina-Garcia2015a].
+The profuse usage of animal sounds in research is partly due to the fact that they can be easily collected using non-invasive methods.
+Animal sounds can also be obtained in a variety of settings from laboratories to natural areas, while the equipment required for collecting an analyzing these sounds has become increasingly inexpensive and broadly accesible [@Blumstein2011; @sugai2019].
+In addition, the existence of online repositories has facilitated the study of these communication signals at larger taxonomic and geographic scales.
+However, the adoption of bioacustic approaches in research may also imply large amounts of data (i.e.
+lots of recordings), which can be challenging to analyzed manually [@gibb2019].
+As a result, a growing number of computational tools for the automatic detection of animal sounds on is increasingly available [reviewed by @Stowell2022], reflecting the need of automated approaches for efficiently conducting these analyses [@gibb2019].
 
@@
 Fortunately, several tools have been develop to assist researchers in this matter.
-Indeed, new methods for automatic annotation of acoustic signals are continuously developed .
+Indeed, new methods for automatic annotation of sound events are continuously developed .
 -->
@@
 The growing availability of tools for automatic detection of acoustic events, particularly as free software, is expected to further simplify acoustic data processing, making it accessible to a wider range of users and scientific questions.
-However, this diversity of tools also posits a challenge as it can be hard to navigate.
-In this regard, the use of standard approaches for evaluating the performance of automatic detection tools might prove helpful to inform the decision about which method better fit our study system.
-The performance of automatic acoustic signal detection routines has been evaluated using standard indices from signal detection theory (CITATIONS).
-In its basic form, performance is assessed by comparing the output of a detection routine against a "gold standard" reference in which all the target signals have been annotated.
-This comparison allows to quantify the number of correctly detected signals (true positives), wrongly detected signals (false positives) and missed signals (false negatives) as well as additional metrics derived from these indices (e.g.
+However, this diversity of automated detection tools also posits a challenge as it can be difficult to navigate [@Stowell2022].
+In this regard, the use of standard approaches for evaluating the performance of automatic detection tools might prove helpful to inform researchers’ decisions about which method better fits a given question and study system [@knight2017].
+The performance of automated sound event detection routines have typically been evaluated using standard indices from signal detection theory [@balantic2020; @knight2017].
+In its basic form, performance is assessed by comparing the output of a detection routine against a 'gold standard' reference in which all the target sounds have been annotated (hereafter called 'reference annotation').
+This comparison facilitates quantifying the number of correctly detected sounds (true positives), wrongly detected sounds (false positives) and missed sounds (false negatives) as well as additional metrics derived from these indices (e.g.
 recall, precision).
-However, the fact that acoustic signals are not being evaluated as discrete classification units (as opposed to for instance identifying species from pictures) demand additional information to fully diagnose detection performance.
-This is particularly relevant when the precise time position of signals is needed as is common for research in which measures of the acoustic structure of signals is the main goal.
-For instance, the same signal can be detected as several separated signals, the inferred time position can be far off from the the target signal position, or several signals can be detected as a single one.
-Therefore, metrics that account for these other performance dimensions can be a valuable tool to properly diagnose automatic acoustic signal detection.
+However, the fact that sound events are not being evaluated as discrete classification units (as opposed to for instance identifying species from pictures), but are rather embedded within a continuous string of sound, demand additional information to fully diagnose the temporal precision detection performance.
+This is particularly relevant when the precise time position of sounds is needed as is common when measuring the acoustic structure of sounds is the main goal [@Araya-Salas2017f].
+For instance, the same signal can be detected as several separated sounds, the inferred time position can be offset from the the target signal position, or several sounds can be detected as a single one.
+Therefore, metrics that account for these additional performance dimensions are a valuable tool to properly diagnose automatic sound event detection.
 
-Here we present the new R package ohun.
-The package is intended to facilitate the automatic detection of acoustic signals, providing functions to diagnose specific aspects of acoustic detection routines in order to simplify their optimization.
-The package makes use of reference annotations containing the time position of target signals in a training data set to evaluate the performance of detection routines.
-This can be done with routine outputs imported from other software as well as detection run within the package itself.
+Here we present the new R package *ohun*.
+This package is intended to facilitate the automatic detection of sound events, providing functions to diagnose particular aspects of acoustic detection routines in order to simplify their optimization.
+The package makes use of reference annotations containing the time position of target sounds that along with the corresponding sound files, serves as a training data set in which to evaluate the performance of detection routines.
+This can be done with routine outputs imported from other software as well as detection routines run within the *ohun* package itself.
 The package also provides a set of functions to explore acoustic data sets and organize them in a format amenable for detection analyses.
-In addition, it offers implementations of two automatic detection methods commonly used in bioacoustic analysis: energy-based detection and template-based detection.
-We explain how to explore and format acoustic data sets, how acoustic signal detection routines can be evaluated and then showcase the package usage with study cases on zebra-finch songs (*Taenopygia gutata*) and Spix's disc-winged bats (*Thyroptera tricolor*) which represent different recording settings (lab and flight cages) and signal types (sonic mating signals and ultrasonic social calls).
+In addition, it offers implementations of two automatic detection methods commonly used in bioacoustic analysis: energy-based detection and template-based detection [@avi2014; @Hafner2015; @mellinger2000; @Charif2010; @aide2013].
+We explain how to explore and format acoustic data sets, how sound event detection routines can be evaluated and then showcase the package usage with study cases on male Zebra-finch songs (*Taenopygia gutata*) and Spix's disc-winged bat calls (*Thyroptera tricolor*) which represent different recording settings (lab and flight cages) and signal types (sonic mating sounds and ultrasonic social calls).
 
@@
 
-Having some sense of the format and size of the sound files to be analyzed is an important step for avoiding downstream errors and informing our time performance expectations.
-Several functions can facilitate getting acoustic data sets in the right format prior to automatic detection.
+The format and size of the acoustic data to be analyzed is important to validate and standardize in order to  avoid downstream errors, as well as to inform expectations for computational time performance.
+Several functions in *ohun* can facilitate double-checking the format of acoustic data sets prior to automatic detection.
 The function `feature_acoustic_data` prints a summary of the duration, size and format of all the recordings in a folder.
-Here we explore the acoustic data set of Zebra finch's songs (suppl.
-mat):
+Here we explore the acoustic data set of zebra finch's songs (Supporting Information):
 
@@
 
-In this case all recordings have format (.wav files, 44.1 kHz sampling rate, 16 bit resolution and a single channel).
+In this case all recordings have the same format (.wav files, 44.1 kHz sampling rate, 16 bit resolution and a single channel).
 We can also check the duration and size of files.
-This information is important as some tuning parameters of detection routines can behave differently depending on file format (e.g.
-time window size is affected by sampling rate) or simply because some software might can only work on specific sound file formats.
-In addition, long sound files might be hard to analyze in regular computers and might have to be split in shorter clips.
+Format information is important as some tuning parameters of detection routines can behave differently depending on file format (*e.g.* time window size can be affected by sampling rate) or simply because some software might only work on specific sound file formats.
+In addition, long sound files could be hard to analyze in some computers and might have to be split into shorter clips.
 In the latter case the function `split_acoustic_data` can be used to produce those clips:
@@
 
-The output shows to which time segments in the original sound files the clips belong to.
-If an annotation table is supplied (argument 'X') the function will adjust the annotations so they refer to the position of the signals in the clips.
+The output shows the time segments in the original sound files to which the clips belong to.
+If an annotation table is supplied (argument 'X') the function will adjust the annotations so they refer to the position of the sounds in the clips.
 This can be helpful when reference tables have been annotated on the original sound files.
 
-Annotations can also be explored using the function `feature_reference`.
-It returns about the mean and range of signal and gap duration (time intervals between selections), bottom and top frequency and of the number of annotations by sound file.
-If the path to the sound files is supplied the duty cycle (fraction of a sound file corresponding to signals) and peak amplitude (highest amplitude in a detection) are also returned:
+Annotations can also be explored using the function `feature_reference`, which returns the mean and range of signal duration and gap duration (time intervals between selections), bottom and top frequency and of the number of annotations by sound file.
+If the path to the sound files is supplied, then the duty cycle (the fraction of a sound file corresponding to sounds) and peak amplitude (the highest amplitude in a detection) are also returned:
 
@@
 
-The package makes use of signal detection theory indices to evaluate detection performance.
-Signal detection theory deals with the process of recovering signals (i.e.
-target signals) from background noise (not necessarily acoustic noise) and it’s widely used for optimizing this decision making process in the presence of uncertainty.
-During a detection routine, the detected ‘items’ can be classified into 4 classes: true positives (TPs, target signals correctly identified as signal), false positives (FPs, noise incorrectly identified as ‘signal’), false negatives (FNs, signals incorrectly identified as noise) and true negatives (TNs, background noise correctly identified as noise).
-However, TNs cannot always be easily defined in the context of acoustic signal detection as noise cannot always be partitioned in discrete units.
-Hence, the package makes use of TPs, FPs and FNs to calculate three additional indices that can further assist with evaluating the performance of a detection routine.
-
- - Recall: correct detections relative to total detections (a.k.a. true positive rate or sensitivity; TPs / (TPs + FNs))
-- Precision: correct detections relative to total detections (TPs / (TPs + FPs)).
-- F1 score: combines recall and precision as the harmonic mean of these two, providing a single value for evaluating performance (a.k.a. F-measure or Dice similarity coefficient).
+The *ohun* package makes use of signal detection theory indices to evaluate detection performance.
+Signal detection theory deals with the process of recovering sounds (*i.e.* target sounds) from background noise (not necessarily acoustic noise) and it’s widely used for optimizing this decision making process in the presence of uncertainty [@hossin2015].
+During a detection routine, the detected ‘items’ can be classified into 4 classes: true positives (TPs, target sounds correctly identified as signal), false positives (FPs, noise incorrectly identified as ‘signal’), false negatives (FNs, sounds incorrectly identified as noise) and true negatives (TNs, background noise correctly identified as noise).
+However, TNs cannot always be easily defined in the context of sound event detection as noise cannot always be partitioned in discrete units.
+Hence, the package makes use of TPs, FPs and FNs to calculate three additional indices that can further assist with evaluating the performance of a detection routine and are widely used in the context of sound event detection [@knight2017]: recall (correct detections relative to total detections), precision (proportion of target sounds that were correctly detected) and F1 score (combined recall and precision as the harmonic mean of these two, providing a single value for evaluating performance, a.k.a.
+F-score, F-measure or Dice similarity coefficient).
 
 A perfect detection will have no false positives or false negatives, which will result in both recall and precision equal to 1.
-However, perfect detection cannot always be reached and some compromise between detecting all target signals plus some noise (recall = 1 & precision < 1) and detecting only target signals but not all of them (recall < 1 & precision = 1) might be warranted.
-The right balance between these two extremes will be given by the relative costs of missing signals and mistaking noise for signals, given the specific goals of the study.
+However, perfect detection cannot always be achieved.
+Therefore, some compromise between detecting most target sounds plus some noise and excluding noise but missing target sounds might be warranted.
 These indices provide a useful framework for diagnosing and optimizing the performance of a detection routine.
+Researchers can identify an appropriate balance between these two extremes will be given by the relative costs of missing sounds and mistaking noise for target sounds in the context of their specific study goals.
 
-The package offers tools to evaluate the performance of an acoustic signal detection based on the indices described above.
-To accomplish this, annotations derived from a detection routine are compared against a reference table containing the time position of all target signals in the sound files.
-For instance, the following code evaluates a routine run in Raven Pro 1.6 (Cornell Lab of Ornithology) using the "band limited energy detector" option (minimum frequency: 0.8 kHz; maximum frequency: 22 kHz; minimum duration: 0.03968 s; maximum duration: 0.54989s; minimum separation: 0.02268 s) on a subset of the Zebra finch recordings described below:
+*ohun* offers tools to evaluate the performance of sound event detection methods based on the indices described above.
+To accomplish this, annotations derived from a detection routine are compared against a reference table containing the time position of all target sounds in the sound files.
+For instance, the following code evaluates a routine run in Raven Pro 1.6 [@Charif2010] using the "band limited energy detector" option (minimum frequency: 0.8 kHz; maximum frequency: 22 kHz; minimum duration: 0.03968 s; maximum duration: 0.54989s; minimum separation: 0.02268 s) on a subset of the zebra finch recordings described below:
 
@@
 
-The output shows the indices described above. The function also allows to detail those indices by sound file. Here we show the first 10 files:
+The output shows the indices described above.
+Plus three additional metrics specific to sound event detection: 'split positives', 'merge positives' and 'overlap to true positives'.
+'Split positives' is the number of reference sounds overlapped by more than one detection, 'merge positives' the number of detected sounds overlapping with another detection and 'overlap to true positives' quantifies the mean overlap between detections and reference signal (1 means complete overlap).
+The function also allows to detail those indices by sound file.
+Here we show the first 10 files:
 
@@
 
-Diagnostics from routines using different tuning parameters can be used to identify the parameter values that optimize a detection.
-These process of evaluating different routines for detection optimization is incorporated into the two signal detection approaches provided natively by ohun, which we depict in the following section.
-Note that the detection with Raven Pro does not necessarily reflect the best performance of this software and has been included only as an example on evaluating detection from external sources rather than a direct comparison of performance to ohun.
+Diagnostics from routines using different tuning parameters can be used to identify the parameter values that optimize detection.
+These process of evaluating different routines for detection optimization is incorporated into the two signal detection approaches provided natively by *ohun*, which we depict in the following section.
+Note that the detection with Raven Pro does not necessarily reflect the best performance of this software and has been included only as an example on evaluating detection from external sources, rather than a direct comparison of performance between Raven Pro and *ohun*.
 
-## Signal detection with ohun
+## Signal detection with *ohun*
 
-The package offers two methods for automatic signal detection: template-based and energy-based detection.
-These methods are better suited for highly stereotyped or good signal-to-noise ratio (SNR) signals, respectively.
-If the target signals do not fit these requirements, more elaborate methods (i.e.
+The package offers two methods for automated signal detection: template-based and energy-based detection.
+These methods are better suited for highly stereotyped or good signal-to-noise ratio (SNR) sounds, respectively.
+If the target sounds do not fit these requirements, more elaborate methods (i.e.
 machine/deep learning approaches) are warranted.
@@
 Bats were captured at their roosting sites (furled leaves of Zingeberaceae plants).
-Each individual bat was released in a large flight cage (9 x 4 x 3 m) for a 5 min period and their ultrasonic inquiry calls were recorded using a condenser microphone (CM16, Avisoft Bioacoustics, Glienike/Nordbahn, Germany) through an Avisoft UltraSoundGate 116Hm plugged into a laptop computer running Avisoft-Recorder software.
+Each individual bat was released in a large flight cage (9 x 4 x 3 m) for a 5-minute period and their ultrasonic inquiry calls were recorded using a condenser microphone (CM16, Avisoft Bioacoustics, Glienike/Nordbahn, Germany) through an Avisoft UltraSoundGate 116Hm plugged into a laptop computer running Avisoft-Recorder software.
 Recordings were made at a sampling rate of 500 kHz and an amplitude resolution of 16 bits.
 
-Recordings were manually annotated using Raven Pro 1.6 (Cornell Lab of Ornithology).
-Annotations were created by visual inspection of spectrograms, in which the start and end of signals were determined by the location of the continuous traces of power spectral entropy of the target signals.
+Recordings were manually annotated using Raven Pro 1.6 [@Charif2010].
+Annotations were created by visual inspection of spectrograms, in which the start and end of sounds were determined by the location of the continuous traces of power spectral entropy of the target sounds.
 A total of `r nrow(manual_ref_thy)` calls were annotated (~`r round(nrow(manual_ref_thy) / 30, 0)` calls per recording).
-Annotations were made with a time window of 200 samples and 70% of overlap, and were then imported into R using the package Rraven (@Araya-Salas2017e).
+Annotations were made with a time window of 200 samples and 70% of overlap, and were then imported into R using the package Rraven [@Araya-Salas2017e].
 
-Inquiry calls of Spix's disc-winged bats are structurally stereotyped (@Chaverri2010b).
-Most variation is found between individuals, although the basic form of a short, downward broadband frequency modulation is always shared (Fig.
+Inquiry calls of Spix's disc-winged bats are structurally stereotyped [@Chaverri2010b].
+Most variation is found among individuals, although the basic form of a short, downward broadband frequency modulation is always shared (Fig.
 BAT-SPECTRO, Araya-Salas et al 2021 ontogeny).
@@
 
-Template-based detection is a useful approach when there are few structural differences in the target signals.
-We used this approach in ohun to detect inquiry calls.
+Template-based detection is a useful approach when there are minimal structural differences in the target sounds  [e.g.
+when signals are produced in a highly stereotyped manner, @balantic2020; @knight2017].
+It uses spectrographic cross-correlation to find sounds resembling an example target sounds (i.e.
+template) across sound files, applying a correlation threshold to separate detections from background noise.
+We used this approach in *ohun* to detect inquiry calls.
 To do this, we tested the performance of three acoustic templates on a training subset of five sound files.
-The function `get_templates` finds several signals representative of the variation in signal structure.
-This function measures several spectrographic parameters which are then summarized using Principal Component Analysis.
+First we used the function `get_templates` to find several sounds representative of the variation in signal structure.
+This function measures several spectral parameters which are then summarized using Principal Component Analysis.
 The first two components are used to project the acoustic space.
-On this space the function defines sub-spaces as equal-size slices of a circle centered at the centroid of the acoustic space.
+In this space the function defines sub-spaces as equal-size slices of a sphere centered at the centroid of the acoustic space.
 Templates are then selected as those closer to the centroid within each of the sub-spaces, including the centroid for the entire acoustic space.
@@
 
-The output of the `get_templates` includes an acoustic space plot (FIG.
-ACOUSTIC-SPACE) in which the position of the signals selected as templates is highlighted.
-In the following code we used those templates for detecting calls.
+The output of the `get_templates` function includes an acoustic space plot (FIG.
+ACOUSTIC-SPACE) in which the position of the sounds selected as templates is highlighted.
+Users can also provide their own acoustic space dimensions (argument 'acoustic.space').
+In the following code we used those templates for detecting bat social calls.
 The code iterates a template-based detection on the training data set across a range of correlation thresholds, in order to find the combination of threshold and template with the best performance.
@@
 Note that the correlation vectors are estimated first (i.e.
-vectors of correlation values across sound files, `template_correlator()`) and then the correlation thresholds are optimized on them (`optimize_template_detector()`).
+vectors of correlation values across sound files, `template_correlator()`) and then the correlation thresholds are optimized on these vectors (`optimize_template_detector()`).
 The output of `optimize_template_detector()` contains the detection performance indices for each combination of templates and thresholds.
@@
 Performance diagnostic of template-based detections using four templates across several threshold values.
-Only the two highest performance iterations for each template are shown.")
+Only the two highest performance iterations for each template are shown.", digits = 3)
 
@@
 
-We used recordings from 18 zebra finch males recorded at obtained from the Rockefeller University Field Research Center Song Library (http://ofer.sci.ccny.cuny.edu/songs, @Tchernichovski2021).
+We used recordings from 18 zebra finch males recorded at obtained from the Rockefeller University Field Research Center Song Library [http://ofer.sci.ccny.cuny.edu/songs, @Tchernichovski2021].
 Recordings contain undirected vocalizations (songs or calls) of single males recorded in sound attenuation chambers using Sound Analysis Pro.
-Zebra finch vocalizations are composed by multiple elements (i.e.
-distinct patterns of continuous traces of power spectral entropy in the spectrogram separated by time gaps) that can vary importantly in key features as duration and frequency range (ZEBRAFINCH-SPECTRO).
-However, as the recorded signals show a good signal-to-noise ratio, they can potentially be detected using an energy-based approach.
+Zebra finch vocalizations are composed of multiple elements (i.e.
+distinct patterns of continuous traces of power spectral entropy in the spectrogram separated by time gaps) that can vary importantly in key features as duration and frequency range (ZEBRAFINCH-SPECTRO) and are clearly not as stereotyped as the Spix’s disc-winged bats.
+However, as the recorded sounds show a good signal-to-noise ratio, such that signals in each recording can potentially be detected using an energy-based approach that does not rely on matching the acoustic structure of a template.
 ```{r catalog zebra f, out.width = "100%", echo = FALSE, fig.align= "left",fig.cap="Fig.
@@
 Reference annotations were made manually on the oscillogram with the spectrogram and audio as a guide using Raven Lite 2.0.1 (Cornell Lab of Ornithology).
-The following code loads the reference annotations and split them in two data set for training (3 sound files) and testing (15 sound files):
+The following code loads the reference annotations and split them into two data sets for training (3 sound files) and testing (15 sound files):
 ```{r, eval = TRUE, echo = FALSE}
@@
 
-Now we can optimize the detection parameters using the function `optimize_energy_detector`.
+The detection parameters can be optimized using the function `optimize_energy_detector`.
 This function runs a detection for all possible combinations of tuning parameters.
-The code below tries three minimum duration and maximum duration values and two hold time values (which merges signals within the specified time interval):
+The code below tries three minimum duration and maximum duration values and two hold time values (which merges sounds within the specified time interval):
 
@@
 
-We now can use the tuning parameter values that produce the best performance to detect signals on the test data set:
+We now can use the tuning parameter values that yielded the best performance to detect sounds on the test data set:
 
@@
 Note that in the example we used a small subset of sound files for training.
-More training data might be needed for optimizing a detection routine on larger data sets or recordings with more variable signals or background noise levels.
+More training data might be needed for optimizing a detection routine on larger data sets or recordings with more variable sounds or background noise levels.
 
@@
 
-The package offers additional tools to simplify acoustic signal detection.
-Detected signals can be labeled as false or true positives with the function 'label_detection'.
+The *ohun* package offers additional tools to simplify sound event detection.
+Detected sounds can be labeled as false or true positives with the function 'label_detection'.
 This allows users to explore the structure of false positives and figure out ways to exclude them.
-The function 'filter_detections' can be used to remove ambiguous signals (i.e.
-those labeled as split or merged detection) keeping only those that maximize a specific criterium (i.e.
-the highest template correlacion).
-Finally, note that  several templates representing the range of variation in signal structure can be used to detect semi-stereotyped signals when running template-based detection ('template_detection' function).
+The function 'filter_detections' can be used to remove ambiguous sounds (i.e.
+those labeled as split or merged detections) keeping only those that maximize a specific criterium (i.e.
+the highest template correlation).
+Finally, note that several templates representing the range of variation in signal structure can be used to detect semi-stereotyped sounds when running template-based detection ('template_detection' function).
 
@@
 
-Here we have shown how to evaluate the performance of acoustic signal detection routines using the package ohun.
+Here we have shown how to evaluate the performance of sound event detection routines using the package *ohun*.
 The package can evaluate detection outputs imported from other software, as well as its own detection routines.
-The latter can be iterated over combinations of tuning parameters in order to find those values that optimize a detection.
+The latter can be iterated over combinations of tuning parameters in order to find those values that optimize detection.
 Despite signal detection indices being commonly reported when presenting new automatic detection methods, to our knowledge widely applicable performance evaluating routines have not been made available in a free, open source platform.
-Providing a common framework for acoustic signal detection evaluation can simplify comparing performances of different tools and the selection of those better suited to our system and research question.
-Note that the tools offered by ohun for diagnosing a detection should not be necessarily limited to acoustic data.
-Any type of automatic detection in which the time of occurrence of discrete events needs to be found can be evaluated and optimized by comparing it to a reference annotation, for instance, detection of specific behaviors in video analysis of animal motor activity (Sturman et al 2020; Hsu & Yttri 2021; DeepEthogram).
-The package can also be used to assess the repeatability of manually made annotations by comparing annotations from different observers on the same data sets.
+Providing a common framework for the evaluation of sound event detection can simplify comparing the performance of different tools and the selection of those tools better suited to a given research question and study system.
+The tools offered by *ohun* for diagnosing detection performance should not be necessarily limited to acoustic data.
+*ohun* can also be used for cases in which the time of occurrence of discrete events needs to be identified such as detecting specific behaviors in video analysis of animal motor activity [*e.g.* @sturman2020; @hsu2021; @bohnslav2021].
+The detection of such motor events in video can also be evaluated and optimized in comparison to a reference annotation, as we have shown above for acoustic signals.
 
-The package provides two native detection methods: template-based  and energy-based detection.
-Compared to new deep learning approaches for finding the occurrence of acoustic signals, the two native methods are relatively simple tools.
-However, these methods have been widely used by the bioacoustic community and, under the appropriate conditions can reach adequate performance, as we have shown in our two study cases.
-Deep learning methods tend to require larger computational power, larger training data sets and, in some cases, more complex training routines.
+The *ohun* package provides two native detection methods: template-based and energy-based detection.
+Compared to new deep learning approaches for finding the occurrence of sound events, the two native methods are relatively simple tools.
+However, these methods have been widely used by the bioacoustic community [@avi2014; @Hafner2015; @mellinger2000; @Charif2010; @aide2013] and can reach adequate performance under the appropriate conditions, as we have shown in our two study cases and has been shown elsewhere [@knight2017].
+Deep learning methods tend to require greater computational power, larger training data sets and, in some cases, more complex training routines.
 This might bring unnecessary difficulties when dealing with when less challenging detection tasks.
-Therefore the availability of a wide range of approaches is desired in order to simplify finding the most appropriate tool for the intricacies of our study system and research goals.
-Note that the tools offered in ohun can be used in a pipeline in which detected signals are further classified using more elaborated discrimination algorithms.
-Acoustic parameters can be used to quantify the structure of the signals and distinguish target from non-target signals.
+Therefore the availability of a wide range of approaches can simplify finding the most appropriate tool for the intricacies of a study system and research goals, as well as having tools accessible for a broader community of researchers.
+The tools offered in *ohun* can also be used in a subsequent pipeline in which detected sounds are further classified using more elaborated discrimination algorithms [@balantic2020].
+In addition, detection performance can be improved by using acoustic structure measurements that can be used to distinguish target from non-target sound events.
 
@@
 large recordings and/or many files).
-These are some tips that can help make a routine more time-efficient.
+We provide some additional tips that can help make a routine more time-efficient.
 1.
 Always test procedures on small data subsets.
-Make sure you are getting decent results on a small subset of recordings before trying to scale up the analysis.
+Make sure that you obtain decent results on a small subset of recordings before trying to scale up the analysis.
 2.
@@
 Sampling rate matters.
-Detecting signals on low sampling rate files is faster, so we must avoid having Nyquist frequencies much higher than the highest frequency of the target signals.
+Detecting sounds on low sampling rate files is faster, so we must avoid having Nyquist frequencies much higher than the highest frequency of the target sounds.
 5.
 Try using a computer with lots of RAM memory or a computer cluster for working on large amounts of data.
-These rules are not restricted to ohun and can also be helpful to speed-up routines in other software packages.
+Lastly, we underscore that these tips are not restricted to *ohun* and can also be helpful to speed up routines in other software packages.
 
-There are some additional things to be considered when aiming to automatically detect acoustic signals.
-When running automatic signals detection try to use your knowledge about the signal structure to determine the initial range.
-This can be extremely helpful for narrowing down possible parameter values, particularly for energy-based detection.
-As a general rule, if people have a hard time figuring out where a target signal occurs, detection algorithms will also have a hard time.
+There are some additional things to be considered when aiming to automatically detect sound events.
+When running energy-based detection routines try to use your knowledge about the signal structure to determine the initial range of tunning parameters.
+This can be extremely helpful for narrowing down possible parameter values.
+As a general rule, if human observers have a hard time detecting where a target sound occurs in a sound file, detection algorithms will likely yield low detection performance.
 In cases in which occurrences are ambiguous low performances are expected.
-Make sure reference tables contain all target signals and only the target signals, otherwise performance optimization can be misleading as the performance of a detection cannot be better than the reference itself.
-Lastly, avoid having overlapping signals or several signals as a single one (like a multi-syllable vocalization) in the reference table when running an energy-based detector, as they are likely to be identified as separated units.
+Ensure reference annotations contain all target sounds and only the target sounds, otherwise performance optimization can be misleading as the performance of a given detection method cannot be better than the reference itself.
+Lastly, avoid having overlapping sounds or several sounds as a single detection (like a multi-syllable vocalization) in the reference annotation when running an energy-based detector, as they are likely to be identified as separated units.
 
@@
 Nazareth Rojas, Silvia Chaves-Ramírez, Mariela Sánchez-Chavarría, Miriam Gioiosa, Cristian Castillo-Salazar, and José Pablo Barrantes for their help collecting acoustic data for Spix’s disc-winged bats and Jorge Elizondo for his help on manually annotating recordings.
-We also thank the Centro Biológico Hacienda Barú for their continuous support of our research.
-This study was partly funded by a CONARE-Max Planck grant from the Consejo Nacional de Rectores.
+We also thank the Centro Biológico Hacienda Barú for their continuous support of our research and Mijail Rojas and Andrey Sequeira for support on early stages of package development.
+This study was partly funded by a CONARE-Max Planck grant from the Consejo Nacional de Rectores and the research activity C0754 at Centro de Investigación en Neurociencias, Universidad de Costa Rica.
+A.R-G.
+is supported by the Walt Halperin Endowed Professorship and the Washington Research Foundation as Distinguished Investigator.
 
@@
 
-All sampling protocols followed guidelines approved by the American Society of Mammalogists for capture, handling and care of mammals (@Sikes2016) and the ASAB/ABS Guidelines for the use of animals in research.
+All sampling protocols followed guidelines approved by the American Society of Mammalogists for capture, handling and care of mammals [@Sikes2016] and the ASAB/ABS Guidelines for the use of animals in research.
 This study was conducted in accordance with the ethical standards for animal welfare of the Costa Rican Ministry of Environment and Energy, Sistema Nacional de Áreas de Conservación, permit no.
@@
 
-### Supplementary materials
+### Supporting Information
 
-Supplementary material associated with this article is available in XXXXXXXXXXXXXXXXX
+Supplementary information associated with this article is available in XXXXXXXXXXXXXXXXX
+
 ## References
-
-Sturman, O., von Ziegler, L., Schläppi, C.
-et al.
-Deep learning-based behavioral analysis reaches human accuracy and is capable of outperforming commercial solutions.
-Neuropsychopharmacol.
-45, 1942–1952 (2020).
-https://doi.org/10.1038
-
-Hsu, A.I., Yttri, E.A. B-SOiD, an open-source unsupervised algorithm for identification and fast prediction of behaviors. Nat Commun 12, 5188 (2021). https://doi.org/10.1038/s41467-021-25420-x

### 2022-08-16 manuscript/manuscript.Rmd
update manuscript

@@
 ---
-title: 'ohun: an R package for diagnosing and optimizing automatic sound event detection'
+title: 'ohun: an R package for diagnosing and optimizing automated sound event detection'
 author:
@@
 Gloriana Chaverri ^3^, ^7^
-Juan Carlos Brenes ^1^ &
+Juan Carlos Brenes ^1^
 Fabiola Chirino ^1^ &
@@
 In this regard, the use of standard approaches for evaluating the performance of automatic detection tools might prove helpful to inform researchers’ decisions about which method better fits a given question and study system [@knight2017].
-The performance of automated sound event detection routines have typically been evaluated using standard indices from signal detection theory [@balantic2020; @knight2017].
+The performance of automated sound event detection routines have typically been evaluated using standard indices from signal detection theory [@macmillan2004; @balantic2020; @knight2017].
 In its basic form, performance is assessed by comparing the output of a detection routine against a 'gold standard' reference in which all the target sounds have been annotated (hereafter called 'reference annotation').
@@
 During a detection routine, the detected ‘items’ can be classified into 4 classes: true positives (TPs, target sounds correctly identified as signal), false positives (FPs, noise incorrectly identified as ‘signal’), false negatives (FNs, sounds incorrectly identified as noise) and true negatives (TNs, background noise correctly identified as noise).
-However, TNs cannot always be easily defined in the context of sound event detection as noise cannot always be partitioned in discrete units.
+However, TNs cannot always be easily defined in the context of sound event detection as noise might not be partitioned in discrete units.
 Hence, the package makes use of TPs, FPs and FNs to calculate three additional indices that can further assist with evaluating the performance of a detection routine and are widely used in the context of sound event detection [@knight2017]: recall (correct detections relative to total detections), precision (proportion of target sounds that were correctly detected) and F1 score (combined recall and precision as the harmonic mean of these two, providing a single value for evaluating performance, a.k.a.

### 2023-02-14 manuscript/manuscript.Rmd
update figs and manuscript

@@
 
-Animal acoustic signals are widely used in a diversity of research areas.
-This is partly due to the relative ease with which sounds can be registered across a wide range of taxonomic groups and research settings.
-However, bioacoustics research can easily generate large amounts of data which might prove challenging to analyze in a timely manner.
-Many tools for the automated detection of sounds are currently available.
-Choosing the right approach for a specific task can be difficult as there are no available tools that can provide a common framework for evaluating detection performance.
-Here we present *ohun* an R package intended to facilitate the automated detection of sound events.
-*ohun* provides functions to diagnose and optimize detection routines as well as to compare the performance of different detection approaches.
-The package makes use of reference annotations containing the time position of target sounds in a training data set to evaluate the performance of detection routines, using common signal detection theory indices.
-This can be done with routine outputs imported from other software and from detection run within the package itself.
-The package also provides a set of functions to organize acoustic data sets in a format amenable for detection analyses.
-It also includes an implementation of two automatic detection methods commonly used in bioacoustic analysis: energy-based detection and template-based detection.
-We show how ohun can be used to automatically detect vocal signals with case studies of adult male zebra finch (*Taenopygia gutata*) songs and Spix's disc-winged bat (*Thyroptera tricolor*) ultrasonic social calls.
+Animal acoustic signals are widely used in diverse research areas due to the relative ease with which sounds can be registered across a wide range of taxonomic groups and research settings.
+However, bioacoustics research can quickly generate large data sets, which might prove challenging to analyze promptly.
+Although many tools are available for the automated detection of sounds, choosing the right approach can be difficult only a few tools provide a framework for evaluating detection performance.
+Here we present ohun, an R package intended to facilitate automated sound detection.
+ohun provides functions to diagnose and optimize detection routines and compare performance among different detection approaches.
+The package uses reference annotations containing the time position of target sounds in a training data set to evaluate detection routines performance using common signal detection theory indices.
+This can be done both with routine outputs imported from other software and detections run within the package.
+The package also provides functions to organize acoustic data sets in a format amenable to detection analyses.
+ohun also includes energy-based and template-based detection methods –two commonly used automatic approaches in bioacoustic research.
+We show how ohun automatically can be used to  detect vocal signals with case studies of adult male zebra finch (Taenopygia gutata) songs and Spix’s disc-winged bat (Thyroptera tricolor) ultrasonic social calls.
 We also include examples of how to evaluate the detection performance of ohun and external software.
@@
 The function 'filter_detections' can be used to remove ambiguous sounds (i.e.
-those labeled as split or merged detections) keeping only those that maximize a specific criterium (i.e.
+those labeled as split or merged detections) keeping only those that maximize a specific criterion (i.e.
 the highest template correlation).

### 2023-05-24 manuscript/manuscript.Rmd
many changes

@@
 output:
+  word_document:
+    toc: no
+    toc_depth: '3'
+  pdf_document:
+    toc: yes
+    toc_depth: '3'
   html_document:
@@
       smooth_scroll: no
-  word_document:
-    toc: yes
-    toc_depth: '3'
-  pdf_document:
-    toc: yes
-    toc_depth: '3'
 always_allow_html: yes
@@
 
+# options to customize chunk outputs
+knitr::opts_chunk$set(
+  class.source = "numberLines lineAnchors", # for code line numbers
+  tidy.opts = list(width.cutoff = 65), 
+  tidy = TRUE,
+  message = FALSE,
+  warning = FALSE
+ )
 ```
@@
 
-```{r word count, eval = TRUE, echo = FALSE, results='asis'}
+```{r word count, eval = FALSE, echo = FALSE, results='asis'}
 
@@
 
-head(split_info)
+as.matrix(split_info[1:6,])
 ```
@@
 
-```{r, eval = TRUE, echo = TRUE}
+```{r, eval = TRUE, echo = FALSE}
 
@@
 
-diagnose_detection(reference = manual_ref_tae, detection = raven_detec, by = "tuning_parameters")
+dd <- diagnose_detection(reference = manual_ref_tae, detection = raven_detec, by = "tuning_parameters")
+
+as.matrix(dd)
 
@@
 
-```{r, eval = TRUE, echo = TRUE}
+```{r, eval = FALSE, echo = TRUE}
 
@@
 
+```{r, eval = TRUE, echo = TRUE}
+
+diag_raven <- diagnose_detection(reference = manual_ref_tae, detection = raven_detec, by = "tuning_parameters", by.sound.file = TRUE)
+
+as.matrix(diag_raven[1:10, ])
+
+```
 Diagnostics from routines using different tuning parameters can be used to identify the parameter values that optimize detection.
@@
 source("~/Dropbox/R_package_testing/ohun/R/get_templates.R")
+source("~/Dropbox/R_package_testing/ohun/R/internal_functions.R")
 
@@
 
-diagnose_detection(reference = test_ref, detection = detec_test)
+dd <- diagnose_detection(reference = test_ref, detection = detec_test)
+
+as.matrix(dd)
 ```
@@
 
-sub_opt_det_train
+as.matrix(sub_opt_det_train)
 ```

### 2023-05-24 manuscript/manuscript.tex
many changes

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={ohun: an R package for diagnosing and optimizing automated sound event detection},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{ohun: an R package for diagnosing and optimizing automated sound
event detection}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

{
\setcounter{tocdepth}{3}
\tableofcontents
}
Marcelo Araya-Salas \textsuperscript{1,} \textsuperscript{2,}
\textsuperscript{3} *, Grace Smith-Vidaurre \textsuperscript{4,}
\textsuperscript{5,} \textsuperscript{6}, Gloriana Chaverri
\textsuperscript{3}, \textsuperscript{7} Juan Carlos Brenes
\textsuperscript{1} Fabiola Chirino \textsuperscript{1} \& Alejandro
Rico-Guevara \textsuperscript{8,} \textsuperscript{9}

~

\textsuperscript{1} Centro de Investigación en Neurociencias,
Universidad de Costa Rica, San José, Costa Rica

\textsuperscript{2} Escuela de Biología, Universidad de Costa Rica, San
José, Costa Rica

\textsuperscript{3} Sede del Sur, Universidad de Costa Rica, Golfito,
Costa Rica

\textsuperscript{4} Laboratory of Neurogenetics of Language, Rockefeller
University, New York, NY, USA

\textsuperscript{5} Rockefeller University Field Center, Millbrook, NY,
USA

\textsuperscript{6} Department of Biological Sciences, University of
Cincinnati, Cincinnati, OH, USA

\textsuperscript{7} Smithsonian Tropical Research Institute, Panama
City, Panamá

\textsuperscript{8} Department of Biology, University of Washington,
Seattle, USA

\textsuperscript{9} Burke Museum of Natural History and Culture,
University of Washington, Seattle, USA

* \emph{To whom correspondence should be addressed}

\textbf{Keywords}: bioacoustics, automatic detection, animal
vocalizations

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

~

Word count: 4279 (including code chunks: 5925)

Check
\href{https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html}{this
link} for details on how to add citations/bibliography in Rmarkdown

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

Animal acoustic signals are widely used in diverse research areas due to
the relative ease with which sounds can be registered across a wide
range of taxonomic groups and research settings. However, bioacoustics
research can quickly generate large data sets, which might prove
challenging to analyze promptly. Although many tools are available for
the automated detection of sounds, choosing the right approach can be
difficult only a few tools provide a framework for evaluating detection
performance. Here we present ohun, an R package intended to facilitate
automated sound detection. ohun provides functions to diagnose and
optimize detection routines and compare performance among different
detection approaches. The package uses reference annotations containing
the time position of target sounds in a training data set to evaluate
detection routines performance using common signal detection theory
indices. This can be done both with routine outputs imported from other
software and detections run within the package. The package also
provides functions to organize acoustic data sets in a format amenable
to detection analyses. ohun also includes energy-based and
template-based detection methods --two commonly used automatic
approaches in bioacoustic research. We show how ohun automatically can
be used to detect vocal signals with case studies of adult male zebra
finch (Taenopygia gutata) songs and Spix's disc-winged bat (Thyroptera
tricolor) ultrasonic social calls. We also include examples of how to
evaluate the detection performance of ohun and external software.
Finally, we provide some general suggestions to improve detection
performance.

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Animal acoustic signals are widely used to address a variety of
questions in highly diverse areas, ranging from neurobiology
(\protect\hyperlink{ref-Burgdorf2011}{Burgdorf \emph{et al.} 2011};
\protect\hyperlink{ref-schoneich2020}{S 2020}) to taxonomy
(\protect\hyperlink{ref-Kohler2017}{Köhler \emph{et al.} 2017};
\protect\hyperlink{ref-gew2019}{Gwee \emph{et al.} 2019}), community
ecology (\protect\hyperlink{ref-zsebok2021}{Zsebők \emph{et al.} 2021};
\protect\hyperlink{ref-tiwari2022}{Tiwari \& Diwakar 2022}) and
evolutionary biology
(\protect\hyperlink{ref-Medina-Garcia2015a}{Medina-García \emph{et al.}
2015}; \protect\hyperlink{ref-Odom2021}{Odom \emph{et al.} 2021}). The
profuse usage of animal sounds in research is partly due to the fact
that they can be easily collected using non-invasive methods. Animal
sounds can also be obtained in a variety of settings from laboratories
to natural areas, while the equipment required for collecting an
analyzing these sounds has become increasingly inexpensive and broadly
accesible (\protect\hyperlink{ref-Blumstein2011}{Blumstein \emph{et al.}
2011}; \protect\hyperlink{ref-sugai2019}{Sugai \emph{et al.} 2019}). In
addition, the existence of online repositories has facilitated the study
of these communication signals at larger taxonomic and geographic
scales. However, the adoption of bioacustic approaches in research may
also imply large amounts of data (i.e.~lots of recordings), which can be
challenging to analyzed manually (\protect\hyperlink{ref-gibb2019}{Gibb
\emph{et al.} 2019}). As a result, a growing number of computational
tools for the automatic detection of animal sounds on is increasingly
available (reviewed by \protect\hyperlink{ref-Stowell2022}{Stowell
2022}), reflecting the need of automated approaches for efficiently
conducting these analyses (\protect\hyperlink{ref-gibb2019}{Gibb
\emph{et al.} 2019}).

The growing availability of tools for automatic detection of acoustic
events, particularly as free software, is expected to further simplify
acoustic data processing, making it accessible to a wider range of users
and scientific questions. However, this diversity of automated detection
tools also posits a challenge as it can be difficult to navigate
(\protect\hyperlink{ref-Stowell2022}{Stowell 2022}). In this regard, the
use of standard approaches for evaluating the performance of automatic
detection tools might prove helpful to inform researchers' decisions
about which method better fits a given question and study system
(\protect\hyperlink{ref-knight2017}{Knight \emph{et al.} 2017}). The
performance of automated sound event detection routines have typically
been evaluated using standard indices from signal detection theory
(\protect\hyperlink{ref-macmillan2004}{Macmillan \& Creelman 2004};
\protect\hyperlink{ref-knight2017}{Knight \emph{et al.} 2017};
\protect\hyperlink{ref-balantic2020}{Balantic \& Donovan 2020}). In its
basic form, performance is assessed by comparing the output of a
detection routine against a `gold standard' reference in which all the
target sounds have been annotated (hereafter called `reference
annotation'). This comparison facilitates quantifying the number of
correctly detected sounds (true positives), wrongly detected sounds
(false positives) and missed sounds (false negatives) as well as
additional metrics derived from these indices (e.g.~recall, precision).
However, the fact that sound events are not being evaluated as discrete
classification units (as opposed to for instance identifying species
from pictures), but are rather embedded within a continuous string of
sound, demand additional information to fully diagnose the temporal
precision detection performance. This is particularly relevant when the
precise time position of sounds is needed as is common when measuring
the acoustic structure of sounds is the main goal
(\protect\hyperlink{ref-Araya-Salas2017f}{Araya-Salas \& Smith-Vidaurre
2017}). For instance, the same signal can be detected as several
separated sounds, the inferred time position can be offset from the the
target signal position, or several sounds can be detected as a single
one. Therefore, metrics that account for these additional performance
dimensions are a valuable tool to properly diagnose automatic sound
event detection.

Here we present the new R package \emph{ohun}. This package is intended
to facilitate the automatic detection of sound events, providing
functions to diagnose particular aspects of acoustic detection routines
in order to simplify their optimization. The package makes use of
reference annotations containing the time position of target sounds that
along with the corresponding sound files, serves as a training data set
in which to evaluate the performance of detection routines. This can be
done with routine outputs imported from other software as well as
detection routines run within the \emph{ohun} package itself. The
package also provides a set of functions to explore acoustic data sets
and organize them in a format amenable for detection analyses. In
addition, it offers implementations of two automatic detection methods
commonly used in bioacoustic analysis: energy-based detection and
template-based detection
(\protect\hyperlink{ref-mellinger2000}{Mellinger \& Clark 2000};
\protect\hyperlink{ref-Charif2010}{Charif \emph{et al.} 2010};
\protect\hyperlink{ref-aide2013}{Aide \emph{et al.} 2013};
\protect\hyperlink{ref-avi2014}{Bioacoustics 2014};
\protect\hyperlink{ref-Hafner2015}{Hafner \& Katz 2015}). We explain how
to explore and format acoustic data sets, how sound event detection
routines can be evaluated and then showcase the package usage with study
cases on male Zebra-finch songs (\emph{Taenopygia gutata}) and Spix's
disc-winged bat calls (\emph{Thyroptera tricolor}) which represent
different recording settings (lab and flight cages) and signal types
(sonic mating sounds and ultrasonic social calls).

\hypertarget{formatting-acoustic-data-sets}{%
\subsection{Formatting acoustic data
sets}\label{formatting-acoustic-data-sets}}

The format and size of the acoustic data to be analyzed is important to
validate and standardize in order to avoid downstream errors, as well as
to inform expectations for computational time performance. Several
functions in \emph{ohun} can facilitate double-checking the format of
acoustic data sets prior to automatic detection. The function
\texttt{feature\_acoustic\_data} prints a summary of the duration, size
and format of all the recordings in a folder. Here we explore the
acoustic data set of zebra finch's songs (Supporting Information):

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# working directory}
\NormalTok{path\_zebra\_finch }\OtherTok{\textless{}{-}} \StringTok{"path\_to\_zebra\_finch\_files"}

\FunctionTok{feature\_acoustic\_data}\NormalTok{(}\AttributeTok{path =}\NormalTok{ path\_zebra\_finch)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Features of the acoustic data set in '/home/m/Dropbox/Projects/ohun_paper/data/raw/taeniopygia':
\end{verbatim}

\begin{verbatim}
## 
## * 18 sound files
\end{verbatim}

\begin{verbatim}
## 
## * 1 file format(s) (.wav (18))
\end{verbatim}

\begin{verbatim}
## 
## * 1 sampling rate(s) (44.1 kHz (18))
\end{verbatim}

\begin{verbatim}
## 
## * 1 bit depth(s) (16 bits (18))
\end{verbatim}

\begin{verbatim}
## 
## * 1 number of channels (1 channel(s) (18))
\end{verbatim}

\begin{verbatim}
## 
## * File duration range: 2.63-21.76 s (mean: 12.22 s)
\end{verbatim}

\begin{verbatim}
## 
## * File size range: 0.23-1.92 MB (mean: 1.08 MB)
\end{verbatim}

\begin{verbatim}
## 
##  (detailed information by sound file can be obtained with 'warbleR::info_sound_files()')
\end{verbatim}

In this case all recordings have the same format (.wav files, 44.1 kHz
sampling rate, 16 bit resolution and a single channel). We can also
check the duration and size of files. Format information is important as
some tuning parameters of detection routines can behave differently
depending on file format (\emph{e.g.} time window size can be affected
by sampling rate) or simply because some software might only work on
specific sound file formats. In addition, long sound files could be hard
to analyze in some computers and might have to be split into shorter
clips. In the latter case the function \texttt{split\_acoustic\_data}
can be used to produce those clips:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{split\_info }\OtherTok{\textless{}{-}} \FunctionTok{split\_acoustic\_data}\NormalTok{(}\AttributeTok{path =}\NormalTok{ path\_zebra\_finch, }\AttributeTok{sgmt.dur =} \DecValTok{5}\NormalTok{)}

\FunctionTok{head}\NormalTok{(split\_info)}
\end{Highlighting}
\end{Shaded}

\begin{table}
\centering
\begin{tabular}[t]{c|c|c|c}
\hline
original.sound.files & sound.files & start & end\\
\hline
Ag13\_43421.27975590\_11\_17\_7\_46\_15.wav & Ag13\_43421.27975590\_11\_17\_7\_46\_15-1.wav & 0 & 5.000\\
\hline
Ag13\_43421.27975590\_11\_17\_7\_46\_15.wav & Ag13\_43421.27975590\_11\_17\_7\_46\_15-2.wav & 5 & 10.000\\
\hline
Ag13\_43421.27975590\_11\_17\_7\_46\_15.wav & Ag13\_43421.27975590\_11\_17\_7\_46\_15-3.wav & 10 & 15.000\\
\hline
Ag13\_43421.27975590\_11\_17\_7\_46\_15.wav & Ag13\_43421.27975590\_11\_17\_7\_46\_15-4.wav & 15 & 20.000\\
\hline
Ag13\_43421.27975590\_11\_17\_7\_46\_15.wav & Ag13\_43421.27975590\_11\_17\_7\_46\_15-5.wav & 20 & 21.761\\
\hline
Blk109Brn\_43559.32349131\_4\_4\_8\_59\_9.wav & Blk109Brn\_43559.32349131\_4\_4\_8\_59\_9-1.wav & 0 & 5.000\\
\hline
\end{tabular}
\end{table}

The output shows the time segments in the original sound files to which
the clips belong to. If an annotation table is supplied (argument `X')
the function will adjust the annotations so they refer to the position
of the sounds in the clips. This can be helpful when reference tables
have been annotated on the original sound files.

Annotations can also be explored using the function
\texttt{feature\_reference}, which returns the mean and range of signal
duration and gap duration (time intervals between selections), bottom
and top frequency and of the number of annotations by sound file. If the
path to the sound files is supplied, then the duty cycle (the fraction
of a sound file corresponding to sounds) and peak amplitude (the highest
amplitude in a detection) are also returned:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# read reference annotations}
\NormalTok{manual\_ref\_tae }\OtherTok{\textless{}{-}} \FunctionTok{read.csv}\NormalTok{(}\FunctionTok{file.path}\NormalTok{(path\_zebra\_finch, }\StringTok{"manual\_selections\_Taeniopygia.csv"}\NormalTok{))}

\CommentTok{\# explore annotations}
\FunctionTok{feature\_reference}\NormalTok{(}\AttributeTok{reference =}\NormalTok{ manual\_ref\_tae, }\AttributeTok{path =}\NormalTok{ path\_zebra\_finch)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##                  min   mean     max
## sel.duration   15.54 103.30  319.41
## gap.duration   80.19 352.26 3024.23
## annotations     2.00  32.83   66.00
## duty.cycle      0.09   0.29    0.61
## peak.amplitude 43.28  65.19   88.69
## bottom.freq     0.50   0.50    0.50
## top.freq       10.00  10.00   10.00
\end{verbatim}

\hypertarget{diagnosing-detection-performance}{%
\subsection{Diagnosing detection
performance}\label{diagnosing-detection-performance}}

The \emph{ohun} package makes use of signal detection theory indices to
evaluate detection performance. Signal detection theory deals with the
process of recovering sounds (\emph{i.e.} target sounds) from background
noise (not necessarily acoustic noise) and it's widely used for
optimizing this decision making process in the presence of uncertainty
(\protect\hyperlink{ref-hossin2015}{Hossin \& Sulaiman 2015}). During a
detection routine, the detected `items' can be classified into 4
classes: true positives (TPs, target sounds correctly identified as
signal), false positives (FPs, noise incorrectly identified as
`signal'), false negatives (FNs, sounds incorrectly identified as noise)
and true negatives (TNs, background noise correctly identified as
noise). However, TNs cannot always be easily defined in the context of
sound event detection as noise might not be partitioned in discrete
units. Hence, the package makes use of TPs, FPs and FNs to calculate
three additional indices that can further assist with evaluating the
performance of a detection routine and are widely used in the context of
sound event detection (\protect\hyperlink{ref-knight2017}{Knight
\emph{et al.} 2017}): recall (correct detections relative to total
detections), precision (proportion of target sounds that were correctly
detected) and F1 score (combined recall and precision as the harmonic
mean of these two, providing a single value for evaluating performance,
a.k.a. F-score, F-measure or Dice similarity coefficient).

A perfect detection will have no false positives or false negatives,
which will result in both recall and precision equal to 1. However,
perfect detection cannot always be achieved. Therefore, some compromise
between detecting most target sounds plus some noise and excluding noise
but missing target sounds might be warranted. These indices provide a
useful framework for diagnosing and optimizing the performance of a
detection routine. Researchers can identify an appropriate balance
between these two extremes will be given by the relative costs of
missing sounds and mistaking noise for target sounds in the context of
their specific study goals.

\emph{ohun} offers tools to evaluate the performance of sound event
detection methods based on the indices described above. To accomplish
this, annotations derived from a detection routine are compared against
a reference table containing the time position of all target sounds in
the sound files. For instance, the following code evaluates a routine
run in Raven Pro 1.6 (\protect\hyperlink{ref-Charif2010}{Charif \emph{et
al.} 2010}) using the ``band limited energy detector'' option (minimum
frequency: 0.8 kHz; maximum frequency: 22 kHz; minimum duration: 0.03968
s; maximum duration: 0.54989s; minimum separation: 0.02268 s) on a
subset of the zebra finch recordings described below:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{raven\_detec }\OtherTok{\textless{}{-}} \FunctionTok{read.csv}\NormalTok{(}\StringTok{"combined\_raven\_detection.csv"}\NormalTok{)}

\FunctionTok{diagnose\_detection}\NormalTok{(}\AttributeTok{reference =}\NormalTok{ manual\_ref\_tae, }\AttributeTok{detection =}\NormalTok{ raven\_detec, }\AttributeTok{by =} \StringTok{"tuning\_parameters"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{raven\_detec }\OtherTok{\textless{}{-}} \FunctionTok{read.csv}\NormalTok{(}\StringTok{"./data/processed/combined\_raven\_detection.csv"}\NormalTok{)}

\FunctionTok{diagnose\_detection}\NormalTok{(}\AttributeTok{reference =}\NormalTok{ manual\_ref\_tae, }\AttributeTok{detection =}\NormalTok{ raven\_detec, }\AttributeTok{by =} \StringTok{"tuning\_parameters"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção
\end{verbatim}

\begin{table}
\centering
\begin{tabular}[t]{c|c|c|c|c|c|c|c|c|c|c}
\hline
tuning\_parameters & total.detections & true.positives & false.positives & false.negatives & split.positives & merged.positives & overlap.to.true.positives & recall & precision & f1.score\\
\hline
band: 1-10 kHz; sep: 0.005 s & 489 & 489 & 85 & 107 & 33 & 99 & 0.911 & 0.820 & 0.852 & 0.836\\
\hline
band: 1-10 kHz; sep: 0.02322 s & 79 & 79 & 76 & 512 & 0 & 556 & 0.994 & 0.134 & 0.510 & 0.212\\
\hline
band: 1-15 kHz; sep: 0.02322 s & 430 & 430 & 202 & 161 & 21 & 236 & 0.927 & 0.728 & 0.680 & 0.703\\
\hline
band: 1-22 kHz; sep: 0.02322.txt s & 494 & 494 & 526 & 101 & 51 & 153 & 0.870 & 0.830 & 0.484 & 0.612\\
\hline
\end{tabular}
\end{table}

The output shows the indices described above. Plus three additional
metrics specific to sound event detection: `split positives', `merge
positives' and `overlap to true positives'. `Split positives' is the
number of reference sounds overlapped by more than one detection, `merge
positives' the number of detected sounds overlapping with another
detection and `overlap to true positives' quantifies the mean overlap
between detections and reference signal (1 means complete overlap). The
function also allows to detail those indices by sound file. Here we show
the first 10 files:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{diag\_raven }\OtherTok{\textless{}{-}} \FunctionTok{diagnose\_detection}\NormalTok{(}\AttributeTok{reference =}\NormalTok{ manual\_ref\_tae, }\AttributeTok{detection =}\NormalTok{ raven\_detec, }\AttributeTok{by =} \StringTok{"tuning\_parameters"}\NormalTok{, }\AttributeTok{by.sound.file =} \ConstantTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{head}\NormalTok{(diag\_raven, }\DecValTok{10}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}
\centering
\begin{tabular}[t]{c|c|c|c|c|c|c|c|c|c|c|c}
\hline
tuning\_parameters & sound.files & total.detections & true.positives & false.positives & false.negatives & split.positives & merged.positives & overlap.to.true.positives & recall & precision & f1.score\\
\hline
band: 1-10 kHz; sep: 0.005 s & Ag13\_43421.27975590\_11\_17\_7\_46\_15.wav & 40 & 35 & 5 & 0 & 0 & 0 & 0.988 & 1.000 & 0.875 & 0.933\\
\hline
band: 1-10 kHz; sep: 0.005 s & BRN7\_43435.27985312\_12\_1\_7\_46\_25.wav & 34 & 27 & 7 & 24 & 0 & 21 & 0.980 & 0.529 & 0.794 & 0.635\\
\hline
band: 1-10 kHz; sep: 0.005 s & Blk109Brn\_43559.32349131\_4\_4\_8\_59\_9.wav & 30 & 29 & 1 & 5 & 0 & 6 & 0.920 & 0.853 & 0.967 & 0.906\\
\hline
band: 1-10 kHz; sep: 0.005 s & DB118\_43568.33139566\_4\_13\_9\_12\_19.wav & 18 & 18 & 0 & 1 & 0 & 2 & 0.989 & 0.947 & 1.000 & 0.973\\
\hline
band: 1-10 kHz; sep: 0.005 s & DB15HP\_43450.29217192\_12\_16\_8\_6\_57.wav & 23 & 21 & 2 & 0 & 0 & 0 & 0.989 & 1.000 & 0.913 & 0.955\\
\hline
band: 1-10 kHz; sep: 0.005 s & DB7\_43357.58119484\_9\_14\_16\_8\_39.wav & 18 & 18 & 0 & 1 & 0 & 2 & 0.941 & 0.947 & 1.000 & 0.973\\
\hline
band: 1-10 kHz; sep: 0.005 s & DG124DB\_43559.30160960\_4\_4\_8\_22\_40.wav & 39 & 29 & 10 & 0 & 0 & 0 & 0.986 & 1.000 & 0.744 & 0.853\\
\hline
band: 1-10 kHz; sep: 0.005 s & GRY37HP\_43442.27431670\_12\_8\_7\_37\_11.wav & 54 & 48 & 6 & 6 & 0 & 8 & 0.936 & 0.889 & 0.889 & 0.889\\
\hline
band: 1-10 kHz; sep: 0.005 s & Gold183\_43555.5549813\_3\_31\_1\_32\_29.wav & 2 & 2 & 0 & 0 & 0 & 0 & 1.000 & 1.000 & 1.000 & 1.000\\
\hline
band: 1-10 kHz; sep: 0.005 s & Gry35HP\_43455.29800260\_12\_21\_8\_16\_40.wav & 18 & 18 & 0 & 8 & 0 & 13 & 0.995 & 0.692 & 1.000 & 0.818\\
\hline
\end{tabular}
\end{table}

Diagnostics from routines using different tuning parameters can be used
to identify the parameter values that optimize detection. These process
of evaluating different routines for detection optimization is
incorporated into the two signal detection approaches provided natively
by \emph{ohun}, which we depict in the following section. Note that the
detection with Raven Pro does not necessarily reflect the best
performance of this software and has been included only as an example on
evaluating detection from external sources, rather than a direct
comparison of performance between Raven Pro and \emph{ohun}.

\hypertarget{signal-detection-with-ohun}{%
\subsection{\texorpdfstring{Signal detection with
\emph{ohun}}{Signal detection with ohun}}\label{signal-detection-with-ohun}}

The package offers two methods for automated signal detection:
template-based and energy-based detection. These methods are better
suited for highly stereotyped or good signal-to-noise ratio (SNR)
sounds, respectively. If the target sounds do not fit these
requirements, more elaborate methods (i.e.~machine/deep learning
approaches) are warranted.

\hypertarget{study-cases}{%
\subsection{Study cases}\label{study-cases}}

\hypertarget{template-detection-on-ultrasonic-social-calls-of-spixs-disc-winged-bats}{%
\subsubsection{Template detection on ultrasonic social calls of Spix's
disc-winged
bats}\label{template-detection-on-ultrasonic-social-calls-of-spixs-disc-winged-bats}}

We recorded 30 individuals of Spix's disc-winged bats (\emph{Thyroptera
tricolor}) at Baru Biological Station, in south-western Costa Rica in
January 2020. Bats were captured at their roosting sites (furled leaves
of Zingeberaceae plants). Each individual bat was released in a large
flight cage (9 x 4 x 3 m) for a 5-minute period and their ultrasonic
inquiry calls were recorded using a condenser microphone (CM16, Avisoft
Bioacoustics, Glienike/Nordbahn, Germany) through an Avisoft
UltraSoundGate 116Hm plugged into a laptop computer running
Avisoft-Recorder software. Recordings were made at a sampling rate of
500 kHz and an amplitude resolution of 16 bits.

Recordings were manually annotated using Raven Pro 1.6
(\protect\hyperlink{ref-Charif2010}{Charif \emph{et al.} 2010}).
Annotations were created by visual inspection of spectrograms, in which
the start and end of sounds were determined by the location of the
continuous traces of power spectral entropy of the target sounds. A
total of 644 calls were annotated (\textasciitilde21 calls per
recording). Annotations were made with a time window of 200 samples and
70\% of overlap, and were then imported into R using the package Rraven
(\protect\hyperlink{ref-Araya-Salas2017e}{Araya-Salas 2020}).

Inquiry calls of Spix's disc-winged bats are structurally stereotyped
(\protect\hyperlink{ref-Chaverri2010b}{Chaverri \emph{et al.} 2010}).
Most variation is found among individuals, although the basic form of a
short, downward broadband frequency modulation is always shared (Fig.
BAT-SPECTRO, Araya-Salas et al 2021 ontogeny).

\begin{verbatim}
## Error in knitr::include_graphics(file.path(rootdir, "data/raw/thyroptera/Catalog_p1.jpeg")): Cannot find the file(s): "../data/raw/thyroptera/Catalog_p1.jpeg"
\end{verbatim}

Template-based detection is a useful approach when there are minimal
structural differences in the target sounds
(\protect\hyperlink{ref-knight2017}{Knight \emph{et al.} 2017};
e.g.~when signals are produced in a highly stereotyped manner,
\protect\hyperlink{ref-balantic2020}{Balantic \& Donovan 2020}). It uses
spectrographic cross-correlation to find sounds resembling an example
target sounds (i.e.~template) across sound files, applying a correlation
threshold to separate detections from background noise. We used this
approach in \emph{ohun} to detect inquiry calls. To do this, we tested
the performance of three acoustic templates on a training subset of five
sound files. First we used the function \texttt{get\_templates} to find
several sounds representative of the variation in signal structure. This
function measures several spectral parameters which are then summarized
using Principal Component Analysis. The first two components are used to
project the acoustic space. In this space the function defines
sub-spaces as equal-size slices of a sphere centered at the centroid of
the acoustic space. Templates are then selected as those closer to the
centroid within each of the sub-spaces, including the centroid for the
entire acoustic space. The user needs to define the number of sub-spaces
in which the acoustic space will be split.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# read manual annotations}
\NormalTok{manual\_ref\_thy }\OtherTok{\textless{}{-}} \FunctionTok{read.csv}\NormalTok{(}\StringTok{"manual\_annotations\_thyroptera.csv"}\NormalTok{)}

\CommentTok{\# get random subset of 5 sound files for training}
\FunctionTok{set.seed}\NormalTok{(}\DecValTok{1}\NormalTok{)}
\NormalTok{train\_files }\OtherTok{\textless{}{-}} \FunctionTok{sample}\NormalTok{(}\FunctionTok{unique}\NormalTok{(manual\_ref\_thy}\SpecialCharTok{$}\NormalTok{sound.files), }\AttributeTok{size =} \DecValTok{5}\NormalTok{)}
\NormalTok{train\_ref }\OtherTok{\textless{}{-}}\NormalTok{ manual\_ref\_thy[manual\_ref\_thy}\SpecialCharTok{$}\NormalTok{sound.files }\SpecialCharTok{\%in\%}\NormalTok{ train\_files, ]}

\CommentTok{\# the rest for testin g}
\NormalTok{test\_files }\OtherTok{\textless{}{-}} \FunctionTok{setdiff}\NormalTok{(manual\_ref\_thy}\SpecialCharTok{$}\NormalTok{sound.files, train\_files)}
\NormalTok{test\_ref }\OtherTok{\textless{}{-}}\NormalTok{ manual\_ref\_thy[manual\_ref\_thy}\SpecialCharTok{$}\NormalTok{sound.files }\SpecialCharTok{\%in\%}\NormalTok{ test\_files, ]}

\CommentTok{\# find templates}
\NormalTok{templates }\OtherTok{\textless{}{-}} \FunctionTok{get\_templates}\NormalTok{(train\_ref, }\AttributeTok{path =}\NormalTok{ data\_path,  }\AttributeTok{bp =} \FunctionTok{c}\NormalTok{(}\DecValTok{10}\NormalTok{, }\DecValTok{50}\NormalTok{), }\AttributeTok{ovlp =} \DecValTok{70}\NormalTok{, }\AttributeTok{wl =} \DecValTok{200}\NormalTok{, }\AttributeTok{n.sub.spaces =} \DecValTok{3}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## The first 2 principal components explained 0.5 of the variance
\end{verbatim}

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/unnamed-chunk-12-1.pdf}
\caption{FIG. ACOUSTIC-SPACE. Acoustic space defined as the the first
two components of a Principal Component Analysis on spectrographic
parameters. Templates are selected as those closer to the centroid for
each of the sub-spaces. Gray dashed lines delimit the region of
sub-spaces. Yellow circles around points highlight the position of the
signals selected as templates.}
\end{figure}

The output of the \texttt{get\_templates} function includes an acoustic
space plot (FIG. ACOUSTIC-SPACE) in which the position of the sounds
selected as templates is highlighted. Users can also provide their own
acoustic space dimensions (argument `acoustic.space'). In the following
code we used those templates for detecting bat social calls. The code
iterates a template-based detection on the training data set across a
range of correlation thresholds, in order to find the combination of
threshold and template with the best performance.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# get correlation vectors}
\NormalTok{corr\_templ\_train }\OtherTok{\textless{}{-}} \FunctionTok{template\_correlator}\NormalTok{(}
  \AttributeTok{templates =}\NormalTok{ templates,}
    \AttributeTok{path =}\NormalTok{ data\_path, }
  \AttributeTok{files =} \FunctionTok{unique}\NormalTok{(train\_ref}\SpecialCharTok{$}\NormalTok{sound.files), }
  \AttributeTok{hop.size =} \DecValTok{10}\NormalTok{, }
  \AttributeTok{ovlp =} \DecValTok{70}
\NormalTok{  )}

\CommentTok{\# evaluate detection for different correlation thresholds}
\NormalTok{opt\_detec\_train }\OtherTok{\textless{}{-}} \FunctionTok{optimize\_template\_detector}\NormalTok{(}
  \AttributeTok{reference =}\NormalTok{ train\_ref, }
  \AttributeTok{template.correlations =}\NormalTok{ corr\_templ\_train,}
  \AttributeTok{threshold =} \FunctionTok{seq}\NormalTok{(}\FloatTok{0.05}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\FloatTok{0.01}\NormalTok{)}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

Note that the correlation vectors are estimated first (i.e.~vectors of
correlation values across sound files, \texttt{template\_correlator()})
and then the correlation thresholds are optimized on these vectors
(\texttt{optimize\_template\_detector()}). The output of
\texttt{optimize\_template\_detector()} contains the detection
performance indices for each combination of templates and thresholds.
Table TEMPLATE PERFORMANCE shows the two highest performance runs for
each template.

\begin{table}

\caption{\label{tab:unnamed-chunk-15}TABLE TEMPLATE PERFORMANCE. Performance diagnostic of template-based detections using four templates across several threshold values. Only the two highest performance iterations for each template are shown.}
\centering
\begin{tabular}[t]{r|l|r|r|r|r|r|r}
\hline
threshold & templates & true.positives & false.positives & false.negatives & recall & precision & f1.score\\
\hline
0.45 & centroid & 81 & 1 & 0 & 1.000 & 0.988 & 0.994\\
\hline
0.50 & centroid & 80 & 0 & 1 & 0.988 & 1.000 & 0.994\\
\hline
0.50 & templ-1 & 77 & 1 & 4 & 0.951 & 0.987 & 0.969\\
\hline
0.45 & templ-1 & 80 & 5 & 1 & 0.988 & 0.941 & 0.964\\
\hline
0.40 & templ-2 & 79 & 0 & 2 & 0.975 & 1.000 & 0.988\\
\hline
0.35 & templ-2 & 80 & 2 & 1 & 0.988 & 0.976 & 0.982\\
\hline
0.45 & templ-3 & 77 & 2 & 4 & 0.951 & 0.975 & 0.963\\
\hline
0.40 & templ-3 & 79 & 5 & 2 & 0.975 & 0.940 & 0.958\\
\hline
\end{tabular}
\end{table}

We can explore the performance of each template in more detail by
looking at the change in F1 score across thresholds (FIG. THRESHOLD vs
F1.SCORE).

\begin{figure}

{\centering \includegraphics{manuscript_files/figure-latex/unnamed-chunk-16-1}

}

\caption{FIG. THRESHOLD vs F1.SCORE. Shows the changes in F1 score across the range of threshold values}\label{fig:unnamed-chunk-16}
\end{figure}

In this example the ``centroid'' template, produced the best performance
(TABLE TEMPLATE PERFORMANCE; FIG. THRESHOLD vs F1.SCORE). Hence, we will
use this template for detecting calls on the rest of the data. The
following code extracts this template from the reference annotation
table and use it to find inquiry calls on the testing data set:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# get correlation vectors for test files}
\NormalTok{corr\_templ\_test }\OtherTok{\textless{}{-}} \FunctionTok{template\_correlator}\NormalTok{(}
  \AttributeTok{templates =}\NormalTok{ templates[templates}\SpecialCharTok{$}\NormalTok{sound.file }\SpecialCharTok{==} \StringTok{"centroid"}\NormalTok{, ],}
    \AttributeTok{path =}\NormalTok{ data\_path, }\AttributeTok{files =} \FunctionTok{unique}\NormalTok{(test\_ref}\SpecialCharTok{$}\NormalTok{sound.files), }
  \AttributeTok{hop.size =} \DecValTok{10}\NormalTok{, }
  \AttributeTok{ovlp =} \DecValTok{70}
\NormalTok{  )}

\CommentTok{\# detect on test files}
\NormalTok{detec\_test }\OtherTok{\textless{}{-}} \FunctionTok{template\_detector}\NormalTok{(}
  \AttributeTok{template.correlations =}\NormalTok{ corr\_templ\_test,}
    \AttributeTok{threshold =} \FloatTok{0.45}
\NormalTok{  )}

\FunctionTok{diagnose\_detection}\NormalTok{(}\AttributeTok{reference =}\NormalTok{ test\_ref, }\AttributeTok{detection =}\NormalTok{ detec\_test)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção
\end{verbatim}

\begin{table}
\centering
\begin{tabular}[t]{c|c|c|c|c|c|c|c|c|c}
\hline
total.detections & true.positives & false.positives & false.negatives & split.positives & merged.positives & overlap.to.true.positives & recall & precision & f1.score\\
\hline
522 & 522 & 10 & 41 & 2 & 32 & 0.907 & 0.927 & 0.981 & 0.953\\
\hline
\end{tabular}
\end{table}

The last line of codes evaluates the detection on the test data set,
which shows a good performance for both recall and precision.

\hypertarget{energy-based-detection-on-zebra-finch-vocalizations}{%
\subsubsection{Energy-based detection on zebra finch
vocalizations}\label{energy-based-detection-on-zebra-finch-vocalizations}}

We used recordings from 18 zebra finch males recorded at obtained from
the Rockefeller University Field Research Center Song Library
(\url{http://ofer.sci.ccny.cuny.edu/songs},
\protect\hyperlink{ref-Tchernichovski2021}{Tchernichovski \emph{et al.}
2021}). Recordings contain undirected vocalizations (songs or calls) of
single males recorded in sound attenuation chambers using Sound Analysis
Pro. Zebra finch vocalizations are composed of multiple elements
(i.e.~distinct patterns of continuous traces of power spectral entropy
in the spectrogram separated by time gaps) that can vary importantly in
key features as duration and frequency range (ZEBRAFINCH-SPECTRO) and
are clearly not as stereotyped as the Spix's disc-winged bats. However,
as the recorded sounds show a good signal-to-noise ratio, such that
signals in each recording can potentially be detected using an
energy-based approach that does not rely on matching the acoustic
structure of a template.

\begin{verbatim}
## Error in knitr::include_graphics(file.path(rootdir, "data/raw/taeniopygia/same.time-Catalog_p1.jpeg")): Cannot find the file(s): "../data/raw/taeniopygia/same.time-Catalog_p1.jpeg"
\end{verbatim}

Reference annotations were made manually on the oscillogram with the
spectrogram and audio as a guide using Raven Lite 2.0.1 (Cornell Lab of
Ornithology). The following code loads the reference annotations and
split them into two data sets for training (3 sound files) and testing
(15 sound files):

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{manual\_ref\_tae }\OtherTok{\textless{}{-}} \FunctionTok{read.csv}\NormalTok{(}\StringTok{"manual\_selections\_Taeniopygia.csv"}\NormalTok{)}

\FunctionTok{set.seed}\NormalTok{(}\DecValTok{450}\NormalTok{)    }
\NormalTok{train\_files }\OtherTok{\textless{}{-}} \FunctionTok{sample}\NormalTok{(}\FunctionTok{unique}\NormalTok{(manual\_ref\_tae}\SpecialCharTok{$}\NormalTok{sound.files), }\DecValTok{3}\NormalTok{)    }
\NormalTok{test\_files }\OtherTok{\textless{}{-}} \FunctionTok{setdiff}\NormalTok{(manual\_ref\_tae}\SpecialCharTok{$}\NormalTok{sound.files, train\_files)}

\NormalTok{train\_ref }\OtherTok{\textless{}{-}}\NormalTok{ manual\_ref\_tae[manual\_ref\_tae}\SpecialCharTok{$}\NormalTok{sound.files }\SpecialCharTok{\%in\%}\NormalTok{ train\_files, ]}
\NormalTok{test\_ref }\OtherTok{\textless{}{-}}\NormalTok{ manual\_ref\_tae[manual\_ref\_tae}\SpecialCharTok{$}\NormalTok{sound.files }\SpecialCharTok{\%in\%}\NormalTok{ test\_files, ]}
\end{Highlighting}
\end{Shaded}

The detection parameters can be optimized using the function
\texttt{optimize\_energy\_detector}. This function runs a detection for
all possible combinations of tuning parameters. The code below tries
three minimum duration and maximum duration values and two hold time
values (which merges sounds within the specified time interval):

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{opt\_det\_train }\OtherTok{\textless{}{-}} \FunctionTok{optimize\_energy\_detector}\NormalTok{(}
  \AttributeTok{reference =}\NormalTok{ train\_ref, }
  \AttributeTok{files =}\NormalTok{ train\_files, }
  \AttributeTok{threshold =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{5}\NormalTok{), }
  \AttributeTok{hop.size =} \FloatTok{11.6}\NormalTok{, }
  \AttributeTok{smooth =} \FunctionTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\DecValTok{10}\NormalTok{), }
  \AttributeTok{hold.time =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{5}\NormalTok{), }
  \AttributeTok{min.duration =} \FunctionTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\DecValTok{15}\NormalTok{, }\DecValTok{25}\NormalTok{), }
  \AttributeTok{max.duration =} \FunctionTok{c}\NormalTok{(}\DecValTok{275}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{325}\NormalTok{), }
  \AttributeTok{bp =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.5}\NormalTok{, }\DecValTok{10}\NormalTok{)}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

The output (\texttt{opt\_det\_train}) shows the performance indices for
each of those combinations. Hhere we show the 10 combinations with the
highest F1 score:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# subset with highest performance}
\NormalTok{opt\_det\_train }\OtherTok{\textless{}{-}}\NormalTok{ opt\_det\_train[}\FunctionTok{order}\NormalTok{(opt\_det\_train}\SpecialCharTok{$}\NormalTok{f1.score, }\AttributeTok{decreasing =} \ConstantTok{TRUE}\NormalTok{), ]}

\FunctionTok{head}\NormalTok{(opt\_det\_train, }\DecValTok{10}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}
\centering
\begin{tabular}[t]{c|c|c|c|c|c|c|c|c|c|c}
\hline
threshold & smooth & hold.time & min.duration & max.duration & true.positives & false.positives & false.negatives & recall & precision & f1.score\\
\hline
1 & 5 & 5 & 25 & 300 & 105 & 9 & 0 & 1.000 & 0.921 & 0.959\\
\hline
1 & 5 & 5 & 25 & 325 & 105 & 9 & 0 & 1.000 & 0.921 & 0.959\\
\hline
1 & 5 & 0 & 25 & 300 & 105 & 12 & 0 & 1.000 & 0.897 & 0.946\\
\hline
1 & 5 & 0 & 25 & 325 & 105 & 12 & 0 & 1.000 & 0.897 & 0.946\\
\hline
1 & 10 & 0 & 25 & 300 & 105 & 13 & 0 & 1.000 & 0.890 & 0.942\\
\hline
1 & 10 & 0 & 25 & 325 & 105 & 13 & 0 & 1.000 & 0.890 & 0.942\\
\hline
1 & 5 & 5 & 15 & 300 & 105 & 15 & 0 & 1.000 & 0.875 & 0.933\\
\hline
1 & 5 & 5 & 15 & 325 & 105 & 15 & 0 & 1.000 & 0.875 & 0.933\\
\hline
5 & 5 & 5 & 15 & 275 & 96 & 8 & 9 & 0.914 & 0.923 & 0.919\\
\hline
5 & 5 & 5 & 15 & 300 & 96 & 8 & 9 & 0.914 & 0.923 & 0.919\\
\hline
\end{tabular}
\end{table}

We now can use the tuning parameter values that yielded the best
performance to detect sounds on the test data set:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{best\_param }\OtherTok{\textless{}{-}}\NormalTok{ opt\_det\_train[}\FunctionTok{which.max}\NormalTok{(opt\_det\_train}\SpecialCharTok{$}\NormalTok{f1.score), ]}

\NormalTok{det\_test }\OtherTok{\textless{}{-}} \FunctionTok{energy\_detector}\NormalTok{(}
  \AttributeTok{files =}\NormalTok{ test\_files, }
  \AttributeTok{threshold =}\NormalTok{ best\_param}\SpecialCharTok{$}\NormalTok{threshold, }
  \AttributeTok{hop.size =} \FloatTok{11.6}\NormalTok{, }
  \AttributeTok{smooth =}\NormalTok{ best\_param}\SpecialCharTok{$}\NormalTok{smooth, }
  \AttributeTok{hold.time =}\NormalTok{ best\_param}\SpecialCharTok{$}\NormalTok{hold.time, }
  \AttributeTok{min.duration =}\NormalTok{ best\_param}\SpecialCharTok{$}\NormalTok{min.duration, }
  \AttributeTok{max.duration =}\NormalTok{ best\_param}\SpecialCharTok{$}\NormalTok{max.duration, }
  \AttributeTok{bp =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.5}\NormalTok{, }\DecValTok{10}\NormalTok{)}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

As our reference annotations include all sound files we can evaluate the
performance of the detection on the test set as well:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{diagnose\_detection}\NormalTok{(}\AttributeTok{reference =}\NormalTok{ test\_ref, }\AttributeTok{detection =}\NormalTok{ det\_test, }\AttributeTok{by.sound.file =} \ConstantTok{FALSE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção
\end{verbatim}

\begin{table}
\centering
\begin{tabular}[t]{c|c|c|c|c|c|c|c|c|c}
\hline
total.detections & true.positives & false.positives & false.negatives & split.positives & merged.positives & overlap.to.true.positives & recall & precision & f1.score\\
\hline
463 & 463 & 27 & 27 & 4 & 32 & 0.97 & 0.945 & 0.945 & 0.945\\
\hline
\end{tabular}
\end{table}

~

The performance on the test data set was also acceptable with a F1 score
of 0.95. Note that in the example we used a small subset of sound files
for training. More training data might be needed for optimizing a
detection routine on larger data sets or recordings with more variable
sounds or background noise levels.

\hypertarget{additional-tools}{%
\subsubsection{Additional tools}\label{additional-tools}}

The \emph{ohun} package offers additional tools to simplify sound event
detection. Detected sounds can be labeled as false or true positives
with the function `label\_detection'. This allows users to explore the
structure of false positives and figure out ways to exclude them. The
function `filter\_detections' can be used to remove ambiguous sounds
(i.e.~those labeled as split or merged detections) keeping only those
that maximize a specific criterion (i.e.~the highest template
correlation). Finally, note that several templates representing the
range of variation in signal structure can be used to detect
semi-stereotyped sounds when running template-based detection
(`template\_detection' function).

\hypertarget{discusion}{%
\subsection{Discusion}\label{discusion}}

Here we have shown how to evaluate the performance of sound event
detection routines using the package \emph{ohun}. The package can
evaluate detection outputs imported from other software, as well as its
own detection routines. The latter can be iterated over combinations of
tuning parameters in order to find those values that optimize detection.
Despite signal detection indices being commonly reported when presenting
new automatic detection methods, to our knowledge widely applicable
performance evaluating routines have not been made available in a free,
open source platform. Providing a common framework for the evaluation of
sound event detection can simplify comparing the performance of
different tools and the selection of those tools better suited to a
given research question and study system. The tools offered by
\emph{ohun} for diagnosing detection performance should not be
necessarily limited to acoustic data. \emph{ohun} can also be used for
cases in which the time of occurrence of discrete events needs to be
identified such as detecting specific behaviors in video analysis of
animal motor activity (\emph{e.g.}
\protect\hyperlink{ref-sturman2020}{Sturman \emph{et al.} 2020};
\protect\hyperlink{ref-hsu2021}{Hsu \& Yttri 2021};
\protect\hyperlink{ref-bohnslav2021}{Bohnslav \emph{et al.} 2021}). The
detection of such motor events in video can also be evaluated and
optimized in comparison to a reference annotation, as we have shown
above for acoustic signals.

The \emph{ohun} package provides two native detection methods:
template-based and energy-based detection. Compared to new deep learning
approaches for finding the occurrence of sound events, the two native
methods are relatively simple tools. However, these methods have been
widely used by the bioacoustic community
(\protect\hyperlink{ref-mellinger2000}{Mellinger \& Clark 2000};
\protect\hyperlink{ref-Charif2010}{Charif \emph{et al.} 2010};
\protect\hyperlink{ref-aide2013}{Aide \emph{et al.} 2013};
\protect\hyperlink{ref-avi2014}{Bioacoustics 2014};
\protect\hyperlink{ref-Hafner2015}{Hafner \& Katz 2015}) and can reach
adequate performance under the appropriate conditions, as we have shown
in our two study cases and has been shown elsewhere
(\protect\hyperlink{ref-knight2017}{Knight \emph{et al.} 2017}). Deep
learning methods tend to require greater computational power, larger
training data sets and, in some cases, more complex training routines.
This might bring unnecessary difficulties when dealing with when less
challenging detection tasks. Therefore the availability of a wide range
of approaches can simplify finding the most appropriate tool for the
intricacies of a study system and research goals, as well as having
tools accessible for a broader community of researchers. The tools
offered in \emph{ohun} can also be used in a subsequent pipeline in
which detected sounds are further classified using more elaborated
discrimination algorithms (\protect\hyperlink{ref-balantic2020}{Balantic
\& Donovan 2020}). In addition, detection performance can be improved by
using acoustic structure measurements that can be used to distinguish
target from non-target sound events.

Detection routines can take a long time when working with large amounts
of acoustic data (e.g.~large recordings and/or many files). We provide
some additional tips that can help make a routine more time-efficient.
1. Always test procedures on small data subsets. Make sure that you
obtain decent results on a small subset of recordings before trying to
scale up the analysis. 2. Template-based detection is almost always
faster than energy-based detection. 3. Run routines in parallel.
Parallelization (i.e.~the ability to distribute tasks over several cores
in your computer) can significantly speed-up routines. All functions for
automatic detection and performance evaluation allow users to run
analysis in parallel (see \texttt{parallel} argument in those
functions). Hence a computer with several cores can be helpful for
improving efficiency. 4. Sampling rate matters. Detecting sounds on low
sampling rate files is faster, so we must avoid having Nyquist
frequencies much higher than the highest frequency of the target sounds.
5. Try using a computer with lots of RAM memory or a computer cluster
for working on large amounts of data. Lastly, we underscore that these
tips are not restricted to \emph{ohun} and can also be helpful to speed
up routines in other software packages.

There are some additional things to be considered when aiming to
automatically detect sound events. When running energy-based detection
routines try to use your knowledge about the signal structure to
determine the initial range of tunning parameters. This can be extremely
helpful for narrowing down possible parameter values. As a general rule,
if human observers have a hard time detecting where a target sound
occurs in a sound file, detection algorithms will likely yield low
detection performance. In cases in which occurrences are ambiguous low
performances are expected. Ensure reference annotations contain all
target sounds and only the target sounds, otherwise performance
optimization can be misleading as the performance of a given detection
method cannot be better than the reference itself. Lastly, avoid having
overlapping sounds or several sounds as a single detection (like a
multi-syllable vocalization) in the reference annotation when running an
energy-based detector, as they are likely to be identified as separated
units.

\hypertarget{acknowledgements}{%
\subsubsection{Acknowledgements}\label{acknowledgements}}

Nazareth Rojas, Silvia Chaves-Ramírez, Mariela Sánchez-Chavarría, Miriam
Gioiosa, Cristian Castillo-Salazar, and José Pablo Barrantes for their
help collecting acoustic data for Spix's disc-winged bats and Jorge
Elizondo for his help on manually annotating recordings. We also thank
the Centro Biológico Hacienda Barú for their continuous support of our
research and Mijail Rojas and Andrey Sequeira for support on early
stages of package development. This study was partly funded by a
CONARE-Max Planck grant from the Consejo Nacional de Rectores and the
research activity C0754 at Centro de Investigación en Neurociencias,
Universidad de Costa Rica. A.R-G. is supported by the Walt Halperin
Endowed Professorship and the Washington Research Foundation as
Distinguished Investigator.

\hypertarget{ethical-note}{%
\subsubsection{Ethical note}\label{ethical-note}}

All sampling protocols followed guidelines approved by the American
Society of Mammalogists for capture, handling and care of mammals
(\protect\hyperlink{ref-Sikes2016}{Sikes \emph{et al.} 2016}) and the
ASAB/ABS Guidelines for the use of animals in research. This study was
conducted in accordance with the ethical standards for animal welfare of
the Costa Rican Ministry of Environment and Energy, Sistema Nacional de
Áreas de Conservación, permit no. SINAC-ACOPAC-RES-INV-008-2017 (Decree
No.~32553-MINAE). Protocols were also approved by the University of
Costa Rica's Institutional Animal Care and Use Committee
(CICUA-42-2018).

\hypertarget{supporting-information}{%
\subsubsection{Supporting Information}\label{supporting-information}}

Supplementary information associated with this article is available in
XXXXXXXXXXXXXXXXX

\hypertarget{references}{%
\subsection*{References}\label{references}}
\addcontentsline{toc}{subsection}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-aide2013}{}}%
Aide, T.M., Corrada-Bravo, C., Campos-Cerqueira, M., Milan, C., Vega, G.
\& Alvarez, R. (2013).
\href{https://doi.org/10.7717/PEERJ.103}{Real-time bioacoustics
monitoring and automated species identification}. \emph{PeerJ},
\textbf{2013}.

\leavevmode\vadjust pre{\hypertarget{ref-Araya-Salas2017e}{}}%
Araya-Salas, M. (2020). Rraven: Connecting r and raven bioacoustic
software. R package version 1.0.9.

\leavevmode\vadjust pre{\hypertarget{ref-Araya-Salas2017f}{}}%
Araya-Salas, M. \& Smith-Vidaurre, G. (2017). warbleR: An r package to
streamline analysis of animal acoustic signals. \emph{Methods in Ecology
and Evolution}, \textbf{8}, 184--191. Retrieved from
\url{http://onlinelibrary.wiley.com/doi/10.1111/2041-210X.12624/abstract}

\leavevmode\vadjust pre{\hypertarget{ref-balantic2020}{}}%
Balantic, C.M. \& Donovan, T.M. (2020). Statistical learning mitigation
of false positives from template-detected data in automated acoustic
wildlife monitoring. \emph{BIoacoustics}, \textbf{29}, 293--321.
Retrieved from
\url{https://www.tandfonline.com/action/journalInformation?journalCode=tbio20}

\leavevmode\vadjust pre{\hypertarget{ref-avi2014}{}}%
Bioacoustics, A. (2014). Avisoft SasLab pro.

\leavevmode\vadjust pre{\hypertarget{ref-Blumstein2011}{}}%
Blumstein, D.T., Mennill, D.J., Clemins, P., Girod, L., Yao, K.,
Patricelli, G., Deppe, J.L., Krakauer, A.H., Clark, C., Cortopassi,
K.A., Hanser, S.F., Mccowan, B., Ali, A.M. \& Kirschel, A.N.G. (2011).
\href{https://doi.org/10.1111/j.1365-2664.2011.01993.x}{Acoustic
monitoring in terrestrial environments using microphone arrays:
Applications, technological considerations and prospectus}.
\emph{Journal of Applied Ecology}, \textbf{48}, 758--767.

\leavevmode\vadjust pre{\hypertarget{ref-bohnslav2021}{}}%
Bohnslav, J.P., Wimalasena, N.K., Clausing, K.J., Dai, Y.Y.,
Yarmolinsky, D.A., Cruz, T., Kashlan, A.D., Chiappe, M.E., Orefice,
L.L., Woolf, C.J. \& Harvey, C.D. (2021).
\href{https://doi.org/10.7554/ELIFE.63377}{DeepEthogram, a machine
learning pipeline for supervised behavior classification from raw
pixels}. \emph{eLife}, \textbf{10}.

\leavevmode\vadjust pre{\hypertarget{ref-Burgdorf2011}{}}%
Burgdorf, J., Panksepp, J. \& Moskal, J.R. (2011).
\href{https://doi.org/10.1016/j.neubiorev.2010.11.011}{Frequency-modulated
50kHz ultrasonic vocalizations: A tool for uncovering the molecular
substrates of positive affect}. \emph{Neuroscience and Biobehavioral
Reviews}, \textbf{35}, 1831--1836.

\leavevmode\vadjust pre{\hypertarget{ref-Charif2010}{}}%
Charif, R., Waack, A. \& Strickman, L. (2010). \emph{Raven pro 1.4
user's manual.} Cornell Lab of Ornithology.

\leavevmode\vadjust pre{\hypertarget{ref-Chaverri2010b}{}}%
Chaverri, G., Gillam, E.H. \& Vonhof, M.J. (2010).
\href{https://doi.org/10.1098/rsbl.2009.0964}{Social calls used by a
leaf-roosting bat to signal location}. \emph{Biology Letters},
\textbf{6}, 441--444.

\leavevmode\vadjust pre{\hypertarget{ref-gibb2019}{}}%
Gibb, R., Browning, E., Glover-Kapfer, P., Jones, K.E. \& Jones, C.E.K.
(2019). Emerging opportunities and challenges for passive acoustics in
ecological assessment and monitoring. \emph{Wiley Online Library},
\textbf{10}, 169--185. Retrieved from
\url{https://besjournals.onlinelibrary.wiley.com/doi/abs/10.1111/2041-210X.13101}

\leavevmode\vadjust pre{\hypertarget{ref-gew2019}{}}%
Gwee, C., Eaton, J., Garg, K. \& Alström, P. (2019). Cryptic diversity
in cyornis (aves: Muscicapidae) jungle-flycatchers flagged by simple
bioacoustic approaches. \emph{Zoological Journal}, \textbf{186}.
Retrieved from
\url{https://academic.oup.com/zoolinnean/article-abstract/186/3/725/5372980}

\leavevmode\vadjust pre{\hypertarget{ref-Hafner2015}{}}%
Hafner, S. \& Katz, J. (2015). monitoR: Acoustic template detection in
r. \emph{R package version 1.0.3}. Retrieved from
\href{http://www.uvm.edu/rsenr/vtcfwru/R/?Page=monitoR/monitoR.htmURL:\%20http://www.uvm.edu/rsenr/vtcfwru/R/?Page=monitoR/monitoR.htmURL:\%20http://www.uvm.edu/rsenr/vtcfwru/R/?Page=monitoR/monitoR.htm}{http://www.uvm.edu/rsenr/vtcfwru/R/?Page=monitoR/monitoR.htmURL:
http://www.uvm.edu/rsenr/vtcfwru/R/?Page=monitoR/monitoR.htmURL:
http://www.uvm.edu/rsenr/vtcfwru/R/?Page=monitoR/monitoR.htm}

\leavevmode\vadjust pre{\hypertarget{ref-hossin2015}{}}%
Hossin, M. \& Sulaiman, M. (2015). A review on evaluation metrics for
data classification evaluations. \emph{International journal of data
mining}. Retrieved from
\url{https://www.academia.edu/download/37219940/5215ijdkp01.pdf}

\leavevmode\vadjust pre{\hypertarget{ref-hsu2021}{}}%
Hsu, A. \& Yttri, E. (2021). B-SOiD, an open-source unsupervised
algorithm for identification and fast prediction of behaviors.
\emph{Nature Communications}. Retrieved from
\url{https://www.nature.com/articles/s41467-021-25420-x}

\leavevmode\vadjust pre{\hypertarget{ref-knight2017}{}}%
Knight, E.C., Hannah, K.C., Foley, G.J., Scott, C.D., Brigham, R.M. \&
Bayne, E. (2017).
\href{https://doi.org/10.5751/ACE-01114-120214}{Recommendations for
acoustic recognizer performance assessment with application to five
common automated signal recognition programs}. \emph{Avian Conservation
and Ecology, Published online: Nov 27, 2017 \textbar{}
doi:10.5751/ACE-01114-120214}, \textbf{12}.

\leavevmode\vadjust pre{\hypertarget{ref-Kohler2017}{}}%
Köhler, J., Jansen, M., Rodríguez, A., Kok, P.J.R., Felipe, T.L.,
Emmrich, M., Glaw, F., Haddad, C.F.B., Mark-Oliver, R., Vences, M.,
Toledo, L.F., Emmrich, M., Glaw, F., Haddad, C.F.B., Rödel, M.O. \&
Vences, M. (2017).
\emph{\href{https://doi.org/10.11646/zootaxa.4251.1.1}{The use of
bioacoustics in anuran taxonomy: Theory, terminology, methods and
recommendations for best practice}}.

\leavevmode\vadjust pre{\hypertarget{ref-macmillan2004}{}}%
Macmillan, N.A. \& Creelman, C.D. (2004).
\href{https://doi.org/10.4324/9781410611147/DETECTION-THEORY-NEIL-MACMILLAN-DOUGLAS-CREELMAN}{Detection
theory: A user's guide: 2nd edition}. \emph{Detection Theory: A User's
Guide: 2nd edition}, 1--445.

\leavevmode\vadjust pre{\hypertarget{ref-Medina-Garcia2015a}{}}%
Medina-García, A., Araya-Salas, M. \& Wright, T.F. (2015). Does vocal
learning accelerate acoustic diversification? Evolution of contact calls
in neotropical parrots. \emph{Journal of Evolutionary Biology},
\textbf{28}, 1782--1792. Retrieved from
\url{http://doi.wiley.com/10.1111/jeb.12694}

\leavevmode\vadjust pre{\hypertarget{ref-mellinger2000}{}}%
Mellinger, D.K. \& Clark, C.W. (2000).
\href{https://doi.org/10.1121/1.429434}{Recognizing transient
low-frequency whale sounds by spectrogram correlation}. \emph{The
Journal of the Acoustical Society of America}, \textbf{107}, 3518--3529.

\leavevmode\vadjust pre{\hypertarget{ref-Odom2021}{}}%
Odom, K.J., Araya-Salas, M., Morano, J.L., Ligon, R.A., Leighton, G.M.,
Taff, C.C., Dalziell, A.H., Billings, A.C., Germain, R.R., Pardo, M.,
Andrade, L.G. de, Hedwig, D., Keen, S.C., Shiu, Y., Charif, R.A.,
Webster, M.S. \& Rice, A.N. (2021).
\href{https://doi.org/10.1111/brv.12695}{Comparative bioacoustics: A
roadmap for quantifying and comparing animal sounds across diverse
taxa}. \emph{Biological Reviews}, \textbf{96}, 1135--1159.

\leavevmode\vadjust pre{\hypertarget{ref-schoneich2020}{}}%
S, S. (2020). Neuroethology of acoustic communication in field
crickets-from signal generation to song recognition in an insect brain.
\emph{Progress in Neurobiology}. Retrieved from
\url{https://www.sciencedirect.com/science/article/pii/S0301008220301374}

\leavevmode\vadjust pre{\hypertarget{ref-Sikes2016}{}}%
Sikes, R.S., Animal Care, the \& American Society of Mammalogists, U.C.
of the. (2016). 2016 guidelines of the american society of mammalogists
for the use of wild mammals in research and education. \emph{Journal of
Mammalogy}, \textbf{97}, 663--688. Retrieved from
\href{https://academic.oup.com/jmammal/article/97/3/663/2459909\%20http://www.mammalsociety.org/committees/animal-care-and-use\#tab3}{https://academic.oup.com/jmammal/article/97/3/663/2459909
http://www.mammalsociety.org/committees/animal-care-and-use\#tab3}

\leavevmode\vadjust pre{\hypertarget{ref-Stowell2022}{}}%
Stowell, D. (2022). Computational bioacoustics with deep learning: A
review and roadmap. \emph{PeerJ}, \textbf{10}, e13152. Retrieved from
\url{https://peerj.com/articles/13152}

\leavevmode\vadjust pre{\hypertarget{ref-sturman2020}{}}%
Sturman, O., Ziegler, L. von, Schläppi, C. \& Akyol, F. (2020). Deep
learning-based behavioral analysis reaches human accuracy and is capable
of outperforming commercial solutions. \emph{Neuropsychopharmacology}.
Retrieved from \url{https://www.nature.com/articles/s41386-020-0776-y}

\leavevmode\vadjust pre{\hypertarget{ref-sugai2019}{}}%
Sugai, L., Silva, T., Jr, J.R. \& Llusia, D. (2019). Terrestrial passive
acoustic monitoring: Review and perspectives. \emph{academic.oup.com}.
Retrieved from
\url{https://academic.oup.com/bioscience/article-abstract/69/1/15/5193506}

\leavevmode\vadjust pre{\hypertarget{ref-Tchernichovski2021}{}}%
Tchernichovski, O., Eisenberg-Edidin, S. \& Jarvis, E.D. (2021).
Balanced imitation sustains song culture in zebra finches. \emph{Nature
Communications 2021 12:1}, \textbf{12}, 1--14. Retrieved from
\url{https://www.nature.com/articles/s41467-021-22852-3}

\leavevmode\vadjust pre{\hypertarget{ref-tiwari2022}{}}%
Tiwari, C. \& Diwakar, S. (2022).
\href{https://doi.org/10.1080/09524622.2022.2064330}{The katydid
country: Bioacoustics and ecology of tettigoniid communities from the
indian subcontinent}. \emph{Bioacoustics}.

\leavevmode\vadjust pre{\hypertarget{ref-zsebok2021}{}}%
Zsebők, S., Schmera, D., Miklós Laczi, \textbar, Nagy, G., Vaskuti, É.,
János Török, \textbar, László, \textbar{} \& Garamszegi, Z. (2021). A
practical approach to measuring the acoustic diversity by community
ecology methods. \emph{Wiley Online Library}, \textbf{12}, 874--884.
Retrieved from
\url{https://besjournals.onlinelibrary.wiley.com/doi/abs/10.1111/2041-210X.13558}

\end{CSLReferences}

\end{document}

### 2023-05-24 manuscript/manuscript_v01.Rmd
many changes

--- 2023-02-14 manuscript/manuscript.Rmd
@@
       smooth_scroll: no
-  word_document:
+  pdf_document:
     toc: yes
     toc_depth: '3'
-  pdf_document:
-    toc: yes
+  word_document:
+    toc: no
     toc_depth: '3'
@@
 
+# options to customize chunk outputs
+knitr::opts_chunk$set(
+  class.source = "numberLines lineAnchors", # for code line numbers
+  tidy.opts = list(width.cutoff = 65), 
+  tidy = TRUE,
+  message = TRUE,
+  warning = FALSE
+ )
 ```
@@
 source("~/Dropbox/R_package_testing/ohun/R/split_acoustic_data.R")
+source("~/Dropbox/R_package_testing/ohun/R/diagnose_detection.R")
+source("~/Dropbox/R_package_testing/ohun/R/label_detection.R")
 split_info <- split_acoustic_data(path = path_zebra_finch, sgmt.dur = 5)
@@
 
+raven_detec <- raven_detec[, c("sound.files", "selec", "start", "end", "bottom.freq", "top.freq", "tuning_parameters")]
+
 write.csv(raven_detec, "./data/processed/combined_raven_detection.csv", row.names = FALSE)
@@
 ```{r, eval = FALSE, echo = TRUE}
-
 raven_detec <- read.csv("combined_raven_detection.csv")
 
-diagnose_detection(reference = manual_ref_tae, detection = raven_detec, by = "tuning_parameters")
+head(raven_detec)
+```
+
+```{r, eval = TRUE, echo = FALSE}
+raven_detec <- read.csv("./data/processed/combined_raven_detection.csv")
+
+head(raven_detec)
+```
+
+```{r, eval = FALSE, echo = TRUE}
+diagnose_detection(reference = manual_ref_tae,
+                   detection = raven_detec,
+                   by = "tuning_parameters",
+                   min.overlap = 0.5)
 
@@
 
-```{r, eval = TRUE, echo = TRUE}
+```{r, eval = TRUE, echo = FALSE}
 
-raven_detec <- read.csv("./data/processed/combined_raven_detection.csv")
-
-diagnose_detection(reference = manual_ref_tae, detection = raven_detec, by = "tuning_parameters")
+diagnose_detection(reference = manual_ref_tae, detection = raven_detec, by = "tuning_parameters", min.overlap = 0.5)
 
@@
 Plus three additional metrics specific to sound event detection: 'split positives', 'merge positives' and 'overlap to true positives'.
-'Split positives' is the number of reference sounds overlapped by more than one detection, 'merge positives' the number of detected sounds overlapping with another detection and 'overlap to true positives' quantifies the mean overlap between detections and reference signal (1 means complete overlap).
-The function also allows to detail those indices by sound file.
-Here we show the first 10 files:
+'Split positives' is the number of reduntant detections (i.e.
+those overlapping sounds already overlapped by other detections) by more than one detection, 'merge positives' the number of detected sounds overlapping with another detection and 'overlap to true positives' quantifies the mean overlap between detections and reference signal (1 means complete overlap).
 
-```{r, eval = TRUE, echo = TRUE}
+```{r, eval = FALSE, echo = TRUE}
+
+diag_raven <- diagnose_detection(reference = manual_ref_tae, detection = raven_detec, by = "tuning_parameters")
+
+head(diag_raven, 10)
+
+```
+
+```{r, eval = TRUE, echo = FALSE}
 
@@
 head(diag_raven, 10)
+
+```
+
+The function also allows to detail those indices by sound file. Here we show the first 10 files:
+
+```{r, eval = FALSE, echo = TRUE}
+
+diag_raven <- diagnose_detection(reference = manual_ref_tae, detection = raven_detec, by = "tuning_parameters", by.sound.file = TRUE)
+
+head(diag_raven, 10)
+```
+
+```{r, eval = TRUE, echo = FALSE}
+
+diag_raven <- diagnose_detection(reference = manual_ref_tae, detection = raven_detec, by = "tuning_parameters", by.sound.file = TRUE)
+
+head(diag_raven, 10)
+
 ```
@@
 
-# opt_detec_train <- opt_detec_train[order(opt_detec_train$templates, opt_detec_train$f1.score, decreasing = TRUE), ]
+# opt_detec_train <- opt_detec_train[order(opt_detec_train$templates, opt_detec_train$f.score, decreasing = TRUE), ]
 
-# sub_opt_detec_train <- opt_detec_train[ave(-opt_detec_train$f1.score, opt_detec_train$templates, FUN = rank) <= 4, ]
+# sub_opt_detec_train <- opt_detec_train[ave(-opt_detec_train$f.score, opt_detec_train$templates, FUN = rank) <= 4, ]
 
@@
     X <- opt_detec_train[opt_detec_train$templates == x, ]
-    X <- X[order(X$f1.score, decreasing = TRUE), ]
+    X <- X[order(X$f.score, decreasing = TRUE), ]
     return(X[1:2, ])
@@
 
-kbl <- kable(sub_opt_detec_train[, c("threshold", "templates", "true.positives", "false.positives", "false.negatives", "recall", "precision", "f1.score")], row.names = FALSE, caption = "TABLE TEMPLATE PERFORMANCE.
+kbl <- kable(sub_opt_detec_train[, c("threshold", "templates", "true.positives", "false.positives", "false.negatives", "recall", "precision", "f.score")], row.names = FALSE, caption = "TABLE TEMPLATE PERFORMANCE.
 Performance diagnostic of template-based detections using four templates across several threshold values.
@@
 
-We can explore the performance of each template in more detail by looking at the change in F1 score across thresholds (FIG. THRESHOLD vs F1.SCORE).
+We can explore the performance of each template in more detail by looking at the change in F1 score across thresholds (FIG. THRESHOLD vs F.SCORE).
 
-```{r, eval = TRUE, echo = FALSE, fig.height= 3, fig.align='center', fig.cap= "FIG. THRESHOLD vs F1.SCORE. Shows the changes in F1 score across the range of threshold values"}
+```{r, eval = TRUE, echo = FALSE, fig.height= 3, fig.align='center', fig.cap= "FIG. THRESHOLD vs F.SCORE. Shows the changes in F1 score across the range of threshold values"}
 
-agg_f1 <- aggregate(f1.score ~ threshold + templates, data = opt_detec_train,
+agg_f1 <- aggregate(f.score ~ threshold + templates, data = opt_detec_train,
 mean)
 
-ggplot(agg_f1, aes(x = threshold, y = f1.score, group = templates, color = templates)) +
+ggplot(agg_f1, aes(x = threshold, y = f.score, group = templates, color = templates)) +
     geom_line() + 
@@
 In this example the "centroid" template, produced the best performance (TABLE TEMPLATE PERFORMANCE; FIG.
-THRESHOLD vs F1.SCORE).
+THRESHOLD vs F.SCORE).
 Hence, we will use this template for detecting calls on the rest of the data.
@@
 diagnose_detection(reference = test_ref, detection = detec_test)
+
 ```
@@
   max.duration = c(275, 300, 325), 
-  bp = c(0.5, 10)
+  bp = c(0.5, 10),
+  path = data_path
 )
+
+```
+
+```{r, eval = FALSE, echo = FALSE}
+write.csv(opt_det_train , file = "./data/processed/taeniopygia/detection_optimization.csv")
 
@@
 # subset with highest performance
-opt_det_train <- opt_det_train[order(opt_det_train$f1.score, decreasing = TRUE), ]
+opt_det_train <- opt_det_train[order(opt_det_train$f.score, decreasing = TRUE), ]
 
@@
 # subset with highest performance
-sub_opt_det_train <- opt_det_train[order(opt_det_train$f1.score, decreasing = TRUE), ]
+sub_opt_det_train <- opt_det_train[order(opt_det_train$f.score, decreasing = TRUE), ]
 
-sub_opt_det_train <- sub_opt_det_train[1:10, c("threshold", "smooth", "hold.time", "min.duration", "max.duration", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f1.score")]
+sub_opt_det_train <- sub_opt_det_train[1:10, c("threshold", "smooth", "hold.time", "min.duration", "max.duration", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f.score")]
 
-sub_opt_det_train
+head(sub_opt_det_train, 10)
 ```
@@
 
-best_param <- opt_det_train[which.max(opt_det_train$f1.score), ]
+best_param <- opt_det_train[which.max(opt_det_train$f.score), ]
 
@@
   max.duration = best_param$max.duration, 
-  bp = c(0.5, 10)
+  bp = c(0.5, 10),
+  path = data_path, cores = 10 
 )
@@
 More training data might be needed for optimizing a detection routine on larger data sets or recordings with more variable sounds or background noise levels.
+
+Using macro averaging:
+
+```{r, diagnose 2, eval = TRUE, echo = TRUE}
+
+diagnose_detection(reference = test_ref,
+                   detection = det_test,
+                   by.sound.file = FALSE,
+                   macro.average = TRUE)
+
+```
 

### 2023-05-24 scripts/study_case_spixs_disc-winged_bat_calls.Rmd
update figs and manuscript
many changes

--- 2022-07-13 scripts/study_case_thyroptera.Rmd
@@
 ---
-title: <center><font size="6"><b>Template-based automatic detection on Spix's disc-winged bat social calls</b></font></center>
-subtitle: <center><font size="4"><b>ohun</b></font></center>
-author: <center><font size="4"><a href="http://marceloarayasalas.weebly.com/">Marcelo Araya-Salas, PhD</a></font></center>
-date: <center>`r format(Sys.Date(), "%d-%m-%Y")`</center>
+title: <font size="6"><b>Template-based automatic detection on Spix's disc-winged bat social calls</b></font>
+subtitle: <font size="4"><b>ohun</b></font>
+author: <font size="4"><a href="http://marceloarayasalas.weebly.com/">Marcelo Araya-Salas, PhD</a></font>
+date:  "`r Sys.Date()`"
 output:
@@
 
+<div class="alert alert-warning">
+
+### Data set
+
+The data used in this tutorial (annotations and recordings) can be downloaded from here:
+
+- [https://figshare.com/ndownloader/files/38431409](https://figshare.com/ndownloader/files/38431409)
+
+</div>
+
+&nbsp;
+
 ---
 
-### Load packages
+### Package installation
+
+Install/load the package from CRAN as follows:
+
+```{r, eval = FALSE}
+
+# From CRAN would be
+install.packages("ohun")
+
+#load package
+library(ohun)
+
+```
+
+To install the latest developmental version from [github](https://github.com/) you will need the R package [remotes](https://cran.r-project.org/package=remotes):
+
+```{r, eval = FALSE}
+remotes::install_github("maRce10/ohun")
+
+#load package
+library(ohun)
+
+```
+
+### Load additional packages
 ```{r packages, message = FALSE, warning = FALSE, echo = FALSE, eval = TRUE, message = FALSE, warning=FALSE}
@@
 source("~/Dropbox/R_package_testing/ohun/R/get_templates.R")
+source("~/Dropbox/R_package_testing/ohun/R/internal_functions.R")
 ```
@@
 ```{r, message=FALSE, warning=FALSE}
-library(ohun)
 
@@
 
-This code creates a multipanel image with multiple spectrograms, one for each of the individuals/recordings in the complete data set
+This code creates a multipanel image with multiple spectrograms, one for each of the individuals/recordings in the complete data set (note that the output of the function `catalog` is a image file in the current working directory):
 ```{r Create catalogs, eval = FALSE}
@@
 # create catalogs
-catalog(X = high_snr, flim = c(10, 45), nrow = 6, ncol = 5, ovlp = 90, height = 10, width = 20, same.time.scale = TRUE, mar = 0.005, wl = 512, gr = FALSE, spec.mar = 0, lab.mar = 0.001, rm.axes = TRUE, by.row = TRUE, box = FALSE, pal = viridis, collevels = seq(-100, 0, 5), labels = NULL)
+catalog(X = high_snr, flim = c(10, 45), nrow = 6, ncol = 5, ovlp = 90, height = 10, width = 20, same.time.scale = TRUE, mar = 0.005, wl = 512, gr = FALSE, spec.mar = 0, lab.mar = 0.001, rm.axes = TRUE, by.row = TRUE, box = FALSE, pal = viridis, collevels = seq(-100, 0, 5), labels = NULL, path = data_path)
 
@@
 
-```{r catalog, out.width = "100%", echo = FALSE, fig.align= "center"}
-
-knitr::include_graphics(normalizePath(file.path(data_path, "Catalog_p1.jpeg")))
-
-```
+![](../data/raw/thyroptera/Catalog_p1.jpeg)
 
@@
 ```{r get templates, eval = TRUE, echo = TRUE, fig.height=7}
+
 # find templates
 templates <- get_templates(train_ref, path = data_path,  bp = c(10, 50), fast = TRUE, ovlp = , wl = 200, n.sub.spaces = 3)
+
 ```
@@
 ```{r get templates 2, eval = FALSE, echo = TRUE}
+
 # create ext. selection table
@@
 # create catalogs
-catalog(X = templates_est, flim = c(10, 45), nrow = 2, ncol = 2, ovlp = 90, height = 10, width = 15, same.time.scale = TRUE, mar = 0.005, wl = 512, gr = FALSE, spec.mar = 0.4, lab.mar = 0.8, rm.axes = TRUE, by.row = TRUE, box = TRUE, pal = viridis, parallel = 10, collevels = seq(-100, 0, 5), img.prefix = "templates")
+catalog(X = templates_est, flim = c(10, 45), nrow = 2, ncol = 2, ovlp = 90, height = 10, width = 15, same.time.scale = TRUE, mar = 0.005, wl = 512, gr = FALSE, spec.mar = 0.4, lab.mar = 0.8, rm.axes = TRUE, by.row = TRUE, box = TRUE, pal = viridis, parallel = 10, collevels = seq(-100, 0, 5), img.prefix = "templates", path = data_path)
 
@@
 
-knitr::include_graphics(normalizePath(file.path(data_path, "templates-Catalog_p1.jpeg")))
+knitr::include_graphics("./data/raw/thyroptera/templates-Catalog_p1.jpeg")
 

### 2023-05-24 scripts/study_case_spixs_disc-winged_bat_calls.qmd
many changes

--- 2023-05-24 scripts/study_case_spixs_disc-winged_bat_calls.Rmd
@@
 ---
-title: <font size="6"><b>Template-based automatic detection on Spix's disc-winged bat social calls</b></font>
-subtitle: <font size="4"><b>ohun</b></font>
-author: <font size="4"><a href="http://marceloarayasalas.weebly.com/">Marcelo Araya-Salas, PhD</a></font>
+title: Template-based automatic detection on Spix's disc-winged bat social calls
+subtitle: ohun
+author: <a href="http://marceloarayasalas.weebly.com/">Marcelo Araya-Salas, PhD</a>
 date: "`r Sys.Date()`"
-output:
-  html_document:
-    toc: yes
-    toc_depth: 3
-    toc_float:
-      collapsed: no
-      smooth_scroll: no
-fontsize: 12pt 
-editor_options: 
-  chunk_output_type: console
+toc: true
+toc-depth: 2
+toc-location: left
+number-sections: true
+highlight-style: pygments
+format:
+  html:
+    df-print: kable
+    code-fold: show
+    code-tools: true
+    css: qmd.css
 ---
@@
 
-opts_chunk$set( fig.width = 8, fig.height = 3.5)
+opts_chunk$set( fig.width = 8, fig.height = 3.5, warning = FALSE)
 
@@
 
-source("~/Dropbox/R_package_testing/ohun/R/get_templates.R")
+source("~/Dropbox/R_package_testing/ohun/R/diagnose_detection.R")
+source("~/Dropbox/R_package_testing/ohun/R/label_detection.R")
+source("~/Dropbox/R_package_testing/ohun/R/summarize_diagnostic.R")
 source("~/Dropbox/R_package_testing/ohun/R/internal_functions.R")
@@
 
-This code creates a multipanel image with multiple spectrograms, one for each of the individuals/recordings in the complete data set (note that the output of the function `catalog` is a image file in the current working directory):
+This code creates a multipanel image with multiple spectrograms, one for each of the individuals/recordings in the complete data set (note that the output of the function `catalog` are image files saved in the sound file directory):
 ```{r Create catalogs, eval = FALSE}
@@
 
-![](../data/raw/thyroptera/Catalog_p1.jpeg)
+![](./img/Catalog_thyroptera.jpeg)
 
@@
 
-```{r catalog templates, out.width = "70%", echo = FALSE, fig.align= "center"}
-
-knitr::include_graphics("./data/raw/thyroptera/templates-Catalog_p1.jpeg")
-
-```
+![](./img/templates-Catalog_p1.jpeg)
 
@@
   files = unique(train_ref$sound.files), 
-  parallel = paral, 
+  cores = paral, 
   hop.size = 10, 
@@
   threshold = seq(0.1, 0.9, 0.05), 
-  parallel = paral
+  cores = paral
   )
@@
 
-### Create plot with threshold vs f1 score
+### Create plot with threshold vs F score
 ```{r, eval = TRUE, echo = FALSE}
@@
 
-agg_f1 <- aggregate(f1.score ~ threshold + templates, data = opt_detec_train,
+agg_f <- aggregate(f.score ~ threshold + templates, data = opt_detec_train,
 mean)
 
-ggplot(agg_f1, aes(x = threshold, y = f1.score, group = templates, color = templates)) +
+ggplot(agg_f, aes(x = threshold, y = f.score, group = templates, color = templates)) +
 geom_line() + geom_point() + scale_color_viridis_d(end = 1) + theme_classic()
@@
 
-### Check subset with highest performance (sorted by f1 score)
+### Check subset with highest performance (sorted by F score)
 
@@
 
-opt_detec_train <- opt_detec_train[order(opt_detec_train$f1.score, decreasing = TRUE), ]
+opt_detec_train <- opt_detec_train[order(opt_detec_train$f.score, decreasing = TRUE), ]
 
-sub_opt_detec_train <- opt_detec_train[1:10, c("templates","threshold", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f1.score")]
+sub_opt_detec_train <- opt_detec_train[1:10, c("templates","threshold", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f.score")]
 
@@
 # subset with highest performance
-opt_detec_train <- opt_detec_train[order(opt_detec_train$f1.score, decreasing = TRUE), ]
+opt_detec_train <- opt_detec_train[order(opt_detec_train$f.score, decreasing = TRUE), ]
 
-sub_opt_detec_train <- opt_detec_train[1:10, c("templates","threshold", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f1.score")]
+sub_opt_detec_train <- opt_detec_train[1:10, c("templates","threshold", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f.score")]
 
@@
 
-dd$f1.score
+dd$f.score
 
@@
 
-- Good performance on detecting Spix's disc-winged social calls: F1 score was `r round(dd$f1.score, 3)` for the training data set and `r round(max(opt_detec_train$f1.score), 3)` for the testing data
+- Good performance on detecting Spix's disc-winged social calls: F score was 0.953 for the training data set and 0.955 for the testing data
 

### 2023-05-24 scripts/study_case_zebra_finch_songs.Rmd
update figs and manuscript
many changes

--- 2022-07-13 scripts/study_case_taeniopygia.Rmd
@@
 ---
-title: <center><font size="6"><b>Energy-based automatic detection on zebra finch songs</b></font></center>
-subtitle: <center><font size="4"><b>ohun</b></font></center>
-author: <center><font size="4"><a href="http://marceloarayasalas.weebly.com/">Marcelo Araya-Salas, PhD</a></font></center>
-date: <center>`r format(Sys.Date(), "%d-%m-%Y")`</center>
+title: <font size="6"><b>Energy-based automatic detection on zebra finch songs</b></font>
+subtitle: <font size="4"><b>ohun</b></font>
+author: <font size="4"><a href="http://marceloarayasalas.weebly.com/">Marcelo Araya-Salas, PhD</a></font>
+date: "`r Sys.Date()`"
 output:
@@
 
+<div class="alert alert-warning">
+
+### Data set
+
+The data used in this tutorial (annotations and recordings) can be downloaded from here:
+
+- [https://figshare.com/ndownloader/files/38431340](https://figshare.com/ndownloader/files/38431340)
+
+</div>
+
+&nbsp;
+
 ```{r setup, include = FALSE}
@@
 
-### Load package
+### Package installation
+
+Install/load the package from CRAN as follows:
+
+```{r, eval = FALSE}
+
+# From CRAN would be
+install.packages("ohun")
+
+#load package
+library(ohun)
+
+```
+
+To install the latest developmental version from [github](https://github.com/) you will need the R package [remotes](https://cran.r-project.org/package=remotes):
+
+```{r, eval = FALSE}
+remotes::install_github("maRce10/ohun")
+
+#load package
+library(ohun)
+
+```
+
+### Load additional packages
 ```{r libraries, eval = TRUE, echo = FALSE, message=FALSE, warning=FALSE}
@@
 
-knitr::include_graphics(normalizePath(file.path(data_path, "Catalog_p1.jpeg")))
+# knitr::include_graphics(normalizePath(file.path(data_path, "Catalog_p1.jpeg")))
 
@@
 
+![](../data/raw/taeniopygia/Catalog_p1.jpeg)
 &nbsp;

### 2023-05-24 scripts/study_case_zebra_finch_songs.qmd
many changes

--- 2023-05-24 scripts/study_case_zebra_finch_songs.Rmd
@@
 ---
-title: <font size="6"><b>Energy-based automatic detection on zebra finch songs</b></font>
-subtitle: <font size="4"><b>ohun</b></font>
-author: <font size="4"><a href="http://marceloarayasalas.weebly.com/">Marcelo Araya-Salas, PhD</a></font>
+title: Energy-based automatic detection on zebra finch songs
+subtitle: ohun
+author: <a href="http://marceloarayasalas.weebly.com/">Marcelo Araya-Salas, PhD</a>
 date: "`r Sys.Date()`"
-output:
-  html_document:
-    toc: yes
-    toc_depth: 3
-    toc_float:
-      collapsed: no
-      smooth_scroll: no
-fontsize: 12pt 
-editor_options: 
-  chunk_output_type: console
+toc: true
+toc-depth: 2
+toc-location: left
+number-sections: true
+highlight-style: pygments
+format:
+  html:
+    df-print: kable
+    code-fold: show
+    code-tools: true
+    css: qmd.css
 ---
@@
 
-knitr::opts_knit$set(root.dir = normalizePath(".."))
+knitr::opts_knit$set(root.dir = "..")
 
-knitr::opts_chunk$set(include = TRUE, eval = TRUE)
+knitr::opts_chunk$set(include = TRUE, eval = TRUE, warning = FALSE)
 
@@
 
+source("~/Dropbox/R_package_testing/ohun/R/diagnose_detection.R")
+source("~/Dropbox/R_package_testing/ohun/R/label_detection.R")
+source("~/Dropbox/R_package_testing/ohun/R/summarize_diagnostic.R")
+source("~/Dropbox/R_package_testing/ohun/R/internal_functions.R")
 ```
@@
 
-This code creates a multipanel image with multiple spectrograms, one for each of the individuals/recordings in the complete data set
+This code creates a multipanel image with multiple spectrograms, one for each of the individuals/recordings in the complete data set  (note that the output of the function `catalog` are image files saved in the sound file directory):
 ```{r Create catalogs, eval = FALSE}
@@
 
-```{r catalog, out.width = "100%", echo = FALSE, fig.align= "center"}
+![](img/Catalog_taenopygia.jpeg)
 
-# knitr::include_graphics(normalizePath(file.path(data_path, "Catalog_p1.jpeg")))
-
-```
-
-![](../data/raw/taeniopygia/Catalog_p1.jpeg)
 &nbsp;
@@
 
-### Check subset with highest performance (sorted by f1 score)
+### Check subset with highest performance (sorted by F score)
 ```{r, eval = FALSE}
@@
 # subset with highest performance
-sub_opt_det <- opt_det[order(opt_det$f1.score, decreasing = TRUE), ]
+sub_opt_det <- opt_det[order(opt_det$f.score, decreasing = TRUE), ]
 
-sub_opt_det <- sub_opt_det[1:10, c("threshold", "smooth", "hold.time", "min.duration", "max.duration", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f1.score")]
+sub_opt_det <- sub_opt_det[1:10, c("threshold", "smooth", "hold.time", "min.duration", "max.duration", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f.score")]
 
@@
 
-### Select the highest performance parameters based on f1 score
+### Select the highest performance parameters based on F score
 ```{r, select best params, eval = TRUE, echo = TRUE}
 
-best_param <- opt_det[which.max(opt_det$f1.score), ]
+best_param <- opt_det[which.max(opt_det$f.score), ]
 
@@
 
-- Good performance on detecting zebra finch songs: F1 score was 0.95 for the training data set and 0.96 for the testing data
+- Good performance on detecting zebra finch songs: F score was 0.95 for the training data set and 0.96 for the testing data
 
